### 2024-01-11 paper/main.tex
initial commit of paper

%%%%%%%% ICML 2024 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2024} with \usepackage[nohyperref]{icml2024} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2024}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2024}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Sample-efficient Reinforcement Learning with Latent Representations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CUSTOM (our stuff)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Latin
\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.\@}\xspace}
\newcommand{\ie}{\textit{i.e.\@}\xspace}
\newcommand{\cf}{\textit{cf.\@}\xspace}
\newcommand{\etc}{\textit{etc.\@}\xspace}
\newcommand{\etal}{\textit{et~al.\@}\xspace}

% Vectors
\def\vzero{{\bm{0}}}
\def\vone{{\bm{1}}}
\def\vmu{{\bm{\mu}}}
\def\vtheta{{\bm{\theta}}}
\def\va{{\bm{a}}}
\def\vb{{\bm{b}}}
\def\vc{{\bm{c}}}
\def\vd{{\bm{d}}}
\def\ve{{\bm{e}}}
\def\vf{{\bm{f}}}
\def\vg{{\bm{g}}}
\def\vh{{\bm{h}}}
\def\vi{{\bm{i}}}
\def\vj{{\bm{j}}}
\def\vk{{\bm{k}}}
\def\vl{{\bm{l}}}
\def\vm{{\bm{m}}}
\def\vn{{\bm{n}}}
\def\vo{{\bm{o}}}
\def\vp{{\bm{p}}}
\def\vq{{\bm{q}}}
\def\vr{{\bm{r}}}
\def\vs{{\bm{s}}}
\def\vt{{\bm{t}}}
\def\vu{{\bm{u}}}
\def\vv{{\bm{v}}}
\def\vw{{\bm{w}}}
\def\vx{{\bm{x}}}
\def\vy{{\bm{y}}}
\def\vz{{\bm{z}}}

% Elements of vectors
\def\evalpha{{\alpha}}
\def\evbeta{{\beta}}
\def\evepsilon{{\epsilon}}
\def\evlambda{{\lambda}}
\def\evomega{{\omega}}
\def\evmu{{\mu}}
\def\evpsi{{\psi}}
\def\evsigma{{\sigma}}
\def\evtheta{{\theta}}
\def\eva{{a}}
\def\evb{{b}}
\def\evc{{c}}
\def\evd{{d}}
\def\eve{{e}}
\def\evf{{f}}
\def\evg{{g}}
\def\evh{{h}}
\def\evi{{i}}
\def\evj{{j}}
\def\evk{{k}}
\def\evl{{l}}
\def\evm{{m}}
\def\evn{{n}}
\def\evo{{o}}
\def\evp{{p}}
\def\evq{{q}}
\def\evr{{r}}
\def\evs{{s}}
\def\evt{{t}}
\def\evu{{u}}
\def\evv{{v}}
\def\evw{{w}}
\def\evx{{x}}
\def\evy{{y}}
\def\evz{{z}}

% Matrix
\def\mA{{\bm{A}}}
\def\mB{{\bm{B}}}
\def\mC{{\bm{C}}}
\def\mD{{\bm{D}}}
\def\mE{{\bm{E}}}
\def\mF{{\bm{F}}}
\def\mG{{\bm{G}}}
\def\mH{{\bm{H}}}
\def\mI{{\bm{I}}}
\def\mJ{{\bm{J}}}
\def\mK{{\bm{K}}}
\def\mL{{\bm{L}}}
\def\mM{{\bm{M}}}
\def\mN{{\bm{N}}}
\def\mO{{\bm{O}}}
\def\mP{{\bm{P}}}
\def\mQ{{\bm{Q}}}
\def\mR{{\bm{R}}}
\def\mS{{\bm{S}}}
\def\mT{{\bm{T}}}
\def\mU{{\bm{U}}}
\def\mV{{\bm{V}}}
\def\mW{{\bm{W}}}
\def\mX{{\bm{X}}}
\def\mY{{\bm{Y}}}
\def\mZ{{\bm{Z}}}
\def\mBeta{{\bm{\beta}}}
\def\mPhi{{\bm{\Phi}}}
\def\mLambda{{\bm{\Lambda}}}
\def\mSigma{{\bm{\Sigma}}}

\newcommand{\E}{\mathbb{E}}
\newcommand{\Ls}{\mathcal{L}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\emp}{\tilde{p}}
\newcommand{\lr}{\alpha}
\newcommand{\reg}{\lambda}
\newcommand{\rect}{\mathrm{rectifier}}
\newcommand{\softmax}{\mathrm{softmax}}
\newcommand{\sigmoid}{\sigma}
\newcommand{\softplus}{\zeta}
\newcommand{\KL}{D_{\mathrm{KL}}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\standarderror}{\mathrm{SE}}
\newcommand{\Cov}{\mathrm{Cov}}
% Wolfram Mathworld says $L^2$ is for function spaces and $\ell^2$ is for vectors
% But then they seem to use $L^2$ for vectors throughout the site, and so does
% wikipedia.
\newcommand{\normlzero}{L^0}
\newcommand{\normlone}{L^1}
\newcommand{\normltwo}{L^2}
\newcommand{\normlp}{L^p}
\newcommand{\normmax}{L^\infty}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\twocolumn[
% \icmltitle{Submission and Formatting Instructions for \\
%            International Conference on Machine Learning (ICML 2024)}
\icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2024
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Aidan Scannell}{aalto,fcai}
\icmlauthor{Arno Solin}{aalto}
\icmlauthor{Joni Pajarinen}{aalto}
% \icmlauthor{Firstname1 Lastname1}{equal,yyy}
% \icmlauthor{Firstname2 Lastname2}{equal,yyy,comp}
% \icmlauthor{Firstname3 Lastname3}{comp}
% \icmlauthor{Firstname4 Lastname4}{sch}
% \icmlauthor{Firstname5 Lastname5}{yyy}
% \icmlauthor{Firstname6 Lastname6}{sch,yyy,comp}
% \icmlauthor{Firstname7 Lastname7}{comp}
% %\icmlauthor{}{sch}
% \icmlauthor{Firstname8 Lastname8}{sch}
% \icmlauthor{Firstname8 Lastname8}{yyy,comp}
%\icmlauthor{}{sch}
%\icmlauthor{}{sch}
\end{icmlauthorlist}

\icmlaffiliation{aalto}{Aalto University, Finland}
\icmlaffiliation{fcai}{Finnish Center for Artificial Intelligence}
% \icmlaffiliation{yyy}{Department of XXX, University of YYY, Location, Country}
% \icmlaffiliation{comp}{Company Name, Location, Country}
% \icmlaffiliation{sch}{School of ZZZ, Institute of WWW, Location, Country}

\icmlcorrespondingauthor{Aidan Scannell}{aidan.scannell@aalto.fi}
% \icmlcorrespondingauthor{Firstname1 Lastname1}{first1.last1@xxx.edu}
% \icmlcorrespondingauthor{Firstname2 Lastname2}{first2.last2@www.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
% \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
This document provides a basic paper template and submission guidelines.
Abstracts must be a single paragraph, ideally between 4--6 sentences long.
Gross violations will trigger corrections at the camera-ready phase.
\end{abstract}

\section{Introduction}
\label{intro}

\section{Related Work}
\label{related_work}

\textbf{Sample-efficiency in deep RL}
Many approaches have been proposed to improve sample efficiency in deep RL.
As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.

\textbf{Representation learning for RL}
Learning representations for RL has been investigated for decades
\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
However, these approaches are usually limited to simple environments.
Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
\cite{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.

\citet{yaratsImprovingSampleEfficiency2021} learn a compact representation with a deterministic autoencoder and then train a
SAC agent \citep{haarnojaSoft2018} with the representation. They name their method SAC-AE.
Interestingly, SAC-AE matches the performance of model-based methods like PlaNet \citep{hafnerLearning2019}
and SLAC \citep{leeStochasticLatentActorCritic2020}, whilst being considerably simpler to implement.

\textbf{Bisimulation}
DeepMDP \citep{geladaDeepMDPLearningContinuous2019},
deep bisimulation for control (DBC) \citep{zhangLearningInvariantRepresentations2020}
and DHGP \citep{rezaei-shoshtariContinuousMDPHomomorphisms2022}
% \citet{geladaDeepMDPLearningContinuous2019,zhangLearningInvariantRepresentations2020,rezaei-shoshtariContinuousMDPHomomorphisms2022}
use SSL losses based on bisimulation metrics \citep{larsenBisimulationProbabilisticTesting1989}.
As such, their encoders group states into clusters whilst preserving some properties
(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
These methods use latent-space dynamic models to learn latent representations which encode only task-relevant information.
% Interestingly, the dynamics models are only used for representation learning and are not used for model-based planning.
DeepMDP \citep{geladaDeepMDPLearningContinuous2019} combines its representation learning with
distributional Q learning  \citep[C51,][]{bellemareDistributionalPerspectiveReinforcement2017},
whilst DBC \citep{zhangLearningInvariantRepresentations2020} combine its method with SAC.

% \cite{zhaoSimplifiedTemporalConsistency2023}
% \cite{hansenTemporalDifferenceLearning2022}

% \cite{higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015} use VAE \citep{kingmaAutoEncoding2014}

% \citet{zhangLearningInvariantRepresentations2020} to learn a latent representation which encode only
% task-relevant information.
% Their method is named deep bisimulation for control (DBC) and they combine their representation learning method with SAC.

\textbf{Primacy bias}
Primacy bias in deep RL \citep{nikishinPrimacyBiasDeep2022}
is where an agent overfits on early data and loses its ability to learn from new data,
which arises due to the plasticity in NNs \cite{lyleUnderstandingPlasticityNeural2023} \todo{earlier citation of plasticiy in NNs??}.
\cite{nikishinPrimacyBiasDeep2022} showed that periodically resetting the agent's parameters can help alleviate primacy bias.
\cite{doroSampleEfficientReinforcementLearning2022} built on top of this insight and showed that resetting the
agent's parameters not only alleviates primacy bias but also
enables the number of parameter update's per environment interaction to be increased.
They refer to this as the replay ratio but it has also been referred to as the update to data
ratio (UTD) \citep{chenRandomizedEnsembledDouble2021}.
\citet{nikishinDeepReinforcementLearning2023} propose an alternative approach to overcome primacy bias which
freezes the current NN and creates a new one which learns the changes in predictions.

When leveraging latent representations for RL (\eg SLAC \citep{leeStochasticLatentActorCritic2020},
SAC-AE \citep{yaratsImprovingSampleEfficiency2021}) this issue is exacerbated.
Not only do the agents networks (\eg actor/critic) overfit on early data, but as the encoder is updated during training, the
inputs (latent states) that the networks are trained on change.
As such, the networks are overfit on latent states which no longer exist in the ``latent'' data set.

\textbf{Misc}
\citet{obando-ceronSmallBatchDeep2023a} show that reducing the batch size not only improves
performance and wall-time gains, but also appears to have knock-on effects on exploration and asymptotic behaviour.

\citet{lyuOffPolicyRLAlgorithms2023} show that performing multiple parameter updates for a given batch.
Their method, name sample multiple reuse (SMR), improves sample efficiency
They say not to use if for image-based environments like Atari.

\section{Method}
\label{sec:method}

In humanoid-run and dog walk compare
\begin{itemize}
    \item Show that resetting is beneficial
    \begin{itemize}
        \item But picking when to reset varies across environments
    \end{itemize}
    \item DDPG
    \item DDPG reset every X param updates
    \item DDPG-AE
    \item DDPG-AE reset every X param updates
    \item DDPG-AE auto-resets
    \item DDPG-TAE
    \item DDPG-TAE reset every X param updates
    \item DDPG-TAE auto-resets
\end{itemize}

\text{Ablations}

\begin{itemize}
    \item Partial vs full resets vs $50\%$ resets etc
    \item Interleaved training vs representation first
    \begin{itemize}
        \item Show that representation first is better when resetting
        \item But interleaved better if not resetting
    \end{itemize}
    \item Normalized latent (SimNorm) vs unnormalized
\end{itemize}

\begin{itemize}
    \item Our method for auto-resetting handles sensors breaking during agents lifetime.
    \item Show this by transforming agents observations at some point during training.
\end{itemize}

\begin{itemize}
    \item When to reset actor/critic/encoder?
    \item How to pick UTD?
    \item How to pick learning rate?
\end{itemize}

Simplical embeddings
\cite{lavoieSimplicialEmbeddingsSelfSupervised2022}

\section{Theoretical Analysis}
% \label{}

Effect of domain shift on policy due to encoder being learned online

\citet{zhangLearningInvariantRepresentations2020} have proofs for bounds on value functions with encoder.
We could probably start with some of these.
They build off the idea of bisimulation \citep{larsenBisimulationProbabilisticTesting1989}.
Where encoders group states into clusters whilst preserving properties
(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.

\textbf{Representation learning improves sample efficiency in RL}
\begin{theorem}
\label{thm:bigtheorem}
Given two MDPs $\left( \mathcal{S}_{1}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$ and $\left( \mathcal{S}_{2}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$
with $|\mathcal{S}_{1}| < |\mathcal{S}_{2}|$,
the optimal value function $V : \mathcal{S} \rightarrow \R$ in $\mathcal{S}_{1}$ can be learned with less environment samples than in $\mathcal{S}_{2}$.
\end{theorem}

\begin{proof}
Left as an exercise to the reader.
\end{proof}

\begin{corollary}
If $f:X\to Y$ is bijective,
the cardinality of $X$ is at least as large as that of $Y$.
\end{corollary}

\begin{assumption}
The state space $\mathcal{S}$ is discrete
\label{ass:discrete-state-space}
\end{assumption}

\begin{itemize}
    \item Prove latent space improves sample efficiency
    \begin{itemize}
        \item Assume discrete state space
        \item Assume latent space contains all information from state space needed to predict values
        \item Show smaller state space can be learned with less samples
    \end{itemize}
    \item Prove retraining from scratch is better than continuing training
    \begin{itemize}
        \item Training encoder episode to episode is like non-IID training
    \end{itemize}
\end{itemize}

\begin{itemize}
    \item Assume latent space contains all information from state space needed to predict values
    \item Show retraining better contains information from state space
    \begin{itemize}
        \item Train agent on $\mathcal{D}^{i}$ with old encoder $e_{\phi^{i-1}}$ and compare to new encoder $e_{\phi^{i}}$
    \end{itemize}
\end{itemize}

\section{Experiments}
\label{experiments}

SAC-AE is just SAC but using AE to get latent.
BDC use bisimulation metric so they can obtain bound on the suboptimality of a value
function trained on their representation.

\begin{itemize}
    \item Compare to SAC-AE
    \item Compare to TCRL
    \item Compare to SLAC
    \item Compare to SAC-BDC \citet{zhangLearningInvariantRepresentations2020}
    \item Compare to DeepMDP
    \item Compare to DHGP
\end{itemize}

\begin{itemize}
    \item Show we can use higher UTD?
    \item AE vs VAE vs VQ-VAE
    \item Full reset vs partial reset
    \begin{itemize}
        \item Reset last layer in MuJoCo
        \item Reset a percentage (\eg 50\%) on image observations
    \end{itemize}
    \item Compare to resetting every X parameter updates like in \citep{doroSampleEfficientReinforcementLearning2022}.
    \item How much training should we do after parameter reset?
    \item MuJoCo and Atari
\end{itemize}

\section{Conclusion}
\label{conclusion}

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite

\bibliography{zotero-library}
\bibliographystyle{icml2024}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn
\section{You \emph{can} have an appendix here.}

You can have as much text here as you want. The main body must be at most $8$ pages long.
For the final version, one more page can be added.
If you want, you can use an appendix like this one.

The $\mathtt{\backslash onecolumn}$ command above can be kept in place if you prefer a one-column appendix, or can be removed if you prefer a two-column appendix.  Apart from this possible change, the style (font size, spacing, margins, page numbering, etc.) should be kept the same as the main body.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 and 2024 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2024-01-28 agents/vq_tc_td3.py
fix

@@ -1012,7 +1012,7 @@ class VQ_TC_TD3(Agent):
 
                 # Losses
                 rho = self.rho**t
-                if self.use_fsq:
+                if self.use_fsq and not self.project_latent:
                     next_z_pred = next_z_pred.flatten(-2)
                     next_z_tar = next_z_tar.flatten(-2)
                 if self.use_cosine_similarity_dynamics:

### 2024-01-28 agents/vq_tc_td3.py
put projection compile after target copy params

@@ -537,8 +537,6 @@ class VQ_TC_TD3(Agent):
                 num_codes=fsq_num_codes,
                 out_dim=projection_dim,
             ).to(device)
-            if compile:
-                self.projection = torch.compile(self.projection, mode="default")
             encoder_params += list(self.projection.parameters())
 
             self.projection_target = Projection(
@@ -549,6 +547,7 @@ class VQ_TC_TD3(Agent):
             ).to(device)
             self.projection_target.load_state_dict(self.projection.state_dict())
             if compile:
+                self.projection = torch.compile(self.projection, mode="default")
                 self.projection_target = torch.compile(
                     self.projection_target, mode="default"
                 )

### 2024-01-28 agents/vq_tc_td3.py
add params after compile

@@ -537,8 +537,6 @@ class VQ_TC_TD3(Agent):
                 num_codes=fsq_num_codes,
                 out_dim=projection_dim,
             ).to(device)
-            encoder_params += list(self.projection.parameters())
-
             self.projection_target = Projection(
                 mlp_dims=[1024],
                 levels=fsq_levels,
@@ -551,6 +549,7 @@ class VQ_TC_TD3(Agent):
                 self.projection_target = torch.compile(
                     self.projection_target, mode="default"
                 )
+            encoder_params += list(self.projection.parameters())
 
         self.ae_opt = torch.optim.AdamW(encoder_params, lr=ae_learning_rate)
 

### 2024-01-28 paper/main.tex
started adding basic layout

@@
 
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=0.6\columnwidth]{./figs/diagram-self-predictive.png}}
+\caption{Self-predictice loss}
+\label{diagram-self-predictive}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
 \section{Related Work}
@@
 
-In humanoid-run and dog walk compare
+\todo{Add algorithm}
+
+\begin{align}
+z_{t} &= e_{\theta} (s_{t}) \\
+\hat{z}_{t+1} &= d_{\phi} (z_{t}, a_{t}) \\
+q_{t} &= Q_{\psi} (z_{t}, a_{t}) \\
+a_{t} &\sim \pi_{\eta} (z_{t}),
+\end{align}
+
+\textbf{Representation learning}
+\begin{align}
+  \mathcal{L}(\theta, \phi; \mathcal{D})
+ % &= \E_{s_{t} \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| \hat{z}_{t+1} - z_{t+1} \|_{2}^{2} \right] \\
+&= \E_{(s_{t}, a_{t}, s_{t+1}) \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
+\end{align}
+
+\textbf{Policy and value function learning}
+
+\section{Experiments}
+\label{sec:experiments}
+
 \begin{itemize}
-    \item Show that resetting is beneficial
-    \begin{itemize}
-        \item But picking when to reset varies across environments
-    \end{itemize}
-    \item DDPG
-    \item DDPG reset every X param updates
-    \item DDPG-AE
-    \item DDPG-AE reset every X param updates
-    \item DDPG-AE auto-resets
-    \item DDPG-TAE
-    \item DDPG-TAE reset every X param updates
-    \item DDPG-TAE auto-resets
+    \item Show normalised latent leads to higher sample efficiency. Show results for NTC-TD3/TC-TD3/TD3/SAC/TCRL/TD-MPC with UTD=1.
+    \item Show normalised latent works with higher UTD. Compare NTC-TD3/TC-TD3/TCRL.
+    \item Show task-agnostic representation is good for incremental task learning. Compare NTC-TD3/TCRL-ours/TCRL
 \end{itemize}
@@
 \begin{itemize}
-    \item Partial vs full resets vs $50\%$ resets etc
-    \item Interleaved training vs representation first
-    \begin{itemize}
-        \item Show that representation first is better when resetting
-        \item But interleaved better if not resetting
-    \end{itemize}
-    \item Normalized latent (SimNorm) vs unnormalized
-\end{itemize}
-
-\begin{itemize}
-    \item Our method for auto-resetting handles sensors breaking during agents lifetime.
-    \item Show this by transforming agents observations at some point during training.
-\end{itemize}
-
-\begin{itemize}
-    \item When to reset actor/critic/encoder?
-    \item How to pick UTD?
-    \item How to pick learning rate?
+  \item Show un-normalised isn't as good.
+  \item Target encoder vs online encoder for mapping states to latents for TD3.
+  \item Show cosine similarity slows training.
+  \item Show adding reconstruction loss hurts performance.
 \end{itemize}
@@
 \textbf{Representation learning improves sample efficiency in RL}
-\begin{theorem}
-\label{thm:bigtheorem}
-Given two MDPs $\left( \mathcal{S}_{1}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$ and $\left( \mathcal{S}_{2}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$
-with $|\mathcal{S}_{1}| < |\mathcal{S}_{2}|$,
-the optimal value function $V : \mathcal{S} \rightarrow \R$ in $\mathcal{S}_{1}$ can be learned with less environment samples than in $\mathcal{S}_{2}$.
-\end{theorem}
+% \begin{theorem}
+% \label{thm:bigtheorem}
+% Given two MDPs $\left( \mathcal{S}_{1}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$ and $\left( \mathcal{S}_{2}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$
+% with $|\mathcal{S}_{1}| < |\mathcal{S}_{2}|$,
+% the optimal value function $V : \mathcal{S} \rightarrow \R$ in $\mathcal{S}_{1}$ can be learned with less environment samples than in $\mathcal{S}_{2}$.
+% \end{theorem}
 
-\begin{proof}
-Left as an exercise to the reader.
-\end{proof}
-
-\begin{corollary}
-If $f:X\to Y$ is bijective,
-the cardinality of $X$ is at least as large as that of $Y$.
-\end{corollary}
-
-\begin{assumption}
-The state space $\mathcal{S}$ is discrete
-\label{ass:discrete-state-space}
-\end{assumption}
+% \begin{assumption}
+% The state space $\mathcal{S}$ is discrete
+% \label{ass:discrete-state-space}
+% \end{assumption}
 
@@
     \end{itemize}
-\end{itemize}
-
-\begin{itemize}
-    \item Assume latent space contains all information from state space needed to predict values
-    \item Show retraining better contains information from state space
-    \begin{itemize}
-        \item Train agent on $\mathcal{D}^{i}$ with old encoder $e_{\phi^{i-1}}$ and compare to new encoder $e_{\phi^{i}}$
-    \end{itemize}
-\end{itemize}
-
-\section{Experiments}
-\label{experiments}
-
-SAC-AE is just SAC but using AE to get latent.
-BDC use bisimulation metric so they can obtain bound on the suboptimality of a value
-function trained on their representation.
-
-\begin{itemize}
-    \item Compare to SAC-AE
-    \item Compare to TCRL
-    \item Compare to SLAC
-    \item Compare to SAC-BDC \citet{zhangLearningInvariantRepresentations2020}
-    \item Compare to DeepMDP
-    \item Compare to DHGP
-\end{itemize}
-
-\begin{itemize}
-    \item Show we can use higher UTD?
-    \item AE vs VAE vs VQ-VAE
-    \item Full reset vs partial reset
-    \begin{itemize}
-        \item Reset last layer in MuJoCo
-        \item Reset a percentage (\eg 50\%) on image observations
-    \end{itemize}
-    \item Compare to resetting every X parameter updates like in \citep{doroSampleEfficientReinforcementLearning2022}.
-    \item How much training should we do after parameter reset?
-    \item MuJoCo and Atari
 \end{itemize}

### 2024-01-29 agents/vq_tc_td3.py
put log rank in try/except

@@ -740,13 +740,16 @@ class VQ_TC_TD3(Agent):
 
                     # Log rank of latent
                     def log_rank(name, z):
-                        rank3 = matrix_rank(z, atol=1e-3, rtol=1e-3)
-                        rank2 = matrix_rank(z, atol=1e-2, rtol=1e-2)
-                        rank1 = matrix_rank(z, atol=1e-1, rtol=1e-1)
-                        condition = cond(z)
-                        for j, rank in enumerate([rank1, rank2, rank3]):
-                            wandb.log({f"{name}-rank-{j}": rank.item()})
-                        wandb.log({f"{name}-cond-num": condition.item()})
+                        try:
+                            rank3 = matrix_rank(z, atol=1e-3, rtol=1e-3)
+                            rank2 = matrix_rank(z, atol=1e-2, rtol=1e-2)
+                            rank1 = matrix_rank(z, atol=1e-1, rtol=1e-1)
+                            condition = cond(z)
+                            for j, rank in enumerate([rank1, rank2, rank3]):
+                                wandb.log({f"{name}-rank-{j}": rank.item()})
+                            wandb.log({f"{name}-cond-num": condition.item()})
+                        except:
+                            pass
 
                     z_batch = self.encoder(batch.observations[0])
                     if self.use_fsq:

### 2024-01-29 agents/vq_tc_td3.py
log less fequently

@@ -345,6 +345,7 @@ class VQ_TC_TD3(Agent):
         tau: float = 0.005,
         rho: float = 0.9,  # discount for dynamics
         act_with_target: bool = False,  # if True act with target actor network
+        logging_epoch_freq: int = 499,
         # Reset stuff
         reset_type: str = "last_layer",  # "full" or "last-layer"
         reset_strategy: str = "latent-dist",  #  "latent-dist" or "every-x-param-updates"
@@ -411,6 +412,8 @@ class VQ_TC_TD3(Agent):
         # self.ae_patience = int(ae_patience / self.early_stopper_freq)
         self.ae_patience = ae_patience
         self.ae_min_delta = ae_min_delta
+
+        self.logging_epoch_freq = logging_epoch_freq
 
         self.project_latent = project_latent
         self.projection_dim = projection_dim
@@ -727,7 +730,7 @@ class VQ_TC_TD3(Agent):
                 if wandb.run is not None:
                     wandb.log({"reset": reset_flag})
 
-            if i % 100 == 0:
+            if i % self.logging_epoch_freq == 0:
                 logger.info(
                     f"Iteration {i} | loss {info['encoder_loss']} | rec loss {info['rec_loss']} | tc loss {info['temporal_consitency_loss']} | reward loss {info['reward_loss']} | value dynamics loss {info['value_dynamics_loss']}"
                 )
@@ -800,7 +803,7 @@ class VQ_TC_TD3(Agent):
             batch = replay_buffer.sample(self.ae_batch_size, val=False)
             info.update(self.update_representation_step(batch=batch))
 
-            if i % 100 == 0:
+            if i % self.logging_epoch_freq == 0:
                 logger.info(
                     f"Iteration {i} | loss {info['encoder_loss']} | rec loss {info['rec_loss']} | tc loss {info['temporal_consitency_loss']} | reward loss {info['reward_loss']} | value dynamics loss {info['value_dynamics_loss']}"
                 )
@@ -924,7 +927,7 @@ class VQ_TC_TD3(Agent):
                         self.reset(replay_buffer=replay_buffer)
                         wandb.log({"reset": 1})
 
-            if i % 100 == 0:
+            if i % self.logging_epoch_freq == 0:
                 if wandb.run is not None:
                     wandb.log(info)
                     wandb.log({"reset": 0})

### 2024-01-29 agents/td3.py
use logging_freq

@@ -125,6 +125,7 @@ class TD3(Agent):
         discount: float = 0.99,
         tau: float = 0.005,
         act_with_target: bool = False,  # if True act with target network
+        logging_freq: int = 499,
         device: str = "cuda",
         name: str = "TD3",
         compile: bool = False,
@@ -152,6 +153,7 @@ class TD3(Agent):
         self.discount = discount
         self.tau = tau
         self.act_with_target = act_with_target
+        self.logging_freq = logging_freq
         self.device = device
 
         # Init actor and it's target
@@ -213,7 +215,7 @@ class TD3(Agent):
                 if wandb.run is not None:
                     wandb.log({"reset": reset_flag})
 
-            if i % 100 == 0:
+            if i % self.logging_freq == 0:
                 if wandb.run is not None:
                     wandb.log(info)
                     wandb.log({"reset": reset_flag})

### 2024-01-29 agents/vq_tc_td3.py

@@ -345,7 +345,7 @@ class VQ_TC_TD3(Agent):
         tau: float = 0.005,
         rho: float = 0.9,  # discount for dynamics
         act_with_target: bool = False,  # if True act with target actor network
-        logging_epoch_freq: int = 499,
+        logging_freq: int = 499,
         # Reset stuff
         reset_type: str = "last_layer",  # "full" or "last-layer"
         reset_strategy: str = "latent-dist",  #  "latent-dist" or "every-x-param-updates"
@@ -413,7 +413,7 @@ class VQ_TC_TD3(Agent):
         self.ae_patience = ae_patience
         self.ae_min_delta = ae_min_delta
 
-        self.logging_epoch_freq = logging_epoch_freq
+        self.logging_freq = logging_freq
 
         self.project_latent = project_latent
         self.projection_dim = projection_dim
@@ -730,7 +730,7 @@ class VQ_TC_TD3(Agent):
                 if wandb.run is not None:
                     wandb.log({"reset": reset_flag})
 
-            if i % self.logging_epoch_freq == 0:
+            if i % self.logging_freq == 0:
                 logger.info(
                     f"Iteration {i} | loss {info['encoder_loss']} | rec loss {info['rec_loss']} | tc loss {info['temporal_consitency_loss']} | reward loss {info['reward_loss']} | value dynamics loss {info['value_dynamics_loss']}"
                 )
@@ -803,7 +803,7 @@ class VQ_TC_TD3(Agent):
             batch = replay_buffer.sample(self.ae_batch_size, val=False)
             info.update(self.update_representation_step(batch=batch))
 
-            if i % self.logging_epoch_freq == 0:
+            if i % self.logging_freq == 0:
                 logger.info(
                     f"Iteration {i} | loss {info['encoder_loss']} | rec loss {info['rec_loss']} | tc loss {info['temporal_consitency_loss']} | reward loss {info['reward_loss']} | value dynamics loss {info['value_dynamics_loss']}"
                 )
@@ -927,7 +927,7 @@ class VQ_TC_TD3(Agent):
                         self.reset(replay_buffer=replay_buffer)
                         wandb.log({"reset": 1})
 
-            if i % self.logging_epoch_freq == 0:
+            if i % self.logging_freq == 0:
                 if wandb.run is not None:
                     wandb.log(info)
                     wandb.log({"reset": 0})

### 2024-01-29 paper/main.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\our}{\textsc{vq-td3}\xspace}
+
 % Latin
@@
 
+Simplical embeddings
+\cite{lavoieSimplicialEmbeddingsSelfSupervised2022}
+
 \section{Method}
@@
 
-\todo{Add algorithm}
+In this section, we detail our method, titled Vector Quantized TD3 (\our).
+\our is conceptually simple, it first learns a representation of the observation space and then,
+given this representation, it performs model-free RL (e.g. TD3) on this representation.
+% It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
+% to a latent space $z \in \mathcal{Z}$.
+% It first learns a representation of the observation space and then uses
+See \cref{alg:main_alg}.
+
+We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma,\mathcal{T})$, where an agent receives
+observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.
+The agent then obtains a reward $r_{t} = \mathcal{R} (o_{t}, a_{t})$ and obtains the next observation
+$o_{t+1} = P(\cdot \mid o_{t}, a_{t})$.
+discount factor $\gamma \in [0, 1]$
+
+\our has four parameterized components.
+\begin{algorithm}[tb]
+   \caption{\our}
+   \label{alg:main_alg}
+\begin{algorithmic}
+   \STATE {\bfseries Input:} Encoder $e_{\theta}$, transition $d_{\phi}$, critics $Q_{\psi}$, policy $\pi_{\eta}$
+   \FOR{$i$ {\bfseries to} $N_{\text{episodes}}$}
+    \STATE $\mathcal{D} \leftarrow \mathcal{D} \cup \{o_{t}, a_{t}, o_{t+1}, r_{t+1}\}^{T}_{t=0}$
+    \FOR{$i=1$ {\bfseries to} $T \times UTD$}
+        \STATE Representation update
+        \STATE TD3 update
+    \ENDFOR
+   \ENDFOR
+\end{algorithmic}
+\end{algorithm}
 
 \begin{align}
-z_{t} &= e_{\theta} (s_{t}) \\
-\hat{z}_{t+1} &= d_{\phi} (z_{t}, a_{t}) \\
-q_{t} &= Q_{\psi} (z_{t}, a_{t}) \\
-a_{t} &\sim \pi_{\eta} (z_{t}),
+\text{Encoder: } &z_{t} = e_{\theta} (s_{t}) \label{eq:encoder} \\
+\text{Dynamics: } &\hat{z}_{t+1} = d_{\phi} (z_{t}, a_{t}) \label{eq:transition} \\
+\text{Value: } &q_{t} = Q_{\psi} (z_{t}, a_{t}) \label{eq:value} \\
+\text{Policy: } &a_{t} \sim \pi_{\eta} (z_{t}) , \label{eq:policy}
+\end{align}
+Note that unlike TCRL \citep{zhaoSimplifiedTemporalConsistency2023}, our transition model does
+not predict the reward.
+
+\textbf{Representation learning}
+Our representation learning is based soley on the temporal consistency loss,
+However, unlike previous works, we normalize the latent vector
+Our representation learning consists of an encoder $e_{\theta}$ (which maps observations to
+latent states) and a transition model $d_{\phi}$ which maps
+Our representation learning adopts a self-supervised loss.
+
+\begin{align} \label{eq:rep-loss}
+  \mathcal{L}_{\text{rep}}(\theta, \phi; \mathcal{D})
+ % &= \E_{s_{t} \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| \hat{z}_{t+1} - z_{t+1} \|_{2}^{2} \right] \\
+%&= \E_{(s_{t}, a_{t}, s_{t+1}) \sim \mathcal{D}}
+&= \E_{\tau \sim \mathcal{D}}
+\left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
 \end{align}
 
-\textbf{Representation learning}
-\begin{align}
-  \mathcal{L}(\theta, \phi; \mathcal{D})
- % &= \E_{s_{t} \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| \hat{z}_{t+1} - z_{t+1} \|_{2}^{2} \right] \\
-&= \E_{(s_{t}, a_{t}, s_{t+1}) \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
-\end{align}
+\textbf{Finite Scalar Quantization}
+We follow Finite Scalar Quantization (FSQ, \cite{mentzerFiniteScalarQuantization2023})
+and bound each dimension of our latent space.
+Given the output of our MLP encoder $z \in \R^{d}$,
+
+where each entry in  $f : z \rightarrow \left[ L/2 \right] \mathrm{tanh}(z))$
+
+% Thereby, we have zˆ ∈ C, where C is the implied codebook, given by the product of these per-channel
+% codebook sets, with |C| = L
+% d
+% . The vectors in C can simply be enumerated leading to a bijection from
+% any zˆ to an integer in {1, .
+.
+.
+, Ld}.
+Therefore, VQ can be replaced with FSQ in any neural networkrelated setup where VQ is commonly used, e.g., to train transformers, after appropriately adapting
+% the output and input dimension of the layers before and after VQ, respectively. We generalize the
+% above exposition to the case where the i-th channel is mapped to Li values and get |C| =
+% Qd
+% i=1 Li
 
@@
 
+\begin{align} \label{eq:value-loss}
+  \mathcal{L}_{Q}(\psi ; \mathcal{D}) &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad \forall k \in 1, 2 \\
+  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n})
+\end{align}
+
+\begin{align} \label{eq:policy-loss}
+ \mathcal{L}_{\pi}(\eta; \mathcal{D}) = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
+\end{align}
+
 \section{Experiments}
 \label{sec:experiments}
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/main_plot.pdf}}
+\caption{Comparison to baselines with UTD=1.}
+\label{fig:main_plot}
+\end{center}
+\vskip -0.2in
+\end{figure*}
 
@@
   \item Show adding reconstruction loss hurts performance.
-\end{itemize}
-
-Simplical embeddings
-\cite{lavoieSimplicialEmbeddingsSelfSupervised2022}
-
-\section{Theoretical Analysis}
-% \label{}
-
-Effect of domain shift on policy due to encoder being learned online
-
-\citet{zhangLearningInvariantRepresentations2020} have proofs for bounds on value functions with encoder.
-We could probably start with some of these.
-They build off the idea of bisimulation \citep{larsenBisimulationProbabilisticTesting1989}.
-Where encoders group states into clusters whilst preserving properties
-(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
-
-\textbf{Representation learning improves sample efficiency in RL}
-% \begin{theorem}
-% \label{thm:bigtheorem}
-% Given two MDPs $\left( \mathcal{S}_{1}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$ and $\left( \mathcal{S}_{2}, \mathcal{A}, P, r, \rho_{0}, \gamma \right)$
-% with $|\mathcal{S}_{1}| < |\mathcal{S}_{2}|$,
-% the optimal value function $V : \mathcal{S} \rightarrow \R$ in $\mathcal{S}_{1}$ can be learned with less environment samples than in $\mathcal{S}_{2}$.
-% \end{theorem}
-
-% \begin{assumption}
-% The state space $\mathcal{S}$ is discrete
-% \label{ass:discrete-state-space}
-% \end{assumption}
-
-\begin{itemize}
-    \item Prove latent space improves sample efficiency
-    \begin{itemize}
-        \item Assume discrete state space
-        \item Assume latent space contains all information from state space needed to predict values
-        \item Show smaller state space can be learned with less samples
-    \end{itemize}
-    \item Prove retraining from scratch is better than continuing training
-    \begin{itemize}
-        \item Training encoder episode to episode is like non-IID training
-    \end{itemize}
 \end{itemize}
@@
 
-You can have as much text here as you want. The main body must be at most $8$ pages long.
-For the final version, one more page can be added.
-If you want, you can use an appendix like this one.
-
-The $\mathtt{\backslash onecolumn}$ command above can be kept in place if you prefer a one-column appendix, or can be removed if you prefer a two-column appendix.
-Apart from this possible change, the style (font size, spacing, margins, page numbering, etc.) should be kept the same as the main body.
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=0.97\textwidth]{./figs/all_mujoco_envs.pdf}}
+\caption{Comparison to SAC/TCRL with UTD=1.}
+\label{fig:main_plot}
+\end{center}
+\vskip -0.2in
+\end{figure*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-01-29 paper/main.tex
update fig 1 and added some intro

@@
 
+Reinforcement learning (RL) has potential...
+
+However, applying RL in real-world environments with high dimensional/complex observation spaces is challenging.
+
+Compressing observations in latent state space.
+Hand crafted features
+
+Learning features with representation learning.
+These methods typically learn task-specific representations
+as they train the representation to predict rewards and/or values
+\citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
+
+For example, \citet{zhaoSimplifiedTemporalConsistency2023} use a latent-state consistency loss based upon the Cosine
+similarity loss plus the MSE of reward predictions in the latent space.
+The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
+
+In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method.
+Normalization based upon Finite Scalar Quantization
+Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} prevents representation collapse,
+and {\em (iii)} is task-agnostic.
+
+% In recent years, it has become common to
+
 \begin{figure}[ht]
@@
 \begin{center}
-\centerline{\includegraphics[width=0.6\columnwidth]{./figs/diagram-self-predictive.png}}
+\centerline{\includegraphics[width=\columnwidth]{./figs/method-diagram.png}}
 \caption{Self-predictice loss}
@@
 is where an agent overfits on early data and loses its ability to learn from new data,
-which arises due to the plasticity in NNs \cite{lyleUnderstandingPlasticityNeural2023} \todo{earlier citation of plasticiy in NNs??}.
-\cite{nikishinPrimacyBiasDeep2022} showed that periodically resetting the agent's parameters can help alleviate primacy bias.
-\cite{doroSampleEfficientReinforcementLearning2022} built on top of this insight and showed that resetting the
+which arises due to the plasticity in NNs \citep{lyleUnderstandingPlasticityNeural2023} \todo{earlier citation of plasticiy in NNs??}.
+\citet{nikishinPrimacyBiasDeep2022} showed that periodically resetting the agent's parameters can help alleviate primacy bias.
+\citet{doroSampleEfficientReinforcementLearning2022} built on top of this insight and showed that resetting the
 agent's parameters not only alleviates primacy bias but also
@@
 
+\begin{table}[t]
+\caption{FSQ levels $\mathcal{L}$ to approximate different codebook sizes $|\mathcal{C}|$}
+\label{sample-table}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lccc}
+\toprule
+Target size $|\mathcal{C}|$ & $2^{8}$ & $2^{10}$ & $2^{12}$ \\
+\midrule
+Proposed $\mathcal{L}$ & $[8,6,5]$ & $[8,5,5,5]$ & $[7,5,5,5,5]$ \\
+% \bottomrule
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+
 \begin{figure*}[ht]
@@
   \item Show un-normalised isn't as good.
+  \item Compare different size latent spaces? Different levels $[8,5]$ vs $[8,6,5]$ vs $[8,5,5,5]$ vs $[7,5,5,5,5]$
   \item Target encoder vs online encoder for mapping states to latents for TD3.

### 2024-01-29 paper/main.tex
aligned main eq

@@
 \begin{align}
-\text{Encoder: } &z_{t} = e_{\theta} (s_{t}) \label{eq:encoder} \\
-\text{Dynamics: } &\hat{z}_{t+1} = d_{\phi} (z_{t}, a_{t}) \label{eq:transition} \\
-\text{Value: } &q_{t} = Q_{\psi} (z_{t}, a_{t}) \label{eq:value} \\
-\text{Policy: } &a_{t} \sim \pi_{\eta} (z_{t}) , \label{eq:policy}
+&\text{Encoder: } & z_{t} &= e_{\theta} (s_{t}) \label{eq:encoder} \\
+&\text{Dynamics: } & \hat{z}_{t+1} &= d_{\phi} (z_{t}, a_{t}) \label{eq:transition} \\
+&\text{Value: } & q_{t} &= Q_{\psi} (z_{t}, a_{t}) \label{eq:value} \\
+&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) , \label{eq:policy}
 \end{align}

### 2024-01-29 paper/main.tex
Sketch Fig. 1

@@
 % if you use cleveref..
-\usepackage[capitalize,noabbrev]{cleveref}
+\usepackage[capitalize,noabbrev,nameinlink]{cleveref}
 
@@
 % Vectors
+\usepackage{bm}
 \def\vzero{{\bm{0}}}
@@
 
+\usetikzlibrary{shapes.geometric,positioning,calc}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=\columnwidth]{./figs/method-diagram.png}}
+\begin{figure}[t!]
+\centering\scriptsize
+\resizebox{\columnwidth}{!}{%
+\begin{tikzpicture}
+
+  % Declare that we have a background layer (advanced)
+  \pgfdeclarelayer{background}
+  \pgfsetlayers{background,main}
+
+  % Helpers for small subscripts
+  \newcommand{\sub}[0]{\scalebox{.8}{\ensuremath t}}
+  \newcommand{\subs}[1]{\scalebox{.8}{\ensuremath t{+}#1}}
+
+  % Lengths (this tunes distance)
+  \newlength{\nodedist}\setlength{\nodedist}{2cm}
+
+  % Styles (this tunes colours etc.)
+  \tikzstyle{mynode}=[fill=blue!10!cyan,rounded corners=1pt,font=\tiny,inner sep=0]
+  \tikzstyle{trap}=[mynode,trapezium,minimum width=1.5em, minimum height=2em,
+                    trapezium left angle=-70, trapezium right angle=-70]
+  \tikzstyle{blob}=[mynode,circle,minimum width=2.4em,minimum height=2.4em]
+  \tikzstyle{polval}=[mynode,minimum width=6em,minimum height=2em,align=center,text=white,fill=black!60,inner sep=2pt]
+  \tikzstyle{arr}=[line width=1pt,black,->]
+  \tikzstyle{darr}=[line width=1pt,black,<->,densely dotted]
+
+  % *** FIRST HALF ***
+
+  % Draw nodes
+  \node[blob] (z0) at (0\nodedist,0) {$\vz_{\sub}$};
+  \node[blob] (z1) at (1\nodedist,0) {$\hat\vz_{\subs{1}}$};
+  \node[blob] (z2) at (2\nodedist,0) {$\hat\vz_{\subs{2}}$};
+
+  \node[trap] (e0) at (0\nodedist,\nodedist) {$e_\phi$};
+  \node[trap] (e1) at (1\nodedist,\nodedist) {$e_{\bar\phi}$};
+  \node[trap] (e2) at (2\nodedist,\nodedist) {$e_{\bar\phi}$};
+
+  \node[blob] (zh1) at ($(z1)!0.5!(e1)$) {$\bar\vz_{\subs{1}}$};
+  \node[blob] (zh2) at ($(z2)!0.5!(e2)$) {$\bar\vz_{\subs{2}}$};
+
+  \coordinate (m0) at ($(z0)!.5!(z1)$);
+  \node[blob] (a0) at ($(m0)!(zh1)!(m0)$) {$\va_{\sub}$};
+  \coordinate (m1) at ($(z1)!.5!(z2)$);
+  \node[blob] (a1) at ($(m1)!(zh2)!(m1)$) {$\va_{\subs{1}}$};
+
+  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/bg}};
+  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/bg}};
+  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/bg}};
+
+  \node[anchor=north] (o0) at (i0.north) {\color{white}$o_{\sub}$};
+  \node[anchor=north] (o1) at (i1.north) {\color{white}$o_{\subs{1}}$};
+  \node[anchor=north] (o2) at (i2.north) {\color{white}$o_{\subs{2}}$};
+
+  \node[above of=i1,align=center] (lab0) {\normalsize\bf 1.\ Learn representation};
+
+  % Draw arrows
+  \draw[arr] (i0) -- (e0);
+  \draw[arr] (i1) -- (e1);
+  \draw[arr] (i2) -- (e2);
+  \draw[arr] (z0) -- (z1);
+  \draw[arr] (z1) -- (z2);
+  \draw[arr] (e0) -- (z0);
+  \draw[arr] (e1) -- (zh1);
+  \draw[arr] (e2) -- (zh2);
+  \draw[arr] (a0) -- (z1);
+  \draw[arr] (a1) -- (z2);
+  \draw[darr] (zh1) -- (z1);
+  \draw[darr] (zh2) -- (z2);
+
+  % *** SECOND HALF ***
+
+  \node[blob] (z) at ($(e2) + (1.5\nodedist,0)$) {$\bar\vz_{\sub}$};
+  \node[polval] (policy) at ($(z) + (-.5\nodedist,-.5\nodedist)$) {{\bf Policy} \\ $\pi_\theta(\va_t \mid \bar\vz_t)$};
+  \node[polval] (value) at ($(z) + (.5\nodedist,-.5\nodedist)$) {{\bf Value} \\ $\mathcal{Q}_\psi(\bar\vz_t, \va_t)$};
+
+  \draw[arr] (z) -- (policy);
+  \draw[arr] (z) -- (value);
+
+  \node[align=center] at ($(z)!(lab0)!(z)$) {\normalsize\bf 2.\ Latent actor-critic\vphantom{p}};
+
+  % Draw backgrounds
+  \begin{pgfonlayer}{background}
+    \draw[draw=none,fill=black!10,rounded corners=5pt] ($(i0.north west) + (-.5cm,1cm)$) 
+      rectangle ($(i2.east)!(z2.north)!(i2.east) + (.5cm,-1cm)$);
+  \end{pgfonlayer}
+
+\end{tikzpicture}}
+%\centerline{\includegraphics[width=\columnwidth]{./figs/method-diagram.png}}
 \caption{Self-predictice loss}
 \label{diagram-self-predictive}
-\end{center}
 \vskip -0.2in

### 2024-01-31 paper/main.tex

@@
   \begin{pgfonlayer}{background}
-    \draw[draw=none,fill=black!10,rounded corners=5pt] ($(i0.north west) + (-.5cm,1cm)$) 
-      rectangle ($(i2.east)!(z2.north)!(i2.east) + (.5cm,-1cm)$);
+
+    \coordinate (mid) at ($(i2.east)!.5!(policy.west)$);
+
+    \draw[draw=none,fill=black!10,rounded corners=5pt] ($(i0.north west) + (-.25cm,1cm)$) 
+      rectangle ($(mid)!(z2.north)!(mid) + (-3pt,-1cm)$);
+
+    \draw[draw=none,fill=black!05,rounded corners=5pt] ($(mid)!(i2.north east)!(mid) + (3pt,1cm)$) 
+      rectangle ($(value.east)!(z2.north)!(value.east) + (.25cm,-1cm)$);
+
   \end{pgfonlayer}

### 2024-01-31 paper/figs/plotting/main_plot/fetch_ddpg_data.py
initial commit of some plotting stuff

@@ -1,0 +1,136 @@
+# %%
+ddpg_path = "yizhao/sprl/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        "acrobot-swingup": [
+            "1eiys0vf",
+            "1yedk0zx",
+            "1pbg10k5",
+            "zgb6ndd1",
+            "2dh8jyh9",
+        ],
+        "cheetah-run": [
+            "i6sm1a7b",
+            "dqax4n78",
+            "1p4lrio9",
+            "1gfzkhtr",
+            "2t0dn97k",
+        ],
+        "fish-swim": [
+            "1zrla7u4",
+            "3rx4mc18",
+            "xwq0fu1r",
+            "smv9miz4",
+            "nu50jdp0",
+        ],
+        "quadruped-walk": [
+            "du473356",
+            "3kyhmy9n",
+            "1yusrrco",
+            "3198sjan",
+            "3fdn4ijo",
+        ],
+        "walker-walk": [
+            "38b1tu3q",
+            "3qg5ghou",
+            "2a8wgbv4",
+            "kjw1t2hi",
+            "c6m4ztsu",
+        ],
+        "humanoid-walk": [
+            "2xa4zh7i",
+            "u01r5cd8",
+            "2dm45ezi",
+            "2ywqbphi",
+            "33kix5gc",
+        ],
+        "dog-walk": [
+            "1j5y9f66",
+            "10k9mjuf",
+            "er1d5h4k",
+            "t6b44j79",
+            "3500wzl6",
+        ],
+        "dog-run": [
+            "2e43rg7f",
+            "1ev56klj",
+            "1rtay2kk",
+            "1fjb7wss",
+            "2p7hov2d",
+        ],
+    },
+}
+
+# %%
+# Plot wandb
+import numpy as np
+from typing import Dict
+
+import matplotlib.pyplot as plt
+
+# plt.style.use('seaborn')
+plt.style.use("seaborn-v0_8")
+
+import wandb
+
+api = wandb.Api(timeout=19)
+import pandas as pd
+
+# pet-pytorch
+titles = [
+    "env_step",
+    "episode",
+    "episode_reward",
+    "time",
+]
+keys = [
+    "eval/.env_step",
+    "eval/.episode",
+    "eval/.episode_reward",
+    "eval/.eval_total_time",
+]
+
+rename = {a: b for a, b in zip(keys, titles)}
+
+
[…]

### 2024-01-31 paper/figs/plotting/main_plot/fetch_redq_utd10_data.py

@@ -1,0 +1,65 @@
+#%%
+redq_path = 'yizhao/redq_aligned/'
+redq_utd10_data = {
+    'path': redq_path,
+    'data': {
+        'acrobot-swingup': ['2avs8qvo','wy9mr8ke','2tcw6uat','xk9hhggf','240610cf'],
+        'cheetah-run': ['3bykfzb8','3t5b2qbg','2xwmvfx1','1bhfo6ql','1sp0l2mc'],
+        'fish-swim': ['2crog1zw','21cobt86','1omunkwt','2mfay1qd','2mym8cu9'],
+        'quadruped-walk': ['2gmklz3o','2fa7fmvq','2lowi443','2mpy5njk','1saqkms5'],
+        'walker-walk': ['19p8pd8a','2v3j1ks2','28drab7y','3r7xgjq5','1h2uhoeg'],
+        'humanoid-walk': ['26x8gnsw','2lif0urn','2nt23llu','1f150tks'],
+        'dog-walk': ['2d1g8awy','2jqf95ei','2dnna3pd','2gxvg107','2e7wc6u3', '2ef2ejbb'],
+        'dog-run': ['l0jxplhr','31sbxbbt','3vsqdij1','17185yhl','10xfrxc7', '2ipeutzk'],
+    }
+}
+
+# %%
+# Plot wandb
+import numpy as np
+from typing import Dict
+
+import matplotlib.pyplot as plt
+plt.style.use('seaborn')
+
+import wandb
+api = wandb.Api(timeout=19)
+import pandas as pd
+
+# pet-pytorch
+titles = ['env_step','episode','episode_reward', 'time',]
+keys = ['eval/.env_step', 'eval/.episode', 'eval/.episode_reward', 'eval/.total_time']
+
+rename = {a : b for a, b in zip(keys, titles)}
+
+def fetch_results(run_path, run_name_list, keys, agent_name):
+
+    data = []
+    for run_name in run_name_list:
+        wandb_run = api.run(run_path + run_name)
+        history = wandb_run.history(keys=keys, pandas=True)
+        history = history.rename(columns=rename)
+        # obtain seed, and env_name
+        seed, env_name = wandb_run.config['_content']['seed'], wandb_run.config['_content']['env']
+
+        # append env_name, seed, and agent_name
+        history['env'] = env_name.replace('_', '-', 1)
+        history['seed'] = seed
+        history['agent'] = agent_name
+
+        data.append(history)
+    return pd.concat(data)
+
+# %%
[…]

### 2024-01-31 paper/figs/plotting/main_plot/fetch_vq_td3_data.py

@@ -1,0 +1,105 @@
+# %%
+ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        "humanoid-run": ["jv7cfy2v"],
+        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
+        "cartpole-swingup": ["w6eykaow"],  # UTD=1
+        "hopper-stand": ["3ozramwe"],  # UTD=1
+        "quadruped-walk": ["8q59zkws"],  # UTD=1
+        # "cheetah-run": ["mkmvf5md"],  # UTD=4
+        "cheetah-run": ["jc3gjwdg"],  # UTD=1
+        "walker-walk": ["26lcg142"],  # UTD=1
+        # "walker-walk": ["nrak2mn1"],  # UTD=4
+        "quadruped-run": ["478dl8sb"],  # UTD=1
+        # 'humanoid-walk': ['2xa4zh7i', 'u01r5cd8', '2dm45ezi', '2ywqbphi', '33kix5gc', ],
+        # "dog-walk": ["gtrhfxn3"],
+        "dog-run": ["ukzsy2t5"],
+        "dog-walk": ["sjqd8i9u"],
+        "dog-run": ["zhsh5oyd"],
+    },
+}
+
+# %%
+# Plot wandb
+import numpy as np
+from typing import Dict
+
+import matplotlib.pyplot as plt
+
+# plt.style.use("seaborn")
+plt.style.use("seaborn-v0_8")
+
+import wandb
+
+api = wandb.Api(timeout=19)
+import pandas as pd
+
+# pet-pytorch
+titles = [
+    "env_step",
+    # "episode",
+    "episode_reward",
+    # "time",
+]
+keys = [
+    "eval/.env_step",
+    # "eval/.episode",
+    "eval/.episodic_return",
+    # "eval/.eval_total_time",
+]
+
+rename = {a: b for a, b in zip(keys, titles)}
+
+
+def fetch_results(run_path, run_name_list, keys, agent_name):
+    data = []
[…]

### 2024-01-31 paper/figs/plotting/main_plot/plot_main.py

@@ -1,0 +1,120 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+COLORS = {
+    "TCRL": "#e41a1c",
+    "SAC": "#377eb8",
+    "REDQ": "#984ea3",
+    "TD-MPC": "#ff7f00",
+    "VQ-TD3": "magenta",
+}
+# %%
+main_envs = [
+    "acrobot-swingup",
+    "cheetah-run",
+    # "fish-swim",
+    "quadruped-walk",
+    "humanoid-run",
+    "walker-walk",
+    "humanoid-walk",
+    "dog-walk",
+    "dog-run",
+]
+
+
+def plot(df, key="episode_reward"):
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    nrow = len(main_envs) // ncol
+    # nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    for idx, env in enumerate(main_envs):
+        data = df[df["env"] == env]
+        row = idx // ncol
+        col = idx % ncol
+        ax = axs[row, col]
+        hue_order = data.agent.unique()
+
+        if idx == 4:
+            sns.lineplot(
+                x="env_step",
+                y=key,
+                data=data,
+                errorbar=("ci", 95),
+                hue="agent",
+                hue_order=hue_order,
+                palette=COLORS,
+                legend="auto",
+                ax=ax,
+            )
+            ax.legend().set_title(None)
+        else:
+            # breakpoint()
+            sns.lineplot(
+                x="env_step",
+                y=key,
+                data=data,
+                errorbar=("ci", 95),
+                hue="agent",
+                hue_order=hue_order,
+                palette=COLORS,
+                legend=False,
+                ax=ax,
[…]

### 2024-01-31 paper/figs/plotting/main_plot/plot_tcrl_alm_td3.py

@@ -1,0 +1,64 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+# plt.style.use('bmh')
+plt.style.use('seaborn-v0_8-whitegrid')
+import seaborn as sns
+
+plt.rcParams['figure.dpi'] = 400
+plt.rcParams['font.size'] = 13
+plt.rcParams['legend.fontsize'] = 12
+plt.rcParams['legend.loc'] = 'lower right'
+COLORS = ['#e41a1c', '#377eb8', '#984ea3', '#ff7f00', '#4daf4a',]
+# %%
+main_envs = ['acrobot-swingup', 'cheetah-run', 'fish-swim', 'quadruped-walk', 'walker-walk', 'humanoid-walk', 'dog-walk', 'dog-run']
+def plot(df, key='episode_reward'):
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    for idx, env in enumerate(main_envs):
+        data = df[df['env'] == env]
+        row = idx // ncol 
+        col = idx % ncol
+        ax = axs[row, col]
+        hue_order = data.agent.unique()
+
+        if idx == 4:
+            sns.lineplot(x='episode', y=key, data=data, errorbar=('ci', 95), hue='agent',hue_order=hue_order, 
+                    palette=COLORS[:4],
+                    legend='auto', ax=ax)
+            ax.legend().set_title(None)
+        else:
+            sns.lineplot(x='episode', y=key, data=data, errorbar=('ci', 95), hue='agent', hue_order=hue_order, 
+                    palette=COLORS[:4],
+                    legend=False, ax=ax)
+          
+        ax.set_title(" ".join([ele.capitalize() for ele in env.split('-')]))
+        ax.set_xlabel('Environment Steps (1e3)')
+        ax.set_ylabel('Episode Return')
+    plt.tight_layout()
+    plt.savefig(f'alm_policy.pdf')
+    plt.show()
+    
+# %%
+# data_path = './'
+# data_list = ['tcrl', 'alm', 'sac']
+
[…]

### 2024-01-31 paper/figs/plotting/plot_all/correct_data.py

@@ -1,0 +1,15 @@
+#%%
+import pandas as pd
+import numpy as np
+# %%
+df1 = pd.read_csv('./tcrl.csv')
+# %%
+# fetch quadruped_run data
+#%%
+tcrl_path = 'yizhao/sprl/'
+tcrl_data = {
+    'path': tcrl_path,
+    'data': {
+        'quadruped-run': ['sfwbbi87', '3bs0g2ut', '6ck2fy07', '2hmiqbct', '3gsnqcr9',],
+     }
+}

### 2024-01-31 paper/figs/plotting/plot_all/plot_all.py

@@ -1,0 +1,91 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+
+# %%
+# first process SAC data
+COLORS = {
+    "TCRL": "#e41a1c",
+    "SAC": "#377eb8",
+    "SAC-our": "#984ea3",
+    "VQ-TD3": "magenta",
+}
+
+
+def plot(df, key="episode_reward"):
+    envs = np.sort(env_list)
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    for idx, env in enumerate(envs):
+        data = df[df["env"] == env]
+        data = data[data["episode"] <= max_ep[env]]
+        row = idx // ncol
+        col = idx % ncol
+        ax = axs[row, col]
+        hue_order = data.agent.unique()
+
+        if idx == 0:
+            sns.lineplot(
+                # x="episode",
+                x="env_step",
+                y=key,
+                data=data,
+                errorbar=("ci", 95),
+                hue="agent",
+                hue_order=hue_order,
+                palette=COLORS,
+                legend="auto",
+                ax=ax,
+            )
+            ax.legend().set_title(None)
+        else:
+            sns.lineplot(
+                x="env_step",
+                # x="episode",
+                y=key,
+                data=data,
+                errorbar=("ci", 95),
+                hue="agent",
+                hue_order=hue_order,
+                palette=COLORS,
+                legend=False,
+                ax=ax,
+            )
+
+        ax.set_title(" ".join([ele.capitalize() for ele in env.split("-")]))
+        ax.set_xlabel("Environment Steps (1e3)")
+        ax.set_ylabel("Episode Return")
+    plt.tight_layout()
+    plt.savefig(f"all_policy_new.pdf")
+    # plt.show()
+
+
+# %%
+data_path = "./"
[…]

### 2024-01-31 paper/figs/plotting/utd_plot/fetch_vq_td3_data.py

@@ -1,0 +1,208 @@
+# %%
+ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        "humanoid-run": ["jv7cfy2v"],
+        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
+        "cartpole-swingup": ["w6eykaow"],  # UTD=1
+        "hopper-stand": ["3ozramwe"],  # UTD=1
+        "quadruped-walk": ["8q59zkws"],  # UTD=1
+        # "cheetah-run": ["mkmvf5md"],  # UTD=4
+        "cheetah-run": [
+            # UTD=1
+            "u9plbcw6",  # seed=64238
+            "qh0aoiu8",  # seed=234
+            "nsbheioo",  # seed=64823
+            "ayf1ov5h",  # seed=7492
+            "72d02rea",  # seed=423
+            # UTD=2
+            "xm9mx9vw",  # seed=64238
+            "fy1d9lzj",  # seed=234
+            "09g92cv8",  # seed=64823
+            "9r14qhgy",  # seed=7492
+            "aalwd7ap",  # seed=423
+            # UTD=4
+            "te46dylh",  # seed=64238
+            "rx1rg4fw",  # seed=234
+            "p94l3jen",  # seed=64823
+            "c5tmc619",  # seed=7492
+            "vgawl7me",  # seed=423
+            # UTD=8
+            "b5cawxro",  # seed=64238
+            "cbhkp2og",  # seed=234
+            "w7dqbbdz",  # seed=64823
+            "pgr6kmbv",  # seed=7492
+            "gcx4iet5",  # seed=423
+        ],
+        "walker-walk": [  # UTD=1
+            "ygmyl5ok",  # seed=64238
+            "mk1dau1k",  # seed=234
+            "iybejvai",  # seed=64823
+            "2ncb40mo",  # seed=7492
+            "trqix916",  # seed=423
+            # UTD=2
+            "o9p2rtps",  # seed=64238
[…]

### 2024-01-31 paper/figs/plotting/utd_plot/make_utd_plot.py

@@ -1,0 +1,119 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+COLORS = {
+    "TCRL": "#e41a1c",
+    "SAC": "#377eb8",
+    "REDQ": "#984ea3",
+    "TD-MPC": "#ff7f00",
+    "VQ-TD3": "magenta",
+}
+# %%
+main_envs = [
+    # "acrobot-swingup",
+    "cheetah-run",
+    "walker-walk",
+    "walker-run",
+    # "fish-swim",
+    # "quadruped-walk",
+    "humanoid-run",
+    # "humanoid-walk",
+    # "dog-walk",
+    # "dog-run",
+]
+
+
+def plot(df, key="episode_reward"):
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    # nrow = len(main_envs) // ncol
+    nrow = 1
+    # nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    for idx, env in enumerate(main_envs):
+        data = df[df["env"] == env]
+        row = idx // ncol
+        col = idx % ncol
+        # ax = axs[row, col]
+        ax = axs[col]
+        hue_order = data.agent.unique()
+        # hue_order = data.utd_ratio.unique()
+        # breakpoint()
+
+        if idx == 0:
+            sns.lineplot(
+                x="env_step",
+                # x="episode",
+                y=key,
+                data=data,
+                errorbar=("ci", 95),
+                hue="agent",
+                style="utd_ratio",
+                hue_order=hue_order,
+                # palette=COLORS,
+                legend="auto",
+                ax=ax,
+            )
+            ax.legend().set_title(None)
+        else:
+            # breakpoint()
+            sns.lineplot(
+                x="env_step",
+                # x="episode",
+                y=key,
+                data=data,
[…]

### 2024-01-31 paper/main.tex
updated title

--- 2024-01-29 paper/main.tex
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Sample-efficient Reinforcement Learning with Latent Representations}
+\icmltitlerunning{Sample-efficient Reinforcement Learning with Implicitly Quantized Representations}
 
@@
 
-\newcommand{\our}{\textsc{vq-td3}\xspace}
+% \newcommand{\our}{\textsc{vq-td3}\xspace}
+\newcommand{\our}{\textsc{iFSQ-RL}\xspace}
 
@@
 %            International Conference on Machine Learning (ICML 2024)}
-\icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
+% \icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
+\icmltitle{Sample-efficient Reinforcement Learning with Implicitly Quantized Representations}
 
@@
 
-\textbf{Policy and value function learning}
-
-\begin{align} \label{eq:value-loss}
-  \mathcal{L}_{Q}(\psi ; \mathcal{D}) &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad \forall k \in 1, 2 \\
-  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n})
-\end{align}
-
-\begin{align} \label{eq:policy-loss}
- \mathcal{L}_{\pi}(\eta; \mathcal{D}) = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
-\end{align}
-
-\section{Experiments}
-\label{sec:experiments}
-
 \begin{table}[t]
@@
 
+\textbf{Policy and value function learning}
+
+\begin{align} \label{eq:value-loss}
+  \mathcal{L}_{Q}(\psi ; \mathcal{D}) &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad \forall k \in 1, 2 \\
+  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n})
+\end{align}
+
+\begin{align} \label{eq:policy-loss}
+ \mathcal{L}_{\pi}(\eta; \mathcal{D}) = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
+\end{align}
+
+\section{Experiments}
+\label{sec:experiments}
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/utd_comparison.pdf}}
+\caption{\textbf{Normalizing the latent-space enables higher UTD} \our's (blue) sample efficiency improves when the UTD ratio is increased from $1 \rightarrow 8$.
+TCRL (which has an un-normalized latent space) does not see an improvement in sample efficiency when when the UTD ratio is increased.
+Results are for four DMC tasks.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:normalization_enables_higher_utd}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\textbf{Normalizing the latent-space enables higher UTD}
+We first show how using bounded latent states enables us to increase the UTD ratio to see improvements in sample efficiency.
+\cref{fig:normalization_enables_higher_utd} shows that as the UTD ratio is increased from $1$ to $8$ \our's sample efficiency
+continues to improve.
+In contrast, when increasing the UTD ratio for TCRL no improvements are observed.
+
 \begin{figure*}[ht]
@@
 \centerline{\includegraphics[width=1.0\textwidth]{./figs/main_plot.pdf}}
-\caption{Comparison to baselines with UTD=1.}
-\label{fig:main_plot}
+\caption{\textbf{Normalizing the latent-space improves sample efficiency} \our is significanly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
+It performs similarly to TD-MPC (orange) whilst being significantly simpler and quicker to train.
+\our even outperforms TCRL, which is the most similar baseline.
+Results are for eight DMC tasks with UTD=1.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:normalization_improves_sample_efficiency}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\textbf{Normalizing the latent-space improves sample efficiency}
+In \cref{fig:normalization_improves_sample_efficiency},
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/utd_comparison.pdf}}
+\caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
+\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:multi-task-pretraining}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\textbf{\our's representation is helpful for multi-task RL}
+In contrast to TCRL and TD-MPC, \our learns a representation that is task-agnostic.
+As such, \our's task-agnostic representations can be used to speed-up learning in the incremental multi-task setting.
+That is, leverage information learned from solving one task to aid learning another task in the same environment.
+For example, pre-training our representation in walker-walk and then using this to speed up training in walker-walk.
+In \cref{fig:multi-task-pretraining} we show that \our benefits from pre-training the representation in walker-walk
+before training on walker-run.
+In contrast, TCRL does not see this improvement.
+This is because its representation loss contains a reward prediction term which makes its representation task-specific.
+These results demonstrate that confirms task-agnostic representation
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/fsq_levels_comparison.pdf}}
+\caption{\textbf{Comparison of different FSQ levels $\mathcal{L}$}.
+Results are for four DMC tasks.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:fsq_levels_comparison}
 \end{center}
@@
 \onecolumn
-\section{You \emph{can} have an appendix here.}
 
+\section{Hyperparameters}
+\cref{tab:hyperparameters} lists all of the hyperparameters for training \our.
+For further details please check the code which is available at XXXXXXX.
+
+\begin{table}[t]
+\caption{Hyperparameters for \our.}
+\label{tab:hyperparameters}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lll}
+\toprule
+Hyperparameter & Value & Description \\
+\midrule
+\textbf{Training} & & \\
+\hline
+Max episode length & 1000 &  Max episode length \\
+Num episodes & $500$ (easy) & Number of training episodes \\
+             & $1000$ (medium) & \\
+             & $15000$ (hard) & \\
+Random episodes & $10$ & Number of random episodes at start \\
+Action repeat & 2 & \\
+Eval every episodes & $5$ & \\
+Num eval episodes & $10$ & \\
+\hline
+\textbf{TD3} & & \\
+\hline
+Actor update freq & $2$  & update actor less frequently than critic \\
+Batch size & $256$ & \\
+Buffer size & $1e6$ & \\
+Exploration noise & $\mathrm{Linear}(1.0,0.1,50)$ (easy) & \\
+                  & $\mathrm{Linear}(1.0,0.1,150)$ (medium) & \\
+                  & $\mathrm{Linear}(1.0,0.1,500)$ (hard) & \\
+Learning rate & $3 \times 10^{-4}$ & \\
+MLP dims & $[512, 512]$ & \\
+Momentum coefficient ($\tau$) & $0.005$ & \\
+Noise clip & $0.3$ & \\
+N-step TD & $1$ & \\
+Policy noise & $0.2$ & \\
+UTD ratio & $1$ & \\
+\hline
+\textbf{Encoder} &  & \\
+\hline
+Discount $\gamma$ & $0.99$ & \\
+Encoder learning rate & $10^{-4}$ & \\
+Encoder MLP dims & $[256]$ & \\
+Encoder momentum coefficient ($\tau$) & 0.005 & \\
+FSQ levels & $[8, 5, 5, 5]$ & \\
+Horizon $(H)$ & $5$ & Horizon for representation learning \\
+Latent dimension ($d$) & $1024$ & \\
+% \bottomrule
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+
+\section{Full results}
 \begin{figure*}[ht]

### 2024-01-31 agents/vq_tc_td3.py
make project_latent work with use_fsq=False

@@ -188,15 +188,19 @@ class Projection(MLPResettable):
     def __init__(
         self,
         mlp_dims: List[int],
-        levels: List[int] = [8, 6, 5],  # target size 2^8, actual size 240
+        levels: Optional[List[int]] = None,
         out_dim: int = 1024,
         latent_dim: int = 1024,
     ):
         self.levels = levels
-        self.latent_dim = (latent_dim, len(levels))
+        if levels is not None:
+            self.latent_dim = (latent_dim, len(levels))
+            in_dim = np.array(self.latent_dim).prod()
+        else:
+            self.latent_dim = latent_dim
+            in_dim = self.latent_dim
         self.act_fn = None
 
-        in_dim = np.array(self.latent_dim).prod()
         mlp = h.mlp(in_dim=in_dim, mlp_dims=mlp_dims, out_dim=out_dim, act_fn=None)
 
         super().__init__(mlp=mlp)
@@ -204,8 +208,9 @@ class Projection(MLPResettable):
         self.reset(reset_type="full")
 
     def forward(self, x):
-        if x.ndim > 2:
-            x = torch.flatten(x, -2, -1)
+        if self.levels is not None:
+            if x.ndim > 2:
+                x = torch.flatten(x, -2, -1)
         z = self.mlp(x)
         # print(f"z {z.shape}")
         return z
@@ -534,16 +539,18 @@ class VQ_TC_TD3(Agent):
 
         if self.project_latent:
             if not use_fsq:
-                raise NotImplementedError
+                projection_levels = None
+            else:
+                projection_levels = fsq_levels
             self.projection = Projection(
                 mlp_dims=[1024],
-                levels=fsq_levels,
+                levels=projection_levels,
                 latent_dim=latent_dim,
                 out_dim=projection_dim,
             ).to(device)
             self.projection_target = Projection(
                 mlp_dims=[1024],
-                levels=fsq_levels,
+                levels=projection_levels,
                 latent_dim=latent_dim,
                 out_dim=projection_dim,
             ).to(device)

### 2024-02-01 paper/figs/plotting/main_plot/fetch_vq_td3_data.py
updated

@@ -4,23 +4,33 @@ ddpg_data = {
     "path": ddpg_path,
     "data": {
         # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
-        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        #### NEW ####
+        "acrobot-swingup": ["lkjyy7t0", "tek7m4yi", "7vqw7n3a", "4v4cxbjf", "omf199u5"],
+        "cheetah-run": ["nwusiqnt", "pxbu4024", "ysrel1rl", "kmj3y2ww", "bc6sqbdl"],
+        "fish-swim": ["5rhervdh", "btrclty9", "royfiwdl", "2bvv2o87", "krfia2hj"],
+        "walker-walk": ["4wydf2sz", "n7l8xac6", "l7nqlsid", "yrwwmyq2", "bpdxi4jw"],
+        "quadruped-walk": ["6gf2r90u", "uvq2ctjr", "4x1r1nqg", "lji8f92s", "eiafkell"],
         "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
-        "humanoid-run": ["jv7cfy2v"],
-        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
-        "cartpole-swingup": ["w6eykaow"],  # UTD=1
-        "hopper-stand": ["3ozramwe"],  # UTD=1
-        "quadruped-walk": ["8q59zkws"],  # UTD=1
-        # "cheetah-run": ["mkmvf5md"],  # UTD=4
-        "cheetah-run": ["jc3gjwdg"],  # UTD=1
-        "walker-walk": ["26lcg142"],  # UTD=1
-        # "walker-walk": ["nrak2mn1"],  # UTD=4
-        "quadruped-run": ["478dl8sb"],  # UTD=1
-        # 'humanoid-walk': ['2xa4zh7i', 'u01r5cd8', '2dm45ezi', '2ywqbphi', '33kix5gc', ],
-        # "dog-walk": ["gtrhfxn3"],
-        "dog-run": ["ukzsy2t5"],
         "dog-walk": ["sjqd8i9u"],
         "dog-run": ["zhsh5oyd"],
+        #### OLD ####
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
[…]

### 2024-02-01 paper/figs/plotting/main_plot/plot_main.py

@@ -15,19 +15,19 @@ plt.rcParams["font.size"] = 13
 plt.rcParams["legend.fontsize"] = 12
 plt.rcParams["legend.loc"] = "lower right"
 COLORS = {
-    "TCRL": "#e41a1c",
+    "TCRL": "#4daf4a",
     "SAC": "#377eb8",
     "REDQ": "#984ea3",
     "TD-MPC": "#ff7f00",
-    "VQ-TD3": "magenta",
+    "iFSQ-RL": "#e41a1c",
 }
 # %%
 main_envs = [
     "acrobot-swingup",
     "cheetah-run",
-    # "fish-swim",
+    "fish-swim",
     "quadruped-walk",
-    "humanoid-run",
+    # "humanoid-run",
     "walker-walk",
     "humanoid-walk",
     "dog-walk",
@@ -46,6 +46,8 @@ def plot(df, key="episode_reward"):
 
     for idx, env in enumerate(main_envs):
         data = df[df["env"] == env]
+        # breakpoint()
+        # data[data["agent"] == "iFSQ-RL"] = data[data["agent"] == "iFSQ-RL"].iloc[::2]
         row = idx // ncol
         col = idx % ncol
         ax = axs[row, col]

### 2024-02-01 paper/main.tex
use humanoid for fig 1

@@
 
-  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/bg}};
-  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/bg}};
-  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/bg}};
+  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/humanoid-1}};
+  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/humanoid-2}};
+  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/humanoid-3}};
 

### 2024-02-01 paper/main.tex
updated figures

@@
 
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/utd_comparison.pdf}}
-\caption{\textbf{Normalizing the latent-space enables higher UTD} \our's (blue) sample efficiency improves when the UTD ratio is increased from $1 \rightarrow 8$.
+% \begin{figure*}[ht]
+% \vskip 0.2in
+% \begin{center}
+% \centerline{\includegraphics[width=1.0\textwidth]{./figs/utd_comparison.pdf}}
+% \caption{\textbf{Normalizing the latent-space enables higher UTD} \our's (blue) sample efficiency improves when the UTD ratio is increased from $1 \rightarrow 8$.
 TCRL (which has an un-normalized latent space) does not see an improvement in sample efficiency when when the UTD ratio is increased.
@@
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:normalization_enables_higher_utd}
-\end{center}
-\vskip -0.2in
-\end{figure*}
+% \label{fig:normalization_enables_higher_utd}
+% \end{center}
+% \vskip -0.2in
+% \end{figure*}
 
-\textbf{Normalizing the latent-space enables higher UTD}
-We first show how using bounded latent states enables us to increase the UTD ratio to see improvements in sample efficiency.
-\cref{fig:normalization_enables_higher_utd} shows that as the UTD ratio is increased from $1$ to $8$ \our's sample efficiency
-continues to improve.
-In contrast, when increasing the UTD ratio for TCRL no improvements are observed.
+% \textbf{Normalizing the latent-space enables higher UTD}
+% We first show how using bounded latent states enables us to increase the UTD ratio to see improvements in sample efficiency.
+% \cref{fig:normalization_enables_higher_utd} shows that as the UTD ratio is increased from $1$ to $8$ \our's sample efficiency
+% continues to improve.
+% In contrast, when increasing the UTD ratio for TCRL no improvements are observed.
 
@@
 \begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/utd_comparison.pdf}}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
+\caption{\textbf{Ablation of normalization}.
+\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:multi-task-pretraining}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=0.3\textwidth]{example-image}}
 \caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.

### 2024-02-01 paper/main.tex
mini updaets

@@
 
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/fsq_levels_comparison.pdf}}
-\caption{\textbf{Comparison of different FSQ levels $\mathcal{L}$}.
+% \begin{figure*}[ht]
+% \vskip 0.2in
+% \begin{center}
+% \centerline{\includegraphics[width=1.0\textwidth]{./figs/fsq_levels_comparison.pdf}}
+% \caption{\textbf{Comparison of different FSQ levels $\mathcal{L}$}.
 Results are for four DMC tasks.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:fsq_levels_comparison}
-\end{center}
-\vskip -0.2in
-\end{figure*}
+% \label{fig:fsq_levels_comparison}
+% \end{center}
+% \vskip -0.2in
+% \end{figure*}
 
@@
 \textbf{Training} & & \\
-\hline
 Max episode length & 1000 &  Max episode length \\
@@
 \textbf{TD3} & & \\
-\hline
 Actor update freq & $2$  & update actor less frequently than critic \\
 Batch size & $256$ & \\
-Buffer size & $1e6$ & \\
+Buffer size & $10^{6}$ & \\
 Exploration noise & $\mathrm{Linear}(1.0,0.1,50)$ (easy) & \\
@@
 \textbf{Encoder} &  & \\
-\hline
 Discount $\gamma$ & $0.99$ & \\
@@
 
-\section{Full results}
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=0.97\textwidth]{./figs/all_mujoco_envs.pdf}}
-\caption{Comparison to SAC/TCRL with UTD=1.}
-\label{fig:main_plot}
-\end{center}
-\vskip -0.2in
-\end{figure*}
+% \section{Full results}
+% \begin{figure*}[ht]
+% \vskip 0.2in
+% \begin{center}
+% \centerline{\includegraphics[width=0.97\textwidth]{./figs/all_mujoco_envs.pdf}}
+% \caption{Comparison to SAC/TCRL with UTD=1.}
+% \label{fig:main_plot}
+% \end{center}
+% \vskip -0.2in
+% \end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-02-01 paper/figs/plotting/norm-ablation-plot/fetch_vq_td3_data.py
initial commit

@@ -1,0 +1,189 @@
+# %%
+ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        # "humanoid-run": ["jv7cfy2v"],
+        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
+        # "cartpole-swingup": ["w6eykaow"],  # UTD=1
+        # "hopper-stand": ["3ozramwe"],  # UTD=1
+        # "quadruped-walk": ["8q59zkws"],  # UTD=1
+        # "cheetah-run": ["mkmvf5md"],  # UTD=4
+        # : [,],
+        "hopper-stand": [
+            # no-norm, d=50
+            "25m4987b",  # seed=1
+            "cet5h22j",  # seed=2
+            "8o4v1gz1",  # seed=3
+            "udj85ac1",  # seed=4
+            "7k9kjfll",  # seed=5
+            # no-norm, d=512
+            "hnbr2b0o",  # seed=1
+            "el8y18hl",  # seed=2
+            "1dx4600a",  # seed=3
+            "vxw4sgkb",  # seed=4
+            "ozh83jpd",  # seed=5
+            # norm, d=512
+            "1wmalfbs",  # seed=1
+            "rj3vhnbq",  # seed=2
+            "j4iwbxq2",  # seed=3
+            "jo9td3sx",  # seed=4
+            "71nsurzh",  # seed=5
+        ],
+        "quadruped-run": [
+            # no-norm, d=50
+            "lfay8jdo",  # seed=1
+            "r40qcslq",  # seed=2
+            "sgc9fa90",  # seed=3
+            "fm07n1cn",  # seed=4
+            "bbmqgmfm",  # seed=5
+            # no-norm, d=512
+            "cuaukeiq",  # seed=1
+            "n8xdctnq",  # seed=2
+            "afuu4r4x",  # seed=3
+            "a9qzumnn",  # seed=4
+            "drhxy0x4",  # seed=5
[…]

### 2024-02-01 paper/figs/plotting/norm-ablation-plot/make_plot.py

@@ -1,0 +1,140 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+COLORS = {
+    "TCRL": "#e41a1c",
+    "SAC": "#377eb8",
+    "REDQ": "#984ea3",
+    "TD-MPC": "#ff7f00",
+    "VQ-TD3": "magenta",
+}
+# %%
+main_envs = [
+    # "acrobot-swingup",
+    "cheetah-run",
+    # "walker-walk",
+    "walker-run",
+    "hopper-stand",
+    # "fish-swim",
+    "quadruped-run",
+    # "humanoid-run",
+    # "humanoid-walk",
+    # "dog-walk",
+    # "dog-run",
+]
+
+# pet-pytorch
+titles = [
+    "env_step",
+    "episode",
+    "episode_reward",
+    "time",
+]
+keys = [
+    "eval/.env_step",
+    "eval/.episode",
+    "eval/.episodic_return",
+    "eval/.elapsed_time",
+]
+
+rename = {a: b for a, b in zip(keys, titles)}
+
+
+def plot(df, key="episode_reward"):
+    # breakpoint()
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    # nrow = len(main_envs) // ncol
+    nrow = 1
+    # nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    df = df.rename(columns=rename)
+
+    for idx, env in enumerate(main_envs):
+        data = df[df["env"] == env]
+        row = idx // ncol
+        col = idx % ncol
+        # ax = axs[row, col]
+        ax = axs[col]
+        # hue_order = data.agent.unique()
+        hue_order = data.name.unique()
+        # hue_order = data.utd_ratio.unique()
+        # breakpoint()
+
+        if idx == 0:
+            sns.lineplot(
+                x="env_step",
+                # x="episode",
+                y=key,
+                data=data,
[…]

### 2024-02-01 paper/figs/plotting/main_plot/fetch_vq_td3_data.py
updated

@@ -10,9 +10,51 @@ ddpg_data = {
         "fish-swim": ["5rhervdh", "btrclty9", "royfiwdl", "2bvv2o87", "krfia2hj"],
         "walker-walk": ["4wydf2sz", "n7l8xac6", "l7nqlsid", "yrwwmyq2", "bpdxi4jw"],
         "quadruped-walk": ["6gf2r90u", "uvq2ctjr", "4x1r1nqg", "lji8f92s", "eiafkell"],
-        "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
-        "dog-walk": ["sjqd8i9u"],
-        "dog-run": ["zhsh5oyd"],
+        "humanoid-walk": [
+            "wur1u276",
+            "l8da5827",
+            "oy053sve",
+            "krctc38z",
+            "anstkvup",
+        ],  # project=False
+        # "humanoid-walk": [
+        #     "bk32bbsi",
+        #     "fr1hpr9b",
+        #     "t9hbnhff",
+        #     "681rgba7",
+        #     "ndd0tr73",
+        # ],  # project=True
+        # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        "dog-walk": [
+            "stawiad8",
+            "iv8p110s",
+            "k454j543",
+            "nihkvll9",
+            "jpkbw5h7",
+        ],  # project=False
+        # "dog-walk": [
+        #     "9k5cqnuq",
+        #     "q28f7f9i",
+        #     "2vf4clxu",
+        #     "8nnpgifp",
+        #     "b5ffm8rt",
+        # ],  # project=True
+        # "dog-walk": ["sjqd8i9u"],
+        "dog-run": [
+            "ohxmhffe",
+            "aikn5eix",
+            "kprj5q9u",
+            "r232jms2",
+            "aox6spn2",
+        ],  # project=False
+        # "dog-run": [
+        #     "7hty9mrz",
+        #     "0no65695",
+        #     "nfuag4aj",
+        #     "otirlyw0",
+        #     "t9oyygr9",
+        # ],  # project=True
+        # "dog-run": ["zhsh5oyd"],
         #### OLD ####
         # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
         # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],

### 2024-02-01 paper/figs/plotting/main_plot/plot_main.py

@@ -84,7 +84,7 @@ def plot(df, key="episode_reward"):
         ax.set_xlabel("Environment Steps (1e3)")
         ax.set_ylabel("Episode Return")
     plt.tight_layout()
-    plt.savefig(f"main_plot_new.pdf")
+    plt.savefig(f"baselines_comparison.pdf")
     # plt.show()
 
 

### 2024-02-01 paper/figs/plotting/norm-ablation-plot/fetch_data.py

@@ -135,7 +135,7 @@ keys = [
 rename = {a: b for a, b in zip(keys, titles)}
 
 
-def fetch_results(run_path, run_name_list, keys, agent_name):
+def fetch_results(run_path, run_name_list, keys):
     data = []
     for run_name in run_name_list:
         wandb_run = api.run(run_path + run_name)
@@ -160,7 +160,7 @@ def fetch_results(run_path, run_name_list, keys, agent_name):
         if wandb_run.config["agent"]["use_fsq"]:
             history["name"] = "iFSQ"
         else:
-            history["name"] = "no-norm"
+            history["name"] = f"no-norm $d={wandb_run.config['agent']['latent_dim']}$"
         history["utd_ratio"] = wandb_run.config["agent"]["utd_ratio"]
 
         data.append(history)
@@ -175,7 +175,7 @@ data = pd.concat(
             run_path=ddpg_data["path"],
             run_name_list=ddpg_data["data"][env],
             keys=keys,
-            agent_name="VQ-TD3",
+            # agent_name="VQ-TD3",
         )
         for env in ddpg_data["data"].keys()
     ]
@@ -183,7 +183,7 @@ data = pd.concat(
 
 # %%
 
-data.to_csv("./vq_td3_main.csv")
+data.to_csv("./ifsq-rl.csv")
 
 
 # %%

### 2024-02-01 paper/figs/plotting/norm-ablation-plot/make_plot.py

@@ -112,7 +112,7 @@ def plot(df, key="episode_reward"):
         ax.set_xlabel("Environment Steps (1e3)")
         ax.set_ylabel("Episode Return")
     plt.tight_layout()
-    plt.savefig(f"normalization-ablation.pdf")
+    plt.savefig(f"../../normalization-ablation.pdf")
     # plt.show()
 
 
@@ -133,7 +133,7 @@ df = [
     # pd.read_csv(f"{data_path}/tdmpc_main.csv"),
     # df_redq,
     # pd.read_csv(f"{data_path}/sac_main.csv"),
-    pd.read_csv(f"{data_path}/vq_td3_main.csv"),
+    pd.read_csv(f"{data_path}/ifsq-rl.csv"),
 ]
 plot(pd.concat(df))
 

### 2024-02-01 paper/figs/plotting/baseline-comparison/make_plot.py
moved

@@ -14,36 +14,47 @@ plt.rcParams["figure.dpi"] = 400
 plt.rcParams["font.size"] = 13
 plt.rcParams["legend.fontsize"] = 12
 plt.rcParams["legend.loc"] = "lower right"
-
-# %%
-# first process SAC data
 COLORS = {
-    "TCRL": "#e41a1c",
+    "TCRL": "#4daf4a",
     "SAC": "#377eb8",
-    "SAC-our": "#984ea3",
-    "VQ-TD3": "magenta",
+    "REDQ": "#984ea3",
+    "TD-MPC": "#ff7f00",
+    "iFSQ-RL": "#e41a1c",
 }
+# %%
+main_envs = [
+    "acrobot-swingup",
+    "cheetah-run",
+    "fish-swim",
+    "quadruped-walk",
+    # "humanoid-run",
+    "walker-walk",
+    "humanoid-walk",
+    "dog-walk",
+    "dog-run",
+]
 
 
 def plot(df, key="episode_reward"):
-    envs = np.sort(env_list)
+    envs = np.sort(df.env.unique())
     ncol = 4
     # assert envs.shape[0] % ncol == 0
-    nrow = envs.shape[0] // ncol
+    nrow = len(main_envs) // ncol
+    # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
-    for idx, env in enumerate(envs):
+    for idx, env in enumerate(main_envs):
         data = df[df["env"] == env]
-        data = data[data["episode"] <= max_ep[env]]
+        # breakpoint()
+        # data[data["agent"] == "iFSQ-RL"] = data[data["agent"] == "iFSQ-RL"].iloc[::2]
         row = idx // ncol
         col = idx % ncol
         ax = axs[row, col]
         hue_order = data.agent.unique()
 
-        if idx == 0:
+        if idx == 4:
             sns.lineplot(
-                # x="episode",
                 x="env_step",
                 y=key,
                 data=data,
@@ -56,9 +67,9 @@ def plot(df, key="episode_reward"):
             )
             ax.legend().set_title(None)
         else:
+            # breakpoint()
             sns.lineplot(
                 x="env_step",
-                # x="episode",
                 y=key,
                 data=data,
[…]

### 2024-02-01 paper/figs/plotting/main_plot/plot_main.py

@@ -1,122 +1,0 @@
-# %%
-import pandas as pd
-import numpy as np
-
-
-import matplotlib.pyplot as plt
-from matplotlib import cm
-
-# plt.style.use('bmh')
-plt.style.use("seaborn-v0_8-whitegrid")
-import seaborn as sns
-
-plt.rcParams["figure.dpi"] = 400
-plt.rcParams["font.size"] = 13
-plt.rcParams["legend.fontsize"] = 12
-plt.rcParams["legend.loc"] = "lower right"
-COLORS = {
-    "TCRL": "#4daf4a",
-    "SAC": "#377eb8",
-    "REDQ": "#984ea3",
-    "TD-MPC": "#ff7f00",
-    "iFSQ-RL": "#e41a1c",
-}
-# %%
-main_envs = [
-    "acrobot-swingup",
-    "cheetah-run",
-    "fish-swim",
-    "quadruped-walk",
-    # "humanoid-run",
-    "walker-walk",
-    "humanoid-walk",
-    "dog-walk",
-    "dog-run",
-]
-
-
-def plot(df, key="episode_reward"):
-    envs = np.sort(df.env.unique())
-    ncol = 4
-    # assert envs.shape[0] % ncol == 0
-    nrow = len(main_envs) // ncol
-    # nrow = envs.shape[0] // ncol
-
-    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
-
-    for idx, env in enumerate(main_envs):
-        data = df[df["env"] == env]
-        # breakpoint()
-        # data[data["agent"] == "iFSQ-RL"] = data[data["agent"] == "iFSQ-RL"].iloc[::2]
-        row = idx // ncol
-        col = idx % ncol
-        ax = axs[row, col]
-        hue_order = data.agent.unique()
-
-        if idx == 4:
-            sns.lineplot(
-                x="env_step",
-                y=key,
-                data=data,
-                errorbar=("ci", 95),
-                hue="agent",
-                hue_order=hue_order,
-                palette=COLORS,
-                legend="auto",
-                ax=ax,
-            )
-            ax.legend().set_title(None)
-        else:
-            # breakpoint()
-            sns.lineplot(
-                x="env_step",
-                y=key,
-                data=data,
-                errorbar=("ci", 95),
-                hue="agent",
[…]

### 2024-02-01 paper/figs/plotting/norm-ablation-plot/fetch_data.py

@@ -1,189 +1,0 @@
-# %%
-ddpg_path = "aalto-ml/lifelong-td3-tc/"
-ddpg_data = {
-    "path": ddpg_path,
-    "data": {
-        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
-        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
-        # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
-        # "humanoid-run": ["jv7cfy2v"],
-        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
-        # "cartpole-swingup": ["w6eykaow"],  # UTD=1
-        # "hopper-stand": ["3ozramwe"],  # UTD=1
-        # "quadruped-walk": ["8q59zkws"],  # UTD=1
-        # "cheetah-run": ["mkmvf5md"],  # UTD=4
-        # : [,],
-        "hopper-stand": [
-            # no-norm, d=50
-            "25m4987b",  # seed=1
-            "cet5h22j",  # seed=2
-            "8o4v1gz1",  # seed=3
-            "udj85ac1",  # seed=4
-            "7k9kjfll",  # seed=5
-            # no-norm, d=512
-            "hnbr2b0o",  # seed=1
-            "el8y18hl",  # seed=2
-            "1dx4600a",  # seed=3
-            "vxw4sgkb",  # seed=4
-            "ozh83jpd",  # seed=5
-            # norm, d=512
-            "1wmalfbs",  # seed=1
-            "rj3vhnbq",  # seed=2
-            "j4iwbxq2",  # seed=3
-            "jo9td3sx",  # seed=4
-            "71nsurzh",  # seed=5
-        ],
-        "quadruped-run": [
-            # no-norm, d=50
-            "lfay8jdo",  # seed=1
-            "r40qcslq",  # seed=2
-            "sgc9fa90",  # seed=3
-            "fm07n1cn",  # seed=4
-            "bbmqgmfm",  # seed=5
-            # no-norm, d=512
-            "cuaukeiq",  # seed=1
-            "n8xdctnq",  # seed=2
-            "afuu4r4x",  # seed=3
-            "a9qzumnn",  # seed=4
-            "drhxy0x4",  # seed=5
[…]

### 2024-02-01 paper/figs/plotting/norm-ablation-plot/make_plot.py

@@ -1,140 +1,0 @@
-# %%
-import pandas as pd
-import numpy as np
-
-
-import matplotlib.pyplot as plt
-from matplotlib import cm
-
-# plt.style.use('bmh')
-plt.style.use("seaborn-v0_8-whitegrid")
-import seaborn as sns
-
-plt.rcParams["figure.dpi"] = 400
-plt.rcParams["font.size"] = 13
-plt.rcParams["legend.fontsize"] = 12
-plt.rcParams["legend.loc"] = "lower right"
-COLORS = {
-    "TCRL": "#e41a1c",
-    "SAC": "#377eb8",
-    "REDQ": "#984ea3",
-    "TD-MPC": "#ff7f00",
-    "VQ-TD3": "magenta",
-}
-# %%
-main_envs = [
-    # "acrobot-swingup",
-    "cheetah-run",
-    # "walker-walk",
-    "walker-run",
-    "hopper-stand",
-    # "fish-swim",
-    "quadruped-run",
-    # "humanoid-run",
-    # "humanoid-walk",
-    # "dog-walk",
-    # "dog-run",
-]
-
-# pet-pytorch
-titles = [
-    "env_step",
-    "episode",
-    "episode_reward",
-    "time",
-]
-keys = [
-    "eval/.env_step",
-    "eval/.episode",
-    "eval/.episodic_return",
-    "eval/.elapsed_time",
-]
-
-rename = {a: b for a, b in zip(keys, titles)}
-
-
-def plot(df, key="episode_reward"):
-    # breakpoint()
-    envs = np.sort(df.env.unique())
-    ncol = 4
-    # assert envs.shape[0] % ncol == 0
-    # nrow = len(main_envs) // ncol
-    nrow = 1
-    # nrow = envs.shape[0] // ncol
-
-    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
-
-    df = df.rename(columns=rename)
-
-    for idx, env in enumerate(main_envs):
-        data = df[df["env"] == env]
-        row = idx // ncol
-        col = idx % ncol
-        # ax = axs[row, col]
-        ax = axs[col]
-        # hue_order = data.agent.unique()
-        hue_order = data.name.unique()
-        # hue_order = data.utd_ratio.unique()
-        # breakpoint()
-
-        if idx == 0:
-            sns.lineplot(
-                x="env_step",
-                # x="episode",
-                y=key,
-                data=data,
[…]

### 2024-02-01 paper/figs/plotting/norm-ablation/fetch_data.py
initial commit

@@ -1,0 +1,189 @@
+# %%
+ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        # "humanoid-run": ["jv7cfy2v"],
+        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
+        # "cartpole-swingup": ["w6eykaow"],  # UTD=1
+        # "hopper-stand": ["3ozramwe"],  # UTD=1
+        # "quadruped-walk": ["8q59zkws"],  # UTD=1
+        # "cheetah-run": ["mkmvf5md"],  # UTD=4
+        # : [,],
+        "hopper-stand": [
+            # no-norm, d=50
+            "25m4987b",  # seed=1
+            "cet5h22j",  # seed=2
+            "8o4v1gz1",  # seed=3
+            "udj85ac1",  # seed=4
+            "7k9kjfll",  # seed=5
+            # no-norm, d=512
+            "hnbr2b0o",  # seed=1
+            "el8y18hl",  # seed=2
+            "1dx4600a",  # seed=3
+            "vxw4sgkb",  # seed=4
+            "ozh83jpd",  # seed=5
+            # norm, d=512
+            "1wmalfbs",  # seed=1
+            "rj3vhnbq",  # seed=2
+            "j4iwbxq2",  # seed=3
+            "jo9td3sx",  # seed=4
+            "71nsurzh",  # seed=5
+        ],
+        "quadruped-run": [
+            # no-norm, d=50
+            "lfay8jdo",  # seed=1
+            "r40qcslq",  # seed=2
+            "sgc9fa90",  # seed=3
+            "fm07n1cn",  # seed=4
+            "bbmqgmfm",  # seed=5
+            # no-norm, d=512
+            "cuaukeiq",  # seed=1
+            "n8xdctnq",  # seed=2
+            "afuu4r4x",  # seed=3
+            "a9qzumnn",  # seed=4
+            "drhxy0x4",  # seed=5
[…]

### 2024-02-01 paper/figs/plotting/norm-ablation/make_plot.py

@@ -1,0 +1,143 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+COLORS = {
+    # "TCRL": "#e41a1c",
+    # "SAC": "#377eb8",
+    # "REDQ": "#984ea3",
+    # "TD-MPC": "#ff7f00",
+    # "VQ-TD3": "magenta",
+    "iFSQ-RL": "#e41a1c",
+    "no-norm $d=50$": "#377eb8",
+    "no-norm $d=512$": "#ff7f00",
+}
+# %%
+main_envs = [
+    # "acrobot-swingup",
+    "cheetah-run",
+    # "walker-walk",
+    "walker-run",
+    "hopper-stand",
+    # "fish-swim",
+    "quadruped-run",
+    # "humanoid-run",
+    # "humanoid-walk",
+    # "dog-walk",
+    # "dog-run",
+]
+
+# pet-pytorch
+titles = [
+    "env_step",
+    "episode",
+    "episode_reward",
+    "time",
+]
+keys = [
+    "eval/.env_step",
+    "eval/.episode",
+    "eval/.episodic_return",
+    "eval/.elapsed_time",
+]
+
+rename = {a: b for a, b in zip(keys, titles)}
+
+
+def plot(df, key="episode_reward"):
+    # breakpoint()
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    # nrow = len(main_envs) // ncol
+    nrow = 1
+    # nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    df = df.rename(columns=rename)
+
+    for idx, env in enumerate(main_envs):
+        data = df[df["env"] == env]
+        row = idx // ncol
+        col = idx % ncol
+        # ax = axs[row, col]
+        ax = axs[col]
+        # hue_order = data.agent.unique()
+        hue_order = data.name.unique()
+        # hue_order = data.utd_ratio.unique()
+        # breakpoint()
+
[…]

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/fetch_data.py

@@ -1,0 +1,186 @@
+# %%
+ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        # "humanoid-run": ["jv7cfy2v"],
+        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
+        # "cartpole-swingup": ["w6eykaow"],  # UTD=1
+        # "hopper-stand": ["3ozramwe"],  # UTD=1
+        # "quadruped-walk": ["8q59zkws"],  # UTD=1
+        # "cheetah-run": ["mkmvf5md"],  # UTD=4
+        # : [,],
+        "hopper-stand": [
+            # reconstruction
+            "25m4987b",  # seed=1
+            "cet5h22j",  # seed=2
+            "8o4v1gz1",  # seed=3
+            "udj85ac1",  # seed=4
+            "7k9kjfll",  # seed=5
+            # no reconstruction
+            "1wmalfbs",  # seed=1
+            "rj3vhnbq",  # seed=2
+            "j4iwbxq2",  # seed=3
+            "jo9td3sx",  # seed=4
+            "71nsurzh",  # seed=5
+        ],
+        "quadruped-run": [
+            # reconstruction
+            "rw26pb1l",  # seed=1
+            "jm549p6m",  # seed=2
+            "7ll3mp6d",  # seed=3
+            "ypgoji3k",  # seed=4
+            "ll54kaha",  # seed=5
+            # no reconstruction
+            "w625abb9",  # seed=1
+            "w1ssoh9i",  # seed=2
+            "wtobbh3o",  # seed=3
+            "4sj3p04r",  # seed=4
+            "wtobbh3o",  # seed=5
+        ],
+        "walker-run": [
+            # reconstruction
+            "qh3xxjz0",  # seed=1
+            "1hjix0d3",  # seed=2
+            "7agg1jz9",  # seed=3
+            "dnqzhj7e",  # seed=4
+            "kabd20ca",  # seed=5
[…]

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/make_plot.py

@@ -1,0 +1,142 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+COLORS = {
+    # "TCRL": "#e41a1c",
+    # "SAC": ,
+    # "REDQ": "#984ea3",
+    # "TD-MPC": "#ff7f00",
+    # "VQ-TD3": "magenta",
+    "iFSQ-RL+rec": "#377eb8",
+    "iFSQ-RL": "#e41a1c",
+}
+# %%
+main_envs = [
+    # "acrobot-swingup",
+    # "cheetah-run",
+    # "walker-walk",
+    "walker-run",
+    "hopper-stand",
+    # "fish-swim",
+    "quadruped-run",
+    # "humanoid-run",
+    # "humanoid-walk",
+    "dog-walk",
+    # "dog-run",
+]
+
+# pet-pytorch
+titles = [
+    "env_step",
+    "episode",
+    "episode_reward",
+    "time",
+]
+keys = [
+    "eval/.env_step",
+    "eval/.episode",
+    "eval/.episodic_return",
+    "eval/.elapsed_time",
+]
+
+rename = {a: b for a, b in zip(keys, titles)}
+
+
+def plot(df, key="episode_reward"):
+    # breakpoint()
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    # nrow = len(main_envs) // ncol
+    nrow = 1
+    # nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    df = df.rename(columns=rename)
+
+    for idx, env in enumerate(main_envs):
+        data = df[df["env"] == env]
+        row = idx // ncol
+        col = idx % ncol
+        # ax = axs[row, col]
+        ax = axs[col]
+        # hue_order = data.agent.unique()
+        hue_order = data.name.unique()
+        # hue_order = data.utd_ratio.unique()
+        # breakpoint()
+
+        if idx == 0:
+            sns.lineplot(
[…]

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/fetch_data.py
updated figures

@@ -14,68 +14,129 @@ ddpg_data = {
         # "cheetah-run": ["mkmvf5md"],  # UTD=4
         # : [,],
         "hopper-stand": [
+            ##### Project=True
+            # # reconstruction
+            # "25m4987b",  # seed=1
+            # "cet5h22j",  # seed=2
+            # "8o4v1gz1",  # seed=3
+            # "udj85ac1",  # seed=4
+            # "7k9kjfll",  # seed=5
+            # # no reconstruction
+            # "1wmalfbs",  # seed=1
+            # "rj3vhnbq",  # seed=2
+            # "j4iwbxq2",  # seed=3
+            # "jo9td3sx",  # seed=4
+            # "71nsurzh",  # seed=5
             # reconstruction
-            "25m4987b",  # seed=1
-            "cet5h22j",  # seed=2
-            "8o4v1gz1",  # seed=3
-            "udj85ac1",  # seed=4
-            "7k9kjfll",  # seed=5
+            "94skam71",  # seed=1
+            "hss7p9jr",  # seed=2
+            "oe1us2qh",  # seed=3
+            "nijlewdk",  # seed=4
+            "4m35m7d4",  # seed=5
             # no reconstruction
-            "1wmalfbs",  # seed=1
-            "rj3vhnbq",  # seed=2
-            "j4iwbxq2",  # seed=3
-            "jo9td3sx",  # seed=4
-            "71nsurzh",  # seed=5
+            "3rmiy5he",  # seed=1
+            "39j1gbp2",  # seed=2
+            "t4hf7g18",  # seed=3
+            "j8apohl8",  # seed=4
+            "5w8halqc",  # seed=5
         ],
         "quadruped-run": [
+            ##### Project=True
+            # # reconstruction
+            # "rw26pb1l",  # seed=1
+            # "jm549p6m",  # seed=2
+            # "7ll3mp6d",  # seed=3
+            # "ypgoji3k",  # seed=4
+            # "ll54kaha",  # seed=5
+            # # no reconstruction
+            # "w625abb9",  # seed=1
+            # "w1ssoh9i",  # seed=2
+            # "wtobbh3o",  # seed=3
+            # "4sj3p04r",  # seed=4
+            # "wtobbh3o",  # seed=5
+            ##### Project=False
             # reconstruction
-            "rw26pb1l",  # seed=1
-            "jm549p6m",  # seed=2
-            "7ll3mp6d",  # seed=3
[…]

### 2024-02-01 paper/main.tex
udpatedc
add reconstruction ablation
Start fig 1

--- 2024-02-01 paper/main.tex
@@
 % \icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
-\icmltitle{Sample-efficient Reinforcement Learning with Implicitly Quantized Representations}
+\icmltitle{Sample-efficient Reinforcement Learning \\ with Implicitly Quantized Representations}
 
@@
 \centering\scriptsize
-\resizebox{\columnwidth}{!}{%
+%\resizebox{\columnwidth}{!}{%
+%\begin{tikzpicture}
+%
+%  % Declare that we have a background layer (advanced)
+%  \pgfdeclarelayer{background}
+%  \pgfsetlayers{background,main}
+%
+%  % Helpers for small subscripts
+%  \newcommand{\sub}[0]{\scalebox{.8}{\ensuremath t}}
+%  \newcommand{\subs}[1]{\scalebox{.8}{\ensuremath t{+}#1}}
+%
+%  % Lengths (this tunes distance)
+%  \newlength{\nodedist}\setlength{\nodedist}{2cm}
+%
+%  % Styles (this tunes colours etc.)
+%  \tikzstyle{mynode}=[fill=blue!10!cyan,rounded corners=1pt,font=\tiny,inner sep=0]
+%  \tikzstyle{trap}=[mynode,trapezium,minimum width=1.5em, minimum height=2em,
+%                    trapezium left angle=-70, trapezium right angle=-70]
+%  \tikzstyle{blob}=[mynode,circle,minimum width=2.4em,minimum height=2.4em]
+%  \tikzstyle{polval}=[mynode,minimum width=6em,minimum height=2em,align=center,text=white,fill=black!60,inner sep=2pt]
+%  \tikzstyle{arr}=[line width=1pt,black,->]
+%  \tikzstyle{darr}=[line width=1pt,black,<->,densely dotted]
+%    
+%  % *** FIRST HALF ***  
+%    
+%  % Draw nodes
+%  \node[blob] (z0) at (0\nodedist,0) {$\vz_{\sub}$};
+%  \node[blob] (z1) at (1\nodedist,0) {$\hat\vz_{\subs{1}}$};
+%  \node[blob] (z2) at (2\nodedist,0) {$\hat\vz_{\subs{2}}$};
+%
+%  \node[trap] (e0) at (0\nodedist,\nodedist) {$e_\phi$};
+%  \node[trap] (e1) at (1\nodedist,\nodedist) {$e_{\bar\phi}$};
+%  \node[trap] (e2) at (2\nodedist,\nodedist) {$e_{\bar\phi}$};
+%
+%  \node[blob] (zh1) at ($(z1)!0.5!(e1)$) {$\bar\vz_{\subs{1}}$};
+%  \node[blob] (zh2) at ($(z2)!0.5!(e2)$) {$\bar\vz_{\subs{2}}$};
+%
+%  \coordinate (m0) at ($(z0)!.5!(z1)$);
+%  \node[blob] (a0) at ($(m0)!(zh1)!(m0)$) {$\va_{\sub}$};
+%  \coordinate (m1) at ($(z1)!.5!(z2)$);
+%  \node[blob] (a1) at ($(m1)!(zh2)!(m1)$) {$\va_{\subs{1}}$};
+%  
+%  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/bg}};
+%  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/bg}};
+%  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/bg}};
+%  
+%  \node[anchor=north] (o0) at (i0.north) {\color{white}$o_{\sub}$};
+%  \node[anchor=north] (o1) at (i1.north) {\color{white}$o_{\subs{1}}$};
+%  \node[anchor=north] (o2) at (i2.north) {\color{white}$o_{\subs{2}}$};
+%  
+%  \node[above of=i1,align=center] (lab0) {\normalsize\bf 1.\ Learn representation};
+%    
+%  % Draw arrows
+%  \draw[arr] (i0) -- (e0);
+%  \draw[arr] (i1) -- (e1);
+%  \draw[arr] (i2) -- (e2);
+%  \draw[arr] (z0) -- (z1);
+%  \draw[arr] (z1) -- (z2);
+%  \draw[arr] (e0) -- (z0);
+%  \draw[arr] (e1) -- (zh1);
+%  \draw[arr] (e2) -- (zh2);
+%  \draw[arr] (a0) -- (z1);
+%  \draw[arr] (a1) -- (z2);
+%  \draw[darr] (zh1) -- (z1);
+%  \draw[darr] (zh2) -- (z2); 
+%  
+%  
+%  % *** SECOND HALF ***
+%  
+%  \node[blob] (z) at ($(e2) + (1.5\nodedist,0)$) {$\bar\vz_{\sub}$};
+%  \node[polval] (policy) at ($(z) + (-.5\nodedist,-.5\nodedist)$) {{\bf Policy} \\ $\pi_\theta(\va_t \mid \bar\vz_t)$};
+%  \node[polval] (value) at ($(z) + (.5\nodedist,-.5\nodedist)$) {{\bf Value} \\ $\mathcal{Q}_\psi(\bar\vz_t, \va_t)$};
+%  
+%  \draw[arr] (z) -- (policy);
+%  \draw[arr] (z) -- (value);
+%  
+%  \node[align=center] at ($(z)!(lab0)!(z)$) {\normalsize\bf 2.\ Latent actor-critic\vphantom{p}};  
+%
+%  % Draw backgrounds
+%  \begin{pgfonlayer}{background}
+%  
+%    \coordinate (mid) at ($(i2.east)!.5!(policy.west)$);  
+%    
+%    \draw[draw=none,fill=black!10,rounded corners=5pt] ($(i0.north west) + (-.25cm,1cm)$) 
+%      rectangle ($(mid)!(z2.north)!(mid) + (-3pt,-1cm)$);
+%      
+%    \draw[draw=none,fill=black!05,rounded corners=5pt] ($(mid)!(i2.north east)!(mid) + (3pt,1cm)$) 
+%      rectangle ($(value.east)!(z2.north)!(value.east) + (.25cm,-1cm)$);
+%      
+%  \end{pgfonlayer}
+%  
+%\end{tikzpicture}}
+
 \begin{tikzpicture}
@@
   % Lengths (this tunes distance)
-  \newlength{\nodedist}\setlength{\nodedist}{2cm}
+  \newlength{\nodedist}\setlength{\nodedist}{3cm}
+  \newlength{\nodedistv}\setlength{\nodedistv}{2.5cm}
 
   % Styles (this tunes colours etc.)
-  \tikzstyle{mynode}=[fill=blue!10!cyan,rounded corners=1pt,font=\tiny,inner sep=0]
+  \tikzstyle{mynode}=[fill=white,draw=black!80,rounded corners=1pt,font=\tiny,inner sep=0]
   \tikzstyle{trap}=[mynode,trapezium,minimum width=1.5em, minimum height=2em,
@@
 
-  % *** FIRST HALF ***
-
   % Draw nodes
@@
 
-  \node[trap] (e0) at (0\nodedist,\nodedist) {$e_\phi$};
-  \node[trap] (e1) at (1\nodedist,\nodedist) {$e_{\bar\phi}$};
-  \node[trap] (e2) at (2\nodedist,\nodedist) {$e_{\bar\phi}$};
+  \node[trap] (e0) at (0\nodedist,\nodedistv) {$e_\phi$};
+  \node[trap] (e1) at (1\nodedist,\nodedistv) {$e_{\bar\phi}$};
+  \node[trap] (e2) at (2\nodedist,\nodedistv) {$e_{\bar\phi}$};
 
@@
 
-  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/humanoid-1}};
-  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/humanoid-2}};
-  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/humanoid-3}};
+  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/bg}};
+  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/bg}};
+  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/bg}};
 
@@
 
-  % *** SECOND HALF ***
-
-  \node[blob] (z) at ($(e2) + (1.5\nodedist,0)$) {$\bar\vz_{\sub}$};
-  \node[polval] (policy) at ($(z) + (-.5\nodedist,-.5\nodedist)$) {{\bf Policy} \\ $\pi_\theta(\va_t \mid \bar\vz_t)$};
-  \node[polval] (value) at ($(z) + (.5\nodedist,-.5\nodedist)$) {{\bf Value} \\ $\mathcal{Q}_\psi(\bar\vz_t, \va_t)$};
-
-  \draw[arr] (z) -- (policy);
-  \draw[arr] (z) -- (value);
-
-  \node[align=center] at ($(z)!(lab0)!(z)$) {\normalsize\bf 2.\ Latent actor-critic\vphantom{p}};
-
-  % Draw backgrounds
-  \begin{pgfonlayer}{background}
-
-    \coordinate (mid) at ($(i2.east)!.5!(policy.west)$);
-
-    \draw[draw=none,fill=black!10,rounded corners=5pt] ($(i0.north west) + (-.25cm,1cm)$) 
-      rectangle ($(mid)!(z2.north)!(mid) + (-3pt,-1cm)$);
-
-    \draw[draw=none,fill=black!05,rounded corners=5pt] ($(mid)!(i2.north east)!(mid) + (3pt,1cm)$) 
-      rectangle ($(value.east)!(z2.north)!(value.east) + (.25cm,-1cm)$);
-
-  \end{pgfonlayer}
-
-\end{tikzpicture}}
-%\centerline{\includegraphics[width=\columnwidth]{./figs/method-diagram.png}}
+\end{tikzpicture}
 \caption{Self-predictice loss}
@@
 \end{figure}
+
+\end{document}
 

### 2024-02-01 paper/main.tex

--- 2024-02-01 paper/main.tex
@@
 \begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/main_plot.pdf}}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/baselines_comparison.pdf}}
 \caption{\textbf{Normalizing the latent-space improves sample efficiency} \our is significanly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
@@
 
-% \section{Full results}
+\section{Full results}
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/reconstruction-loss-ablation.pdf}}
+\caption{\textbf{Reconstruction loss has detrimental impact}.
+\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:multi-task-pretraining}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
 % \begin{figure*}[ht]

### 2024-02-01 paper/main.tex
Fig. 1 start

@@
   % Styles (this tunes colours etc.)
-  \tikzstyle{mynode}=[fill=white,draw=black!80,rounded corners=1pt,font=\tiny,inner sep=0]
-  \tikzstyle{trap}=[mynode,trapezium,minimum width=1.5em, minimum height=2em,
-                    trapezium left angle=-70, trapezium right angle=-70]
+  \tikzstyle{mynode}=[fill=white,draw=black!80,rounded corners=1pt,font=\tiny,inner sep=0,align=center]
+  \tikzstyle{trap}=[mynode,trapezium,text width=6.5em, minimum height=2em,
+                    trapezium left angle=-70, trapezium right angle=-70,inner sep=2pt]
   \tikzstyle{blob}=[mynode,circle,minimum width=2.4em,minimum height=2.4em]
@@
 
-  \node[trap] (e0) at (0\nodedist,\nodedistv) {$e_\phi$};
-  \node[trap] (e1) at (1\nodedist,\nodedistv) {$e_{\bar\phi}$};
-  \node[trap] (e2) at (2\nodedist,\nodedistv) {$e_{\bar\phi}$};
+  \node[trap] (e0) at (0\nodedist,\nodedistv) {Encoder \\ $e_\phi$};
+  \node[trap] (e1) at (1\nodedist,\nodedistv) {Momentum enc.\\$e_{\bar\phi}$};
+  \node[trap] (e2) at (2\nodedist,\nodedistv) {Momentum enc.\\$e_{\bar\phi}$};
 
@@
 
-  \node[mynode,above of=e0] (i0) {\includegraphics[width=4.7em]{figs/humanoid-1}};
-  \node[mynode,above of=e1] (i1) {\includegraphics[width=4.7em]{figs/humanoid-2}};
-  \node[mynode,above of=e2] (i2) {\includegraphics[width=4.7em]{figs/humanoid-3}};
+  \node[mynode,minimum width=8em,minimum height=8em,outer sep=0,above of=e0,node distance=6em,path picture={\node at (path picture bounding box.center){\includegraphics[height=10em]{figs/humanoid-1}};}] (i0) {};
+
+  \node[mynode,minimum width=8em,minimum height=8em,outer sep=0,above of=e1,node distance=6em,path picture={\node at (path picture bounding box.center){\includegraphics[height=10em]{figs/humanoid-3}};}] (i1) {};
+
+  \node[mynode,minimum width=8em,minimum height=8em,outer sep=0,above of=e2,node distance=6em,path picture={\node at (path picture bounding box.center){\includegraphics[height=10em]{figs/humanoid-3}};}] (i2) {};
 
@@
   \node[anchor=north] (o2) at (i2.north) {\color{white}$o_{\subs{2}}$};
-
-  \node[above of=i1,align=center] (lab0) {\normalsize\bf 1.\ Learn representation};
 
@@
 
-\end{document}
+%\end{document}
 

### 2024-02-01 paper/main.tex
Fig. 1 start

@@
 \usetikzlibrary{shapes.geometric,positioning,calc}
+
+% Arno's cube hack
+\newcommand{\cube}{%
+\resizebox{1em}{.9em}{
+\begin{tikzpicture}[baseline=-1ex,cube/.style={very thick,black},
+            grid/.style={very thin,gray},
+            axis/.style={->,black,thick}]
+ \begin{scope}[every node/.append style={yslant=-0.5},yslant=-0.5]
+ [cube/.style={very thick,black},
+            axis/.style={->,blue,thick}]
+   \shade[right color=gray!10, left color=black!50] (0,0) rectangle +(3,3);
+   \node at (0.5,2.5) {};
+   \node at (1.5,2.5) {};
+   \node at (2.5,2.5) {};
+   \node at (0.5,1.5) {};
+   \node at (1.5,1.5) {};
+   \node at (2.5,1.5) {};
+   \node at (0.5,0.5) {};
+   \node at (1.5,0.5) {};
+   \node at (2.5,0.5) {};
+   \draw (0,0) grid (3,3);
+ \end{scope}
+
+ \begin{scope}[every node/.append style={yslant=0.5},yslant=0.5]
+   \shade[right color=gray!70,left color=gray!10] (3,-3) rectangle +(3,3);
+   \node at (3.5,-0.5) {};
+   \node at (4.5,-0.5) {};
+   \node at (5.5,-0.5) {};
+   \node at (3.5,-1.5) {};
+   \node at (4.5,-1.5) {};
+   \node at (5.5,-1.5) {};
+   \node at (3.5,-2.5) {};
+   \node at (4.5,-2.5) {};
+   \node at (5.5,-2.5) {};
+   \draw (3,-3) grid (6,0);
+ \end{scope}
+
+ \begin{scope}[every node/.append style={
+     yslant=0.5,xslant=-1},yslant=0.5,xslant=-1
+   ]
+   \shade[bottom color=gray!10, top color=black!80] (6,3) rectangle +(-3,-3);
+   \node at (3.5,2.5) {};
+   \node at (3.5,1.5) {};
+   \node at (3.5,0.5) {};
+   \node at (4.5,2.5) {};
+   \node at (4.5,1.5) {};
+   \node at (4.5,0.5) {};
+   \node at (5.5,2.5) {};
+   \node at (5.5,1.5) {};
+   \node at (5.5,0.5) {};
+   \draw (3,0) grid (6,3);  
+ \end{scope}
+\end{tikzpicture}}}
 
@@
   \draw[arr] (i2) -- (e2);
-  \draw[arr] (z0) -- (z1);
-  \draw[arr] (z1) -- (z2);
-  \draw[arr] (e0) -- (z0);
+  \draw[arr] (z0) -- node[below,font=\tiny] {Dynamics} (z1);
+  \draw[arr] (z1) -- node[below,font=\tiny] {Dynamics} (z2);
+  \draw[arr] (e0) -- node[above,font=\tiny,rotate=90,yshift=6pt] {Normalization} (z0);
   \draw[arr] (e1) -- (zh1);
@@
 
+  % Add normalization
+  \node[circle,fill=white,inner sep=0] at ($(e0)!.5!(z0)$) {\cube};
+  \node[circle,fill=white,inner sep=0] at ($(e1)!.5!(zh1)$) {\cube};  
+  \node[circle,fill=white,inner sep=0] at ($(e2)!.5!(zh2)$) {\cube};
+
 \end{tikzpicture}
-\caption{Self-predictice loss}
+\caption{Self-predictice loss We denote our normalization scheme by \cube.}
 \label{diagram-self-predictive}

### 2024-02-01 paper/main.tex
Fig 1

@@
 
+  % Minimization
+  \node[blue!50,below of=z1,align=center,font=\scriptsize,xshift=4em] (min) {minimize: $\|\bar\vz_{\bullet}-\hat\vz_{\bullet}\|$};
+  \draw[blue!50,line width=1pt,shorten <=1mm,shorten >=1mm,->] (min) to[bend right=30] ($(z1)!.5!(zh1)$);
+  \draw[blue!50,line width=1pt,shorten <=1mm,shorten >=1mm,->] (min) to[bend left=30] ($(z2)!.5!(zh2)$);
+
 \end{tikzpicture}

### 2024-02-01 paper/figs/plotting/baseline-comparison/fetch_data.py
updated

@@ -10,11 +10,15 @@ ddpg_data = {
         # "walker-walk": ["4wydf2sz", "n7l8xac6", "l7nqlsid", "yrwwmyq2", "bpdxi4jw"],
         # "quadruped-walk": ["6gf2r90u", "uvq2ctjr", "4x1r1nqg", "lji8f92s", "eiafkell"],
         #### Project=False ####
+        # GOOD performance d=128
+        "quadruped-walk": ["8zgkcuhs", "egk6qxco", "oh1xhyqt", "hm3ihqzo", "p4slb4q9"],
+        "cheetah-run": ["l4c3phty", "96xm0zx5", "6tul0lmk", "y7c9k82p", "zx6ykgj9"],
+        # POOR performance d=512
         "acrobot-swingup": ["c6qmc8yx", "spwng8k4", "wgurvdan", "4xfe6bez", "95l445ci"],
-        "cheetah-run": ["7kg8xyj8", "gujw41jb", "r9cu6yvh", "lh1kdlht", "8gymqblt"],
+        # "cheetah-run": ["7kg8xyj8", "gujw41jb", "r9cu6yvh", "lh1kdlht", "8gymqblt"],
         # "fish-swim": ["izb29vhj", "xlgm5s8v", "xc1nf2ss", "ezhafy1m", "5tw9ini4"],
         "walker-walk": ["9xh1l0n1", "9ujhx4vn", "pojcknc3", "2xehepzn", "fvzzgcw2"],
-        "quadruped-walk": ["17h2zwdq", "s6i2xtt3", "fu8wchfa", "tzwhqglg", "7n7tivxv"],
+        # "quadruped-walk": ["17h2zwdq", "s6i2xtt3", "fu8wchfa", "tzwhqglg", "7n7tivxv"],
         "humanoid-walk": [
             "wur1u276",
             "l8da5827",
@@ -51,6 +55,8 @@ ddpg_data = {
             "kprj5q9u",
             "r232jms2",
             # "aox6spn2", # crashed
+            "5ubkon5e",
+            "65wmol3k",
         ],  # project=False
         # "dog-run": [
         #     "7hty9mrz",
@@ -135,6 +141,7 @@ def fetch_results(run_path, run_name_list, keys, agent_name):
         history["env"] = env_name
         history["seed"] = seed
         history["agent"] = agent_name
+        # history["agent"] = agent_name + f" d={wandb_run.config['agent']['latent_dim']}"
[…]

### 2024-02-01 paper/figs/plotting/baseline-comparison/make_plot.py

@@ -20,6 +20,9 @@ COLORS = {
     "REDQ": "#984ea3",
     "TD-MPC": "#ff7f00",
     "iFSQ-RL": "#e41a1c",
+    # "iFSQ-RL d=1024": "magenta",
+    # "iFSQ-RL d=512": "grey",
+    # "iFSQ-RL d=128": "#e41a1c",
 }
 # %%
 main_envs = [
@@ -44,6 +47,7 @@ def plot(df, key="episode_reward"):
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
+    # breakpoint()
     # df["env_step"] = df["env_step"] / 1000
     for idx, env in enumerate(main_envs):
         data = df[df["env"] == env]
@@ -54,7 +58,7 @@ def plot(df, key="episode_reward"):
         ax = axs[row, col]
         hue_order = data.agent.unique()
 
-        if idx == 4:
+        if idx == 3:
             sns.lineplot(
                 # x=int("env_step" / 1000),
                 x="env_step",
@@ -64,7 +68,7 @@ def plot(df, key="episode_reward"):
                 data=data,
                 errorbar=("ci", 95),
                 hue="agent",
-                hue_order=hue_order,
+                # hue_order=hue_order,
                 palette=COLORS,
                 legend="auto",
                 ax=ax,
@@ -78,7 +82,7 @@ def plot(df, key="episode_reward"):
                 data=data,
                 errorbar=("ci", 95),
                 hue="agent",
-                hue_order=hue_order,
+                # hue_order=hue_order,
                 palette=COLORS,
                 legend=False,
                 ax=ax,
@@ -119,7 +123,7 @@ df = [
     pd.read_csv(f"{data_path}/tdmpc_main.csv"),
     df_redq,
     pd.read_csv(f"{data_path}/sac_main.csv"),
-    pd.read_csv(f"{data_path}/vq_td3_main.csv"),
+    pd.read_csv(f"{data_path}/ifsq-rl.csv"),
 ]
 plot(pd.concat(df))
 

### 2024-02-01 paper/figs/plotting/norm-ablation/fetch_data.py

@@ -177,6 +177,20 @@ ddpg_data = {
             "lh1kdlht",  # seed=4
             "8gymqblt",  # seed=5
         ],
+        "acrobot-swingup": [
+            # no norm
+            "vkbkcrjm",
+            "nx50by5c",
+            "4iy9ahwx",
+            "frc6lm7a",
+            "e730qvy7",
+            # norm
+            "c6qmc8yx",
+            "spwng8k4",
+            "wgurvdan",
+            "4xfe6bez",
+            "95l445ci",
+        ],
         # 'humanoid-walk': ['2xa4zh7i', 'u01r5cd8', '2dm45ezi', '2ywqbphi', '33kix5gc', ],
         # "dog-walk": ["gtrhfxn3"],
         # "dog-run": ["ukzsy2t5"],

### 2024-02-01 paper/figs/plotting/norm-ablation/make_plot.py

@@ -28,15 +28,15 @@ COLORS = {
 }
 # %%
 main_envs = [
-    # "acrobot-swingup",
-    "cheetah-run",
+    "acrobot-swingup",
+    # "cheetah-run",
     # "walker-walk",
     "walker-run",
     # "hopper-stand",
     # "fish-swim",
-    "quadruped-run",
+    # "quadruped-run",
     "humanoid-walk",
-    # "humanoid-run",
+    "humanoid-run",
     # "dog-walk",
     # "dog-run",
 ]

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/fetch_data.py

@@ -40,6 +40,13 @@ ddpg_data = {
             "j8apohl8",  # seed=4
             "5w8halqc",  # seed=5
         ],
+        "quadruped-run-d=128": [
+            "csjhne6n",
+            "bm5kosik",
+            "kum2pe4f",
+            "s9p9y44s",
+            "4gsq88ju",
+        ],
         "quadruped-run": [
             ##### Project=True
             # # reconstruction
@@ -62,11 +69,11 @@ ddpg_data = {
             "sats7q1b",  # seed=4
             "foax0pto",  # seed=5
             # no reconstruction
-            "g16tux3h",  # seed=1
-            "82tqwdmq",  # seed=2
-            "msec1uul",  # seed=3
-            "jd6nsd17",  # seed=4
-            "i09us6qm",  # seed=5
+            # "g16tux3h",  # seed=1
+            # "82tqwdmq",  # seed=2
+            # "msec1uul",  # seed=3
+            # "jd6nsd17",  # seed=4
+            # "i09us6qm",  # seed=5
         ],
         "walker-run": [
             ##### Project=True
@@ -96,6 +103,13 @@ ddpg_data = {
             "w65qql3j",  # seed=4
             "wkge5rid",  # seed=5
         ],
+        "cheetah-run-d128": [
+            "l4c3phty",
+            "96xm0zx5",
+            "6tul0lmk",
+            "y7c9k82p",
+            "zx6ykgj9",
+        ],
         "cheetah-run": [
             ##### Project=True
             # # reconstruction
@@ -118,11 +132,25 @@ ddpg_data = {
             "ndtfn06e",  # seed=5
             # no reconstruction
             ##### Project=False
-            "7kg8xyj8",  # seed=1
-            "gujw41jb",  # seed=2
-            "r9cu6yvh",  # seed=3
-            "lh1kdlht",  # seed=4
-            "8gymqblt",  # seed=5
+            # "7kg8xyj8",  # seed=1
+            # "gujw41jb",  # seed=2
+            # "r9cu6yvh",  # seed=3
+            # "lh1kdlht",  # seed=4
+            # "8gymqblt",  # seed=5
+        ],
+        "acrobot-swingup": [
+            # reconstruction
+            "6sfcsm49",
+            "4zz2hrn9",
+            "ab4rl6yl",
+            "4ewms37u",
+            "1q4nt70m",
+            # no reconstruction
+            "c6qmc8yx",
+            "spwng8k4",
[…]

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/make_plot.py

@@ -25,10 +25,10 @@ COLORS = {
 }
 # %%
 main_envs = [
-    # "acrobot-swingup",
+    "acrobot-swingup",
     # "cheetah-run",
     # "walker-walk",
-    "walker-run",
+    # "walker-run",
     "hopper-stand",
     # "fish-swim",
     "quadruped-run",

### 2024-02-01 paper/figs/plotting/task-agnostic-ablation/fetch_data.py
Task agnostic pre-training and reward ablation

@@ -1,0 +1,167 @@
+# %%
+ddpg_path = "kallekku/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        "humanoid-run": [
+            # Pre-training
+            "h0e96eeo",
+            "6tykwpqn",
+            "w446n5he",
+            "fnzw5jyj",
+            "ugcwkueb",
+            # No pre-training:
+            "ek6t7mtb",
+            "febl1741",
+            "i2926y8u",
+            "jntea203",
+            "kjw08gye",
+        ],
+        "hopper-hop": [
+            # Pre-training
+            "dblmpj4q",
+            "hx4frieq",
+            "j13wyns5",
+            "eonukhi9",
+            "s9br2god",
+            # No pre-training
+            "4ax1gct6",
+            "f5d66x17",
+            "jp26qyvm",
+            "n3cb46el",
+            "z4dideh5"
+        ],
+        "quadruped-run": [
+            # Pre-training"
+            "0zviwszy",
+            "dr5wqqs6",
+            "0tb5a3p1",
+            "jchvk2kd",
+            "2l4wphqg",
+            # No pre-training
+            "1ottm721",
+            "67objwkg",
+            "a1rml1p2",
+            "my89a2ui",
+            "0bdbllyg",
+            # Task-specific pre-training
+            "uykvengt",
+            "e6u4u0et",
+            "phlr610t",
+            "32c5rcc6",
+            "2hyd7evc",
+        ],
+        "walker-run": [
+            # Pre-training"
+            "cua4ibb4",
+            "ya38xqbp",
+            "1cp2xqb6",
+            "vokt0o8m",
+            "qwy97zej",
+            # No pre-training
+            "0y5whgm0",
+            "5hhavbpu",
+            "zwxgu2fr",
+            "cbp72vz9",
+            "oi0gnnd1",
+            # Task-specific pre-training
+            "oxisl0xv",
+            "wv0a1i6l",
+            "2mstpcgd",
+            "olnugqts",
+            "1iegxig2",
+        ],
+    },
+}
+
+# %%
+# Plot wandb
+import numpy as np
+from typing import Dict
+
+import matplotlib.pyplot as plt
+
+# plt.style.use("seaborn")
+plt.style.use("seaborn-v0_8")
+
+import wandb
+
+api = wandb.Api(timeout=19)
+import pandas as pd
+
+# pet-pytorch
+titles = [
+    "env_step",
[…]

### 2024-02-01 paper/figs/plotting/task-agnostic-ablation/make_plot.py

@@ -1,0 +1,144 @@
+# %%
+import pandas as pd
+import numpy as np
+
+
+import matplotlib.pyplot as plt
+from matplotlib import cm
+
+# plt.style.use('bmh')
+plt.style.use("seaborn-v0_8-whitegrid")
+import seaborn as sns
+
+plt.rcParams["figure.dpi"] = 400
+plt.rcParams["font.size"] = 13
+plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["legend.loc"] = "lower right"
+COLORS = {
+    # "TCRL": "#e41a1c",
+    # "SAC": "#377eb8",
+    # "REDQ": "#984ea3",
+    # "TD-MPC": "#ff7f00",
+    # "VQ-TD3": "magenta",
+    "iFSQ-RL": "#e41a1c",
+    "iFSQ-RL-spec-trf": "#377eb8",
+    "iFSQ-RL-agn-trf": "#ff7f00",
+}
+# %%
+main_envs = [
+    # "acrobot-swingup",
+    # "cheetah-run",
+    # "walker-walk",
+    "walker-run",
+    # "hopper-stand",
+    # "fish-swim",
+    "quadruped-run",
+    # "humanoid-walk",
+    "humanoid-run",
+    "hopper-hop",
+    # "dog-walk",
+    # "dog-run",
+]
+
+# pet-pytorch
+titles = [
+    "env_step",
+    "episode",
+    "episode_reward",
+    "time",
+]
+keys = [
+    "eval/.env_step",
+    "eval/.episode",
+    "eval/.episodic_return",
+    "eval/.elapsed_time",
+]
+
+rename = {a: b for a, b in zip(keys, titles)}
+
+
+def plot(df, key="episode_reward"):
+    # breakpoint()
+    envs = np.sort(df.env.unique())
+    ncol = 4
+    # assert envs.shape[0] % ncol == 0
+    # nrow = len(main_envs) // ncol
+    nrow = 1
+    # nrow = envs.shape[0] // ncol
+
+    fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+
+    df = df.rename(columns=rename)
+
+    for idx, env in enumerate(main_envs):
+        data = df[df["env"] == env]
+        row = idx // ncol
+        col = idx % ncol
+        # ax = axs[row, col]
+        ax = axs[col]
+        # hue_order = data.agent.unique()
+        hue_order = data.name.unique()
[…]

### 2024-02-01 paper/figs/plotting/baseline-comparison/fetch_data.py
updated

@@ -12,8 +12,9 @@ ddpg_data = {
         #### Project=False ####
         # GOOD performance d=128
         "quadruped-walk": ["8zgkcuhs", "egk6qxco", "oh1xhyqt", "hm3ihqzo", "p4slb4q9"],
-        "cheetah-run": ["l4c3phty", "96xm0zx5", "6tul0lmk", "y7c9k82p", "zx6ykgj9"],
-        # POOR performance d=512
+        "cheetah-run": ["8x29ibzc", "9k7936z8", "e7afvv6h", "81pbgf97", "qfqv9wz3"],
+        # "cheetah-run": ["l4c3phty", "96xm0zx5", "6tul0lmk", "y7c9k82p", "zx6ykgj9"],
+        # POOR performance d=512 project=False
         "acrobot-swingup": ["c6qmc8yx", "spwng8k4", "wgurvdan", "4xfe6bez", "95l445ci"],
         # "cheetah-run": ["7kg8xyj8", "gujw41jb", "r9cu6yvh", "lh1kdlht", "8gymqblt"],
         # "fish-swim": ["izb29vhj", "xlgm5s8v", "xc1nf2ss", "ezhafy1m", "5tw9ini4"],

### 2024-02-01 paper/figs/plotting/norm-ablation/fetch_data.py

@@ -13,6 +13,62 @@ ddpg_data = {
         # "quadruped-walk": ["8q59zkws"],  # UTD=1
         # "cheetah-run": ["mkmvf5md"],  # UTD=4
         # : [,],
+        "quadruped-run-d=128-no-norm": [
+            "q7ia4kx5",
+            "i9b9nl8y",
+            "tvondn1a",
+            "bfbhcgre",
+            "kqvznpvy",
+        ],
+        "quadruped-run-d=128-norm": [
+            "csjhne6n",
+            "bm5kosik",
+            "kum2pe4f",
+            "s9p9y44s",
+            "4gsq88ju",
+        ],
+        "cheetah-run-d128-no-norm": [
+            "5j2it0tc",
+            "v0fz1014",
+            "v1i6mftd",
+            "yfkyy6rt",
+            "txtebz1w",
+        ],
+        "cheetah-run-d128-norm": [
+            "l4c3phty",
+            "96xm0zx5",
+            "6tul0lmk",
+            "y7c9k82p",
+            "zx6ykgj9",
+        ],
+        "walker-run-d=128-no-norm": [
+            "tkja6y1j",
+            "ogmwwcwv",
+            "i27uobpy",
+            "p59ieaw9",
+            "c5a2wliy",
+        ],
+        "walker-run-d=128": [
+            "kb1gk7hm",
+            "egjvexs4",
+            "lwnz0cf1",
+            "zfsg29ej",
+            "5k9okav3",
+        ],
+        "hopper-stand-d=64-no-norm": [
+            "0ymb3imw",
+            "5fibx0fg",
+            "i8ou5zx4",
+            "bijgsw6i",
+            "02uvlbjb",
+        ],
+        "hopper-stand-d=64": [
+            "b978u8ed",
+            "77bswwl0",
+            "rkdmf32p",
+            "ucvgujrp",
+            "zqv19l3p",
+        ],
         "humanoid-run": [
             # no-norm d=1024 project=False
             "uuyjutba",
@@ -41,142 +97,142 @@ ddpg_data = {
             "krctc38z",
             "anstkvup",
         ],  # project=False
-        "hopper-stand": [
-            ###### Project=True
-            # # no-norm, d=50
-            # "25m4987b",  # seed=1
-            # "cet5h22j",  # seed=2
-            # "8o4v1gz1",  # seed=3
[…]

### 2024-02-01 paper/figs/plotting/norm-ablation/make_plot.py

@@ -29,14 +29,14 @@ COLORS = {
 # %%
 main_envs = [
     "acrobot-swingup",
-    # "cheetah-run",
+    "cheetah-run",
     # "walker-walk",
     "walker-run",
     # "hopper-stand",
     # "fish-swim",
-    # "quadruped-run",
-    "humanoid-walk",
-    "humanoid-run",
+    "quadruped-run",
+    # "humanoid-walk",
+    # "humanoid-run",
     # "dog-walk",
     # "dog-run",
 ]

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/fetch_data.py

@@ -13,32 +13,12 @@ ddpg_data = {
         # "quadruped-walk": ["8q59zkws"],  # UTD=1
         # "cheetah-run": ["mkmvf5md"],  # UTD=4
         # : [,],
-        "hopper-stand": [
-            ##### Project=True
-            # # reconstruction
-            # "25m4987b",  # seed=1
-            # "cet5h22j",  # seed=2
-            # "8o4v1gz1",  # seed=3
-            # "udj85ac1",  # seed=4
-            # "7k9kjfll",  # seed=5
-            # # no reconstruction
-            # "1wmalfbs",  # seed=1
-            # "rj3vhnbq",  # seed=2
-            # "j4iwbxq2",  # seed=3
-            # "jo9td3sx",  # seed=4
-            # "71nsurzh",  # seed=5
-            # reconstruction
-            "94skam71",  # seed=1
-            "hss7p9jr",  # seed=2
-            "oe1us2qh",  # seed=3
-            "nijlewdk",  # seed=4
-            "4m35m7d4",  # seed=5
-            # no reconstruction
-            "3rmiy5he",  # seed=1
-            "39j1gbp2",  # seed=2
-            "t4hf7g18",  # seed=3
-            "j8apohl8",  # seed=4
-            "5w8halqc",  # seed=5
+        "quadruped-run-d=128-rec": [
+            "qgadc67o",
+            "8cbgk6x9",
+            "kg64um8c",
+            "qr1oe9y9",
+            "dfl6e0j3",
         ],
         "quadruped-run-d=128": [
             "csjhne6n",
@@ -47,61 +27,12 @@ ddpg_data = {
             "s9p9y44s",
             "4gsq88ju",
         ],
-        "quadruped-run": [
-            ##### Project=True
-            # # reconstruction
-            # "rw26pb1l",  # seed=1
-            # "jm549p6m",  # seed=2
-            # "7ll3mp6d",  # seed=3
-            # "ypgoji3k",  # seed=4
-            # "ll54kaha",  # seed=5
-            # # no reconstruction
-            # "w625abb9",  # seed=1
-            # "w1ssoh9i",  # seed=2
-            # "wtobbh3o",  # seed=3
-            # "4sj3p04r",  # seed=4
-            # "wtobbh3o",  # seed=5
-            ##### Project=False
-            # reconstruction
-            "3yyb8ibb",  # seed=1
-            "v0e4moj8",  # seed=2
[…]

### 2024-02-01 paper/figs/plotting/baseline-comparison/fetch_data.py
update figs

@@ -13,12 +13,13 @@ ddpg_data = {
         # GOOD performance d=128
         "quadruped-walk": ["8zgkcuhs", "egk6qxco", "oh1xhyqt", "hm3ihqzo", "p4slb4q9"],
         "cheetah-run": ["8x29ibzc", "9k7936z8", "e7afvv6h", "81pbgf97", "qfqv9wz3"],
+        "walker-walk": ["8z9o8vp6", "roz89u0n", "h1by1w6s", "zfu00w22", "jvw4myex"],
         # "cheetah-run": ["l4c3phty", "96xm0zx5", "6tul0lmk", "y7c9k82p", "zx6ykgj9"],
         # POOR performance d=512 project=False
         "acrobot-swingup": ["c6qmc8yx", "spwng8k4", "wgurvdan", "4xfe6bez", "95l445ci"],
         # "cheetah-run": ["7kg8xyj8", "gujw41jb", "r9cu6yvh", "lh1kdlht", "8gymqblt"],
         # "fish-swim": ["izb29vhj", "xlgm5s8v", "xc1nf2ss", "ezhafy1m", "5tw9ini4"],
-        "walker-walk": ["9xh1l0n1", "9ujhx4vn", "pojcknc3", "2xehepzn", "fvzzgcw2"],
+        # "walker-walk": ["9xh1l0n1", "9ujhx4vn", "pojcknc3", "2xehepzn", "fvzzgcw2"],
         # "quadruped-walk": ["17h2zwdq", "s6i2xtt3", "fu8wchfa", "tzwhqglg", "7n7tivxv"],
         "humanoid-walk": [
             "wur1u276",
@@ -36,10 +37,10 @@ ddpg_data = {
         # ],  # project=True
         # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
         "dog-walk": [
-            "stawiad8",
+            # "stawiad8",
             "iv8p110s",
-            "k454j543",
-            "nihkvll9",
+            # "k454j543",
+            # "nihkvll9",
             "jpkbw5h7",
         ],  # project=False
         # "dog-walk": [

### 2024-02-01 paper/figs/plotting/norm-ablation/make_plot.py

@@ -32,13 +32,13 @@ main_envs = [
     "cheetah-run",
     # "walker-walk",
     "walker-run",
-    # "hopper-stand",
+    "hopper-stand",
     # "fish-swim",
     "quadruped-run",
-    # "humanoid-walk",
-    # "humanoid-run",
+    "humanoid-walk",
+    "humanoid-run",
     # "dog-walk",
-    # "dog-run",
+    "dog-run",
 ]
 
 # pet-pytorch
@@ -63,8 +63,8 @@ def plot(df, key="episode_reward"):
     envs = np.sort(df.env.unique())
     ncol = 4
     # assert envs.shape[0] % ncol == 0
-    # nrow = len(main_envs) // ncol
-    nrow = 1
+    nrow = len(main_envs) // ncol
+    # nrow = 1
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
@@ -75,8 +75,8 @@ def plot(df, key="episode_reward"):
         data = df[df["env"] == env]
         row = idx // ncol
         col = idx % ncol
-        # ax = axs[row, col]
-        ax = axs[col]
+        ax = axs[row, col]
+        # ax = axs[col]
         # hue_order = data.agent.unique()
         hue_order = data.name.unique()
         # hue_order = data.utd_ratio.unique()

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/make_plot.py

@@ -26,13 +26,13 @@ COLORS = {
 # %%
 main_envs = [
     "acrobot-swingup",
-    # "cheetah-run",
+    "cheetah-run",
     # "walker-walk",
-    # "walker-run",
+    "walker-run",
     "hopper-stand",
-    # "fish-swim",
+    "fish-swim",
     "quadruped-run",
-    # "humanoid-run",
+    "humanoid-run",
     # "humanoid-walk",
     "dog-walk",
     # "dog-run",
@@ -60,8 +60,8 @@ def plot(df, key="episode_reward"):
     envs = np.sort(df.env.unique())
     ncol = 4
     # assert envs.shape[0] % ncol == 0
-    # nrow = len(main_envs) // ncol
-    nrow = 1
+    nrow = len(main_envs) // ncol
+    # nrow = 1
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
@@ -72,8 +72,8 @@ def plot(df, key="episode_reward"):
         data = df[df["env"] == env]
         row = idx // ncol
         col = idx % ncol
-        # ax = axs[row, col]
-        ax = axs[col]
+        ax = axs[row, col]
+        # ax = axs[col]
         # hue_order = data.agent.unique()
         hue_order = data.name.unique()
         # hue_order = data.utd_ratio.unique()

### 2024-02-01 paper/main.tex
add multi task figure
updated intro and experiments

@@
 
-Reinforcement learning (RL) has potential...
+Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+However, applying RL in real-world environments is challenging. The problem is two-fold,
+{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
 
-However, applying RL in real-world environments with high dimensional/complex observation spaces is challenging.
+% Given such a latent space,
+% Whilst these latent states are often lower dimensional this is not always
+Many approaches have been proposed to improve sample efficiency in deep RL.
+As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
+In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+In particular, methods for compressing observations into latent states.%, aka representation learning.
 
-Compressing observations in latent state space.
-Hand crafted features
+Learning representations for RL has been investigated for decades
+\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
+However, these approaches are usually limited to simple environments.
+Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+\cite{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
+attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
+Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
-Learning features with representation learning.
-These methods typically learn task-specific representations
-as they train the representation to predict rewards and/or values
-\citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
-
-For example, \citet{zhaoSimplifiedTemporalConsistency2023} use a latent-state consistency loss based upon the Cosine
-similarity loss plus the MSE of reward predictions in the latent space.
-The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
+To the best of our knowledge, all prior works learn representations with objectives that depend on
+predicting the reward and/or value in the latent space \citet{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+% use a latent-state consistency loss based upon the Cosine
+% For example, \citet{zhaoSimplifiedTemporalConsistency2023} use a latent-state consistency loss based upon the Cosine
+% similarity loss plus the MSE of reward predictions in the latent space.
+% The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
+As such, these methods learn task-specific representations.
+% \citet{geladaDeepMDPLearningContinuous2019,zhangLearningInvariantRepresentations2020,rezaei-shoshtariContinuousMDPHomomorphisms2022}
+use SSL losses based on bisimulation metrics \citep{larsenBisimulationProbabilisticTesting1989}.
+As such, their encoders group states into clusters whilst preserving some properties
+(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
+These methods use latent-space dynamic models to learn latent representations which encode only task-relevant information.
+% Interestingly, the dynamics models are only used for representation learning and are not used for model-based planning.
+DeepMDP \citep{geladaDeepMDPLearningContinuous2019} combines its representation learning with
+distributional Q learning  \citep[C51,][]{bellemareDistributionalPerspectiveReinforcement2017},
+whilst DBC \citep{zhangLearningInvariantRepresentations2020} combine its method with SAC.
 
@@
 \label{sec:experiments}
+In this section, we evaluate \our in a wide variety of continuous control tasks from the DeepMind Control (DMC) suite.
+We aim to answer the following questions:
+\begin{itemize}
+    \item How does \our compare to state-of-the-art model-free and model-based algorithms?
+    \item Is learning a representation with only latent-state consistency really better than including reward predictions and value predictions?
+    \item Can \our's task-agnostic representation aid task-adaptation with its pre-trained representation?
+\end{itemize}
 
@@
 \our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:multi-task-pretraining}
+\label{fig:normalization-ablation}
 \end{center}
@@
 \begin{center}
-\centerline{\includegraphics[width=0.3\textwidth]{example-image}}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/reward-ablation.pdf}}
+\caption{\textbf{Reward prediction is not necessary for representation learning}.
+\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:reward-ablation}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\textwidth]{./figs/task-agnostic-ablation.pdf}}
 \caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
@@
 
-\begin{itemize}
-    \item Show normalised latent leads to higher sample efficiency. Show results for NTC-TD3/TC-TD3/TD3/SAC/TCRL/TD-MPC with UTD=1.
-    \item Show normalised latent works with higher UTD. Compare NTC-TD3/TC-TD3/TCRL.
-    \item Show task-agnostic representation is good for incremental task learning. Compare NTC-TD3/TCRL-ours/TCRL
-\end{itemize}
+% \begin{itemize}
+%     \item Show normalised latent leads to higher sample efficiency. Show results for NTC-TD3/TC-TD3/TD3/SAC/TCRL/TD-MPC with UTD=1.
+%     \item Show normalised latent works with higher UTD. Compare NTC-TD3/TC-TD3/TCRL.
+%     \item Show task-agnostic representation is good for incremental task learning. Compare NTC-TD3/TCRL-ours/TCRL
+% \end{itemize}
 
-\text{Ablations}
+\subsection{Ablations}
 
-\begin{itemize}
-  \item Show un-normalised isn't as good.
-  \item Compare different size latent spaces? Different levels $[8,5]$ vs $[8,6,5]$ vs $[8,5,5,5]$ vs $[7,5,5,5,5]$
-  \item Target encoder vs online encoder for mapping states to latents for TD3.
-  \item Show cosine similarity slows training.
-  \item Show adding reconstruction loss hurts performance.
-\end{itemize}
+\textbf{Reward predict is not necessary for representation learning}
+Next, we test if using reward prediction aids representation learning.
+\cref{fig:reward-ablation} shows that
+
+\textbf{Reconstruction loss has detrimental impact}
+Recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
+has shown that incorporating a reconstruction term into the representation loss can have a negative impact on performance.
+To provide a thorough analysis of \our, we include results where we add a reconstruction term to our
+representation loss \cref{eq:rep-loss}.
+We perform reconstruction at each time step in the horizon.
+\cref{fig:multi-task-pretraining} shows that in no environments does reconstruction aid learning.
+Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on
+overall performance.
+
+% \begin{itemize}
+%   \item Show un-normalised isn't as good.
+%   \item Compare different size latent spaces? Different levels $[8,5]$ vs $[8,6,5]$ vs $[8,5,5,5]$ vs $[7,5,5,5,5]$
+%   \item Target encoder vs online encoder for mapping states to latents for TD3.
+%   \item Show cosine similarity slows training.
+%   \item Show adding reconstruction loss hurts performance.
+% \end{itemize}
 

### 2024-02-01 paper/main.tex
updated

@@
 
-To the best of our knowledge, all prior works learn representations with objectives that depend on
-predicting the reward and/or value in the latent space \citet{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
-% use a latent-state consistency loss based upon the Cosine
-% For example, \citet{zhaoSimplifiedTemporalConsistency2023} use a latent-state consistency loss based upon the Cosine
-% similarity loss plus the MSE of reward predictions in the latent space.
-% The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
-As such, these methods learn task-specific representations.
-% \citet{geladaDeepMDPLearningContinuous2019,zhangLearningInvariantRepresentations2020,rezaei-shoshtariContinuousMDPHomomorphisms2022}
-use SSL losses based on bisimulation metrics \citep{larsenBisimulationProbabilisticTesting1989}.
-As such, their encoders group states into clusters whilst preserving some properties
-(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
-These methods use latent-space dynamic models to learn latent representations which encode only task-relevant information.
-% Interestingly, the dynamics models are only used for representation learning and are not used for model-based planning.
-DeepMDP \citep{geladaDeepMDPLearningContinuous2019} combines its representation learning with
-distributional Q learning  \citep[C51,][]{bellemareDistributionalPerspectiveReinforcement2017},
-whilst DBC \citep{zhangLearningInvariantRepresentations2020} combine its method with SAC.
-
-In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method.
-Normalization based upon Finite Scalar Quantization
-Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} prevents representation collapse,
-and {\em (iii)} is task-agnostic.
-
-% In recent years, it has become common to
-
 \begin{figure}[t!]
@@
   % Minimization
-  \node[blue!50,below of=z1,align=center,font=\scriptsize,xshift=4em] (min) {minimize: $\|\bar\vz_{\bullet}-\hat\vz_{\bullet}\|$};
+  \node[blue!50,below of=z1,align=center,font=\scriptsize,xshift=4em] (min) {minimize: $\|\bar\vz_{\bullet}-\hat\vz_{\bullet}\|^{2}_{2}$};
   \draw[blue!50,line width=1pt,shorten <=1mm,shorten >=1mm,->] (min) to[bend right=30] ($(z1)!.5!(zh1)$);
@@
 \end{tikzpicture}
-\caption{Self-predictice loss We denote our normalization scheme by \cube.}
+\caption{\textbf{\our's model architecture} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
+We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
+To this end, our normalization results in an implicit codebook whose size we can control.
+We denote our normalization scheme by \cube.}
 \label{diagram-self-predictive}
@@
 \end{figure}
+
+To the best of our knowledge, all prior works learn representations with objectives that depend on
+predicting the reward and/or value in the latent space \citet{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+% use a latent-state consistency loss based upon the Cosine
+% For example, \citet{zhaoSimplifiedTemporalConsistency2023} use a latent-state consistency loss based upon the Cosine
+% similarity loss plus the MSE of reward predictions in the latent space.
+% The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
+As such, these methods learn task-specific representations.
+% \citet{geladaDeepMDPLearningContinuous2019,zhangLearningInvariantRepresentations2020,rezaei-shoshtariContinuousMDPHomomorphisms2022}
+use SSL losses based on bisimulation metrics \citep{larsenBisimulationProbabilisticTesting1989}.
+As such, their encoders group states into clusters whilst preserving some properties
+(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
+These methods use latent-space dynamic models to learn latent representations which encode only task-relevant information.
+% Interestingly, the dynamics models are only used for representation learning and are not used for model-based planning.
+DeepMDP \citep{geladaDeepMDPLearningContinuous2019} combines its representation learning with
+distributional Q learning  \citep[C51,][]{bellemareDistributionalPerspectiveReinforcement2017},
+whilst DBC \citep{zhangLearningInvariantRepresentations2020} combine its method with SAC.
+
+In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method.
+Normalization based upon Finite Scalar Quantization
+Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} prevents representation collapse,
+and {\em (iii)} is task-agnostic.
+
+% In recent years, it has become common to
 

### 2024-02-01 paper/main.tex
finished draft of intro

@@
 Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\cite{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
 More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
@@
 We denote our normalization scheme by \cube.}
-\label{diagram-self-predictive}
+\label{fig:diagram-self-predictive}
 \vskip -0.2in
@@
 To the best of our knowledge, all prior works learn representations with objectives that depend on
-predicting the reward and/or value in the latent space \citet{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+predicting the reward and/or value in the latent space
+\citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
 % use a latent-state consistency loss based upon the Cosine
@@
 As such, these methods learn task-specific representations.
-% \citet{geladaDeepMDPLearningContinuous2019,zhangLearningInvariantRepresentations2020,rezaei-shoshtariContinuousMDPHomomorphisms2022}
-use SSL losses based on bisimulation metrics \citep{larsenBisimulationProbabilisticTesting1989}.
-As such, their encoders group states into clusters whilst preserving some properties
-(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
-These methods use latent-space dynamic models to learn latent representations which encode only task-relevant information.
-% Interestingly, the dynamics models are only used for representation learning and are not used for model-based planning.
-DeepMDP \citep{geladaDeepMDPLearningContinuous2019} combines its representation learning with
-distributional Q learning  \citep[C51,][]{bellemareDistributionalPerspectiveReinforcement2017},
-whilst DBC \citep{zhangLearningInvariantRepresentations2020} combine its method with SAC.
 
-In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method.
-Normalization based upon Finite Scalar Quantization
-Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} prevents representation collapse,
+In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method
+(we use TD3 \citep{fujimotoAddressingFunctionApproximation2018}).
+See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
+It is based soley on a self-predictive loss.
+In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
+\citep{mentzerFiniteScalarQuantization2023}.
+As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
+Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} is simple to implement,
 and {\em (iii)} is task-agnostic.
-
-% In recent years, it has become common to
-
-%\end{document}
-
-\section{Related Work}
-\label{related_work}
-
-\textbf{Sample-efficiency in deep RL}
-Many approaches have been proposed to improve sample efficiency in deep RL.
-As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
-\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-
-\textbf{Representation learning for RL}
-Learning representations for RL has been investigated for decades
-\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
-However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\cite{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
-More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
-attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
-
-\citet{yaratsImprovingSampleEfficiency2021} learn a compact representation with a deterministic autoencoder and then train a
-SAC agent \citep{haarnojaSoft2018} with the representation. They name their method SAC-AE.
-Interestingly, SAC-AE matches the performance of model-based methods like PlaNet \citep{hafnerLearning2019}
-and SLAC \citep{leeStochasticLatentActorCritic2020}, whilst being considerably simpler to implement.
-
-\textbf{Bisimulation}
-DeepMDP \citep{geladaDeepMDPLearningContinuous2019},
-deep bisimulation for control (DBC) \citep{zhangLearningInvariantRepresentations2020}
-and DHGP \citep{rezaei-shoshtariContinuousMDPHomomorphisms2022}
-% \citet{geladaDeepMDPLearningContinuous2019,zhangLearningInvariantRepresentations2020,rezaei-shoshtariContinuousMDPHomomorphisms2022}
-use SSL losses based on bisimulation metrics \citep{larsenBisimulationProbabilisticTesting1989}.
-As such, their encoders group states into clusters whilst preserving some properties
-(\eg optimal value, all values, all action values from each state) \citep{liUnifiedTheoryState2006}.
-These methods use latent-space dynamic models to learn latent representations which encode only task-relevant information.
-% Interestingly, the dynamics models are only used for representation learning and are not used for model-based planning.
-DeepMDP \citep{geladaDeepMDPLearningContinuous2019} combines its representation learning with
-distributional Q learning  \citep[C51,][]{bellemareDistributionalPerspectiveReinforcement2017},
-whilst DBC \citep{zhangLearningInvariantRepresentations2020} combine its method with SAC.
-
-% \cite{zhaoSimplifiedTemporalConsistency2023}
-% \cite{hansenTemporalDifferenceLearning2022}
-
-% \cite{higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015} use VAE \citep{kingmaAutoEncoding2014}
-
-% \citet{zhangLearningInvariantRepresentations2020} to learn a latent representation which encode only
-% task-relevant information.
-% Their method is named deep bisimulation for control (DBC) and they combine their representation learning method with SAC.
-
-\textbf{Primacy bias}
-Primacy bias in deep RL \citep{nikishinPrimacyBiasDeep2022}
-is where an agent overfits on early data and loses its ability to learn from new data,
-which arises due to the plasticity in NNs \citep{lyleUnderstandingPlasticityNeural2023} \todo{earlier citation of plasticiy in NNs??}.
-\citet{nikishinPrimacyBiasDeep2022} showed that periodically resetting the agent's parameters can help alleviate primacy bias.
-\citet{doroSampleEfficientReinforcementLearning2022} built on top of this insight and showed that resetting the
-agent's parameters not only alleviates primacy bias but also
-enables the number of parameter update's per environment interaction to be increased.
-They refer to this as the replay ratio but it has also been referred to as the update to data
-ratio (UTD) \citep{chenRandomizedEnsembledDouble2021}.
-\citet{nikishinDeepReinforcementLearning2023} propose an alternative approach to overcome primacy bias which
-freezes the current NN and creates a new one which learns the changes in predictions.
-
-When leveraging latent representations for RL (\eg SLAC \citep{leeStochasticLatentActorCritic2020},
-SAC-AE \citep{yaratsImprovingSampleEfficiency2021}) this issue is exacerbated.
-Not only do the agents networks (\eg actor/critic) overfit on early data, but as the encoder is updated during training, the
-inputs (latent states) that the networks are trained on change.
-As such, the networks are overfit on latent states which no longer exist in the ``latent'' data set.
-
-\textbf{Misc}
-\citet{obando-ceronSmallBatchDeep2023a} show that reducing the batch size not only improves
-performance and wall-time gains, but also appears to have knock-on effects on exploration and asymptotic behaviour.
-
-\citet{lyuOffPolicyRLAlgorithms2023} show that performing multiple parameter updates for a given batch.
-Their method, name sample multiple reuse (SMR), improves sample efficiency
-They say not to use if for image-based environments like Atari.
-
-Simplical embeddings
-\cite{lavoieSimplicialEmbeddingsSelfSupervised2022}
 

### 2024-02-01 paper/main.tex
added algorithm

@@
 % if you use cleveref..
-\usepackage[capitalize,noabbrev,nameinlink]{cleveref}
+% \usepackage[capitalize,noabbrev,nameinlink]{cleveref}
+\usepackage[capitalize,nameinlink]{cleveref}
 
@@
 % \newcommand{\our}{\textsc{vq-td3}\xspace}
-\newcommand{\our}{\textsc{iFSQ-RL}\xspace}
+% \newcommand{\our}{\textsc{iFSQ-RL}\xspace}
+\newcommand{\our}{\textsc{iQRL}\xspace}
 
@@
 
-In this section, we detail our method, titled Vector Quantized TD3 (\our).
-\our is conceptually simple, it first learns a representation of the observation space and then,
-given this representation, it performs model-free RL (e.g. TD3) on this representation.
+In this section, we detail our method, titled \textit{implicitly Quantized Reinforcement Learning} (\our).
+\our is conceptually simple, it (1) learns a representation of the observation space and then,
+(2) performs model-free RL (e.g. TD3) on this representation.
 % It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
@@
 % It first learns a representation of the observation space and then uses
-See \cref{alg:main_alg}.
+See \cref{fig:diagram-self-predictive} and \cref{alg:main_alg}.
 
-We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma,\mathcal{T})$, where an agent receives
-observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.
+We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
+where an agent receives an observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.
 The agent then obtains a reward $r_{t} = \mathcal{R} (o_{t}, a_{t})$ and obtains the next observation
 $o_{t+1} = P(\cdot \mid o_{t}, a_{t})$.
-discount factor $\gamma \in [0, 1]$
+The discount factor is denoted $\gamma \in [0, 1]$.
 
-\our has four parameterized components.
 \begin{algorithm}[tb]
@@
    \label{alg:main_alg}
+   \renewcommand{\algorithmiccomment}[1]{\hfill\textcolor{gray}{\(\triangleright\) #1}}
 \begin{algorithmic}
-   \STATE {\bfseries Input:} Encoder $e_{\theta}$, transition $d_{\phi}$, critics $Q_{\psi}$, policy $\pi_{\eta}$
+   \STATE {\bfseries Input:} Encoder $e_{\theta}$, transition $d_{\phi}$, critics $\{Q_{\psi_{1}}, Q_{\psi_{2}} \}$, policy $\pi_{\eta}$, learning rate $\alpha$, target network update rate $\tau$
    \FOR{$i$ {\bfseries to} $N_{\text{episodes}}$}
-    \STATE $\mathcal{D} \leftarrow \mathcal{D} \cup \{o_{t}, a_{t}, o_{t+1}, r_{t+1}\}^{T}_{t=0}$
-    \FOR{$i=1$ {\bfseries to} $T \times UTD$}
-        \STATE Representation update
-        \STATE TD3 update
+    \STATE Collect data in environment
+    \STATE $\qquad \mathcal{D} \leftarrow \mathcal{D} \cup \{o_{t}, a_{t}, o_{t+1}, r_{t+1}\}^{T}_{t=0}$
+    \FOR{$i=1$ {\bfseries to} $T$}
+        \STATE Update representation
+        \STATE $\qquad [\theta, \phi] \leftarrow [\theta, \phi] + \alpha \nabla \left( \mathcal{L}_{\text{rep}}(\theta, \phi; \mathcal{D}) \right)$  \COMMENT{\cref{eq:rep-loss}}
+        \STATE Update critic
+        \STATE \quad \quad $\psi \leftarrow \psi + \alpha \nabla \left( \mathcal{L}_{Q}(\psi; \mathcal{D}) \right)$ \COMMENT{\cref{eq:value-loss}}
+        \IF{i \% 2 == 0}
+          \STATE Update actor less frequently than critic
+          \STATE $\quad \quad \eta \leftarrow \psi + \alpha \nabla \left( \mathcal{L}_{\pi}(\psi; \mathcal{D}) \right)$  \COMMENT{\cref{eq:policy-loss}}
+          % \STATE $\bar{\eta} \leftarrow \tau \bar{\eta} + (1-\tau) \eta$
+        \ENDIF
+        \STATE Update target networks
+        \STATE $\quad \quad [\bar{\theta}, \bar{\phi}, \bar{\psi}] \leftarrow \tau [\bar{\theta}, \bar{\phi}, \bar{\psi}] + (1-\tau) [{\theta}, {\phi}, {\psi}]$
     \ENDFOR
@@
 
+\our has four parameterized components.
 \begin{align}
@@
 Our representation learning adopts a self-supervised loss.
-
+%
 \begin{align} \label{eq:rep-loss}
-  \mathcal{L}_{\text{rep}}(\theta, \phi; \mathcal{D})
+  \mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
  % &= \E_{s_{t} \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| \hat{z}_{t+1} - z_{t+1} \|_{2}^{2} \right] \\
 %&= \E_{(s_{t}, a_{t}, s_{t+1}) \sim \mathcal{D}}
-&= \E_{\tau \sim \mathcal{D}}
-\left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
+% &= \E_{\tau \sim \mathcal{D}}
+% \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
+&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2}
 \end{align}
-
+%
 \textbf{Finite Scalar Quantization}

### 2024-02-01 paper/figs/plotting/reward-ablation/fetch_data.py
Reward-ablation and task-agnostic ablation updated

@@ -5,163 +5,289 @@ ddpg_data = {
     "data": {
         "quadruped-run": [
             #iFSQ-RL
-            "9mbrf8o3",
-            "lffrf7xy",
-            "zcloyygj",
-            "4lmtpd04",
-            "bud72vyt",
+#            "9mbrf8o3",
+#            "lffrf7xy",
+#            "zcloyygj",
+#            "4lmtpd04",
+#            "bud72vyt",
+            "225j3iqa",
+            "gsh5g5ie",
+            "wkwqz2hp",
+            "1fqkia25",
+            "bamy18xg",
             # iFSQL-RL+rew
-            "6jwu9xzm",
-            "3thseoe1",
-            "02tn6r73",
-            "gonirtj6",
-            "e3ak3zco",
+#            "6jwu9xzm",
+#            "3thseoe1",
+#            "02tn6r73",
+#            "gonirtj6",
+#            "e3ak3zco",
+            "5fd3wano",
+            "zhw57oda",
+            "27qykyze",
+            "ci0ik0ub",
+            "0hlxppg4",
             # iFSQ-RL+cos+rew
-            "sovpuj12",
-            "epkk4mw8",
-            "5mpggr4p",
-            "ams8eapm",
-            "6bg43249",
+#            "sovpuj12",
+#            "epkk4mw8",
+#            "5mpggr4p",
+#            "ams8eapm",
+#            "6bg43249",
+            "l5mr1ykt",
+            "3o91tqjt",
+            "tcecjfc5",
+            "7giv4t1b",
+            "chdxnkyl",
             # TCRL-ours
-            "ijm4oaud",
-            "3bm9cclp",
-            "aa3eaznd",
-            "px4fskkk",
-            "m62ducyj",
+#            "ijm4oaud",
+#            "3bm9cclp",
+#            "aa3eaznd",
+#            "px4fskkk",
+#            "m62ducyj",
+            "8nnny8uj",
+            "7j4amx4h",
+            "ngnrnyes",
+            "9ddmzk2n",
+            "ekef5rwc",
             # TCRL-ours-small
-            "73kkvwmy",
-            "12krzm4k",
-            "mp0d0djz",
-            "bhixjgby",
-            "6x1suc5x",
+#            "73kkvwmy",
+#            "12krzm4k",
+#            "mp0d0djz",
+#            "bhixjgby",
+#            "6x1suc5x",
         ],
         "cheetah-run": [
             #iFSQ-RL
[…]

### 2024-02-01 paper/figs/plotting/task-agnostic-ablation/fetch_data.py

@@ -45,11 +45,16 @@ ddpg_data = {
             "my89a2ui",
             "0bdbllyg",
             # Task-specific pre-training
-            "uykvengt",
-            "e6u4u0et",
-            "phlr610t",
-            "32c5rcc6",
-            "2hyd7evc",
+#            "uykvengt",
+#            "e6u4u0et",
+#            "phlr610t",
+#            "32c5rcc6",
+#            "2hyd7evc",
+            "lc8kap78",
+            "fs0fug53",
+            "rypu71yu",
+            "7lerwfz0",
+            "jcvs7slh",
         ],
         "walker-run": [
             # Pre-training"
@@ -65,11 +70,16 @@ ddpg_data = {
             "cbp72vz9",
             "oi0gnnd1",
             # Task-specific pre-training
-            "oxisl0xv",
-            "wv0a1i6l",
-            "2mstpcgd",
-            "olnugqts",
-            "1iegxig2",
+#            "oxisl0xv",
+#            "wv0a1i6l",
+#            "2mstpcgd",
+#            "olnugqts",
+#            "1iegxig2",
+            "g029b1dc",
+            "2dy32443",
+            "1ogr1lye",
+            "8rtrpoht",
+            "glmvfvho",
         ],
     },
 }

### 2024-02-01 paper/main.tex

@@
 &\text{Value: } & q_{t} &= Q_{\psi} (z_{t}, a_{t}) \label{eq:value} \\
-&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) , \label{eq:policy}
+&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) . \label{eq:policy}
 \end{align}
-Note that unlike TCRL \citep{zhaoSimplifiedTemporalConsistency2023}, our transition model does
-not predict the reward.
+Central to our method is the latent state $z_{t}$
+
+\textbf{Bounding the latent space}
+Typically the latent state $z_{t} \in \R^{d}$
+However, in our work we bound the domain of the latent space, \ie $z_{t} \in [-L, L]^{d}$.
 
@@
 \end{align}
+
+Whilst our method resembles TCRL \citep{zhaoSimplifiedTemporalConsistency2023} it is important to note that
+our transition model does not predict the reward.
+As such, our method learns a task-agnostic representation.
 

### 2024-02-01 paper/figs/plotting/reward-ablation/fetch_data.py
Updated reward and task-agnostic ablations

@@ -348,12 +348,12 @@ def fetch_results(run_path, run_name_list, keys, agent_name=None):
         history["env"] = env_name
         history["seed"] = wandb_run.config["seed"]
         if not wandb_run.config["agent"]["reward_loss"]:
-            history["name"] = "iFSQ-RL"
+            history["name"] = "iQRL"
         elif not wandb_run.config["agent"]["use_cosine_similarity_dynamics"]:
-            history["name"] = "iFSQ-RL-rew"
+            history["name"] = "iQRL+rew"
         elif wandb_run.config["agent"]["use_cosine_similarity_dynamics"] and \
                 wandb_run.config["agent"]["use_fsq"]:
-            history["name"] = "iFSQ-RL-cos-rew"
+            history["name"] = "iQRL+cos+rew"
         elif (not wandb_run.config["agent"]["use_fsq"]) and \
                 wandb_run.config["agent"]["latent_dim"] > 50:
             history["name"] = "TCRL-ours"

### 2024-02-01 paper/figs/plotting/task-agnostic-ablation/fetch_data.py

@@ -139,15 +139,23 @@ def fetch_results(run_path, run_name_list, keys):
         history["time"] = 0
         history["env"] = env_name
         history["seed"] = wandb_run.config["seed"]
-        history = history[history["env_step"] <= 2e6]
+
+        if env_name == "hopper-hop":
+            pass
+        elif env_name == "humanoid-run":
+            history = history[history["env_step"] <= 2e6]
+        elif env_name in ["walker-run", "quadruped-run"]:
+            history = history[history["env_step"] <= 500000]
+        else:
+            raise NotImplementedError
 
         # history["agent"] = agent_name
         if not wandb_run.config["load_pretrained_agent"]:
-            history["name"] = "iFSQ-RL"
+            history["name"] = "iQRL"
         elif wandb_run.config["agent"]["reward_loss"]:
-            history["name"] = "iFSQ-RL-spec-trf"
+            history["name"] = "iQRL-pretrained"
         else:
-            history["name"] = "iFSQ-RL-agn-trf"
+            history["name"] = "iQRL+rew-pretrained"
             # history["name"] = f"no-norm $d={wandb_run.config['agent']['latent_dim']}$"
         history["utd_ratio"] = wandb_run.config["agent"]["utd_ratio"]
 

### 2024-02-01 paper/figs/plotting/task-agnostic-ablation/make_plot.py

@@ -14,15 +14,17 @@ plt.rcParams["figure.dpi"] = 400
 plt.rcParams["font.size"] = 13
 plt.rcParams["legend.fontsize"] = 12
 plt.rcParams["legend.loc"] = "lower right"
+#plt.rcParams["text.usetex"] = True
+
 COLORS = {
     # "TCRL": "#e41a1c",
     # "SAC": "#377eb8",
     # "REDQ": "#984ea3",
     # "TD-MPC": "#ff7f00",
     # "VQ-TD3": "magenta",
-    "iFSQ-RL": "#e41a1c",
-    "iFSQ-RL-spec-trf": "#377eb8",
-    "iFSQ-RL-agn-trf": "#ff7f00",
+    "iQRL": "#e41a1c",
+    "iQRL-pretrained": "#377eb8",
+    "iQRL+rew-pretrained": "#ff7f00",
 }
 # %%
 main_envs = [
@@ -35,7 +37,7 @@ main_envs = [
     "quadruped-run",
     # "humanoid-walk",
     "humanoid-run",
-    "hopper-hop",
+    # "hopper-hop",
     # "dog-walk",
     # "dog-run",
 ]
@@ -60,7 +62,7 @@ rename = {a: b for a, b in zip(keys, titles)}
 def plot(df, key="episode_reward"):
     # breakpoint()
     envs = np.sort(df.env.unique())
-    ncol = 4
+    ncol = 3
     # assert envs.shape[0] % ncol == 0
     # nrow = len(main_envs) // ncol
     nrow = 1

### 2024-02-01 paper/figs/plotting/norm-ablation/fetch_data.py
updated plot

@@ -69,6 +69,14 @@ ddpg_data = {
             "ucvgujrp",
             "zqv19l3p",
         ],
+        "fish-swim-no-norm": [
+            "a2r96nmy",
+            "1cezxqig",
+            "15r32rsc",
+            "rxyyhje0",
+            "8pho3asn",
+        ],
+        "fish-swim": ["3duujm1p", "kl6xd92o", "s4t5viko", "flcmnkt7", "1zjl534e"],
         "humanoid-run": [
             # no-norm d=1024 project=False
             "uuyjutba",
@@ -312,9 +320,9 @@ def fetch_results(run_path, run_name_list, keys):
         history["seed"] = wandb_run.config["seed"]
         # history["agent"] = agent_name
         if wandb_run.config["agent"]["use_fsq"]:
-            history["name"] = "iFSQ-RL"
+            history["name"] = "iQRL"
         else:
-            history["name"] = f"No iFSQ"
+            history["name"] = "iQRL-no-normalization"
             # history["name"] = f"no-norm $d={wandb_run.config['agent']['latent_dim']}$"
         history["utd_ratio"] = wandb_run.config["agent"]["utd_ratio"]
 

### 2024-02-01 paper/figs/plotting/norm-ablation/make_plot.py

@@ -11,17 +11,18 @@ plt.style.use("seaborn-v0_8-whitegrid")
 import seaborn as sns
 
 plt.rcParams["figure.dpi"] = 400
-plt.rcParams["font.size"] = 13
-plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["font.size"] = 15
+plt.rcParams["legend.fontsize"] = 14
 plt.rcParams["legend.loc"] = "lower right"
+plt.rcParams["text.usetex"] = True
 COLORS = {
     # "TCRL": "#e41a1c",
     # "SAC": "#377eb8",
     # "REDQ": "#984ea3",
     # "TD-MPC": "#ff7f00",
     # "VQ-TD3": "magenta",
-    "iFSQ-RL": "#e41a1c",
-    "No iFSQ": "#377eb8",
+    "iQRL": "#e41a1c",
+    "iQRL-no-normalization": "#377eb8",
     "no-norm $d=50$": "#377eb8",
     "no-norm $d=512$": "#ff7f00",
     "no-norm $d=1024$": "#984ea3",
@@ -33,12 +34,12 @@ main_envs = [
     # "walker-walk",
     "walker-run",
     "hopper-stand",
-    # "fish-swim",
+    "fish-swim",
     "quadruped-run",
     "humanoid-walk",
     "humanoid-run",
     # "dog-walk",
-    "dog-run",
+    # "dog-run",
 ]
 
 # pet-pytorch

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/fetch_data.py

@@ -180,6 +180,8 @@ ddpg_data = {
         #     # "lh1kdlht",  # seed=4
         #     # "8gymqblt",  # seed=5
         # ],
+        "fish-swim-rec": ["6vss1lh5", "ugzcaey2", "b8q0nc3f", "geecyirj", "79waw193"],
+        "fish-swim": ["3duujm1p", "kl6xd92o", "s4t5viko", "flcmnkt7", "1zjl534e"],
         "acrobot-swingup": [
             # reconstruction
             "6sfcsm49",
@@ -209,10 +211,10 @@ ddpg_data = {
             "wrggk81m",  # seed=5
             # no reconstruction
             # project=False
-            "stawiad8",
+            # "stawiad8",
             "iv8p110s",
-            "k454j543",
-            "nihkvll9",
+            # "k454j543",
+            # "nihkvll9",
             "jpkbw5h7",
             # project=True
             # "9k5cqnuq",
@@ -287,9 +289,9 @@ def fetch_results(run_path, run_name_list, keys):
         history["seed"] = wandb_run.config["seed"]
         # history["agent"] = agent_name
         if wandb_run.config["agent"]["reconstruction_loss"]:
-            history["name"] = "iFSQ-RL+rec"
+            history["name"] = "iQRL+rec"
         else:
-            history["name"] = "iFSQ-RL"
+            history["name"] = "iQRL"
 
         data.append(history)
         # breakpoint()

### 2024-02-01 paper/figs/plotting/reconstruction-ablation/make_plot.py

@@ -11,17 +11,18 @@ plt.style.use("seaborn-v0_8-whitegrid")
 import seaborn as sns
 
 plt.rcParams["figure.dpi"] = 400
-plt.rcParams["font.size"] = 13
-plt.rcParams["legend.fontsize"] = 12
+plt.rcParams["font.size"] = 15
+plt.rcParams["legend.fontsize"] = 14
 plt.rcParams["legend.loc"] = "lower right"
+plt.rcParams["text.usetex"] = True
 COLORS = {
     # "TCRL": "#e41a1c",
     # "SAC": ,
     # "REDQ": "#984ea3",
     # "TD-MPC": "#ff7f00",
     # "VQ-TD3": "magenta",
-    "iFSQ-RL+rec": "#377eb8",
-    "iFSQ-RL": "#e41a1c",
+    "iQRL+rec": "#377eb8",
+    "iQRL": "#e41a1c",
 }
 # %%
 main_envs = [

### 2024-02-01 paper/figs/plotting/reward-ablation/make_plot.py
Updated task agnostic and reward

@@ -14,7 +14,7 @@ plt.rcParams["figure.dpi"] = 400
 plt.rcParams["font.size"] = 13
 plt.rcParams["legend.fontsize"] = 12
 plt.rcParams["legend.loc"] = "lower right"
-#plt.rcParams["text.usetex"] = True
+plt.rcParams["text.usetex"] = True
 COLORS = {
     # "TCRL": "#e41a1c",
     # "SAC": "#377eb8",
@@ -29,8 +29,8 @@ COLORS = {
 }
 # %%
 main_envs = [
-    "acrobot-swingup",
     "cheetah-run",
+    "acrobot-swingup",
     # "walker-walk",
     "walker-run",
     "hopper-stand",
@@ -63,10 +63,10 @@ rename = {a: b for a, b in zip(keys, titles)}
 def plot(df, key="episode_reward"):
     # breakpoint()
     envs = np.sort(df.env.unique())
-    ncol = 6
+    ncol = 3
     # assert envs.shape[0] % ncol == 0
     # nrow = len(main_envs) // ncol
-    nrow = 1
+    nrow = 2
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
@@ -78,7 +78,7 @@ def plot(df, key="episode_reward"):
         row = idx // ncol
         col = idx % ncol
         # ax = axs[row, col]
-        ax = axs[col]
+        ax = axs[row, col]
         # hue_order = data.agent.unique()
         hue_order = data.name.unique()
         # hue_order = data.utd_ratio.unique()
@@ -86,34 +86,67 @@ def plot(df, key="episode_reward"):
 
         if idx == 0:
             sns.lineplot(
+                # x=int("env_step" / 1000),
                 x="env_step",
+                # x="env_step",
                 # x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
                 hue="name",
-                # style="utd_ratio",
-                hue_order=hue_order,
+                # hue_order=hue_order,
                 palette=COLORS,
                 legend="auto",
                 ax=ax,
             )
             ax.legend().set_title(None)
         else:
-            # breakpoint()
             sns.lineplot(
-                x="env_step",
                 # x="episode",
+                x="env_step",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
                 hue="name",
-                # style="utd_ratio",
-                hue_order=hue_order,
[…]

### 2024-02-01 paper/main.tex
wrote abstract

@@
 \begin{abstract}
-This document provides a basic paper template and submission guidelines.
-Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-Gross violations will trigger corrections at the camera-ready phase.
+  % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
+  % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
+  % is limited by its sample-efficiency.
+  % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
+  % handling complex/high-dimensional observation spaces.
+  Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
+  Not only can it handle complex observation spaces but it can also improve sample efficiency.
+  % Recent works have shown that learning representations can improve sample efficiency in RL.
+  A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
+  It requires an encoder which maps observations to latent states and a transition model which predicts
+  next latent states.
+  It then minimizes the distance between the next state predicted by the transition model and
+  the next state predicted by the encoder.
+  Typically it is combined with other loss terms such as reward or value prediction.
+  This prevents representation collapse at the cost of learning a task-specific representation.
+  In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
+  a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
+  consistency loss.
+  % the leverage the latent-state
+  % Typically these approaches have unbounded latent spaces and are combined with auxillary loss terms such as minimizing the
+  % error of reward or value predictions in the latent space.
+  % of an encoder which maps observations to latent states and a transition model which predicts
+  % next latent states given a latent state and an action.
+  % However, most of th
+  We empirically show that our method prevents representation collapse as it maintains matrix rank.
+  Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
+  and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
+  % Moreover, it kk
+% This document provides a basic paper template and submission guidelines.
+% Abstracts must be a single paragraph, ideally between 4--6 sentences long.
+% Gross violations will trigger corrections at the camera-ready phase.
 \end{abstract}

### 2024-02-01 paper/main.tex
udpates

@@
   Not only can it handle complex observation spaces but it can also improve sample efficiency.
-  % Recent works have shown that learning representations can improve sample efficiency in RL.
   A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
@@
   consistency loss.
-  % the leverage the latent-state
-  % Typically these approaches have unbounded latent spaces and are combined with auxillary loss terms such as minimizing the
-  % error of reward or value predictions in the latent space.
-  % of an encoder which maps observations to latent states and a transition model which predicts
-  % next latent states given a latent state and an action.
-  % However, most of th
   We empirically show that our method prevents representation collapse as it maintains matrix rank.
@@
 \end{abstract}
-
-\section{Introduction}
-\label{intro}
-
-Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
-However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
-deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
-
-% Given such a latent space,
-% Whilst these latent states are often lower dimensional this is not always
-Many approaches have been proposed to improve sample efficiency in deep RL.
-As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
-\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-In particular, methods for compressing observations into latent states.%, aka representation learning.
-
-Learning representations for RL has been investigated for decades
-\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
-However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
-More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
-attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
@@
 \end{tikzpicture}
-\caption{\textbf{\our's model architecture} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
+\caption{\textbf{Overview of \our's representation learning} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
 We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
-To this end, our normalization results in an implicit codebook whose size we can control.
+Our normalization results in an implicit codebook whose size we can control.
+  % Importantly, this prevents representation collapse without needing loss terms to minimize errors in reward and/or value predictions.
+  % As such, our representations are task-agnostic.
   We denote our normalization scheme by \cube.}
@@
 
-To the best of our knowledge, all prior works learn representations with objectives that depend on
+\section{Introduction}
+\label{intro}
+
+Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+However, applying RL in real-world environments is challenging. The problem is two-fold,
+{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+
+% Given such a latent space,
+% Whilst these latent states are often lower dimensional this is not always
+Many approaches have been proposed to improve sample efficiency in deep RL.
+As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
+In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+In particular, we focus on representation learning methods which compress observations into latent states.
+Nevertheless, our representation learning borrows ideas from world models as it uses multi-step predictions in a
+latent-space dynamics model to learn temporally consist representations \citep{zhaoSimplifiedTemporalConsistency2023}.
+% In particular, methods for compressing observations into latent states.%, aka representation learning.
+
+Learning representations for RL has been investigated for decades
+\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
+However, these approaches are usually limited to simple environments.
+Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
+attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
+Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
+
+To the best of our knowledge, all prior works that learn representations with objectives that depend on
 predicting the reward and/or value in the latent space
 \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+This is usually motivated by preventing representation collapse.
 % use a latent-state consistency loss based upon the Cosine
@@
 % The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
-As such, these methods learn task-specific representations.
+However, learning representations using reward and/or value loss terms leads to the representations being task specific.
 
@@
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
-It is based soley on a self-predictive loss.
+It is based soley on the latent-state consistency loss, \ie a commonly use self-supervised loss for RL.
 In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
 \citep{mentzerFiniteScalarQuantization2023}.
-As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} is simple to implement,
-and {\em (iii)} is task-agnostic.
+As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
+% As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
+Importantly, our method {\em (i)} prevents representation collapse (as it learns a high-rank representation),
+{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
+{\em (iii)} is simple to implement, and {\em (iii)} learns a task-agnostic
+representation which could be helpful in downstream tasks.
 

### 2024-02-01 paper/main.tex
Abstract

--- 2024-02-01 paper/main.tex
@@
 \begin{abstract}
+  % Arno's proposal
+  In reinforcement learning (RL) for continuous control, we propose  an efficient representation learning method using a latent-state consistency loss.
+This approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
+By minimizing discrepancies between these states, we achieve task-agnostic representations and prevent representation collapse using only the consistency loss.
+Our method is straightforward, compatible with any model-free RL algorithm, and demonstrates state-of-the-art performance in continuous control benchmarks from DeepMind Control Suite.\looseness-1
+
   % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.

### 2024-02-01 paper/main.tex
updated intro

--- 2024-02-01 paper/main.tex
@@
 We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
-Our normalization results in an implicit codebook whose size we can control.
+Our normalization results in an implicit codebook (whose size we can control) and helps prevent representation collapse.
   % Importantly, this prevents representation collapse without needing loss terms to minimize errors in reward and/or value predictions.
@@
 attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
+% Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
-To the best of our knowledge, all prior works that learn representations with objectives that depend on
-predicting the reward and/or value in the latent space
+Recently, TD-MPC \citep{hansenTemporalDifferenceLearning2022} and TCRL \citep{zhaoSimplifiedTemporalConsistency2023}
+have obtained state-of-the-art performance on continuous control benchmarks by learning
+representations with self-supervised losses.
+In particular, they leverage the latent-state consistency loss.
+This requires an encoder $e_{\theta}$ which maps observations $o_{t}$ to latent states $z_{t}$ and a transition model $d_{\phi}$ which predicts
+next latent states $z_{t+1}$ given an a latent state $z_{t}$ and an action $a_{t}$.
+It then minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
+and the next state predicted by the target encoder $e_{\bar{\phi}}(z_{t+1})$.
+Self-supervised losses are susceptible to a problem known as representation collapse, where the encoder learns a constant
+representation \cite{jingUnderstandingDimensionalCollapse2021}.
+As such, it is common to combine the latent-state consistency loss with other loss terms, such as
+minimizing the reward prediction error in the latent space
 \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
-This is usually motivated by preventing representation collapse.
+This helps to prevent representation collapse at the cost of learning a task-specific representation.
+% Typically it is combined with other loss terms such as reward or value prediction.
+
+% To the best of our knowledge, all prior works that learn representations with objectives that depend on
+% predicting the reward and/or value in the latent space
+% \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+% This is usually motivated by preventing representation collapse.
+% To the best of our knowledge, all prior works that learn representations with objectives that depend on
+% predicting the reward and/or value in the latent space
+% \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+% This is usually motivated by preventing representation collapse.
 % use a latent-state consistency loss based upon the Cosine
@@
 % The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
-However, learning representations using reward and/or value loss terms leads to the representations being task specific.
+% However, learning representations using reward and/or value loss terms leads to the representations being task specific.
 

### 2024-02-01 paper/main.tex
Updated abstract

--- 2024-02-01 paper/main.tex
@@
 \begin{abstract}
-  % Arno's proposal
-  In reinforcement learning (RL) for continuous control, we propose  an efficient representation learning method using a latent-state consistency loss.
-This approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
-By minimizing discrepancies between these states, we achieve task-agnostic representations and prevent representation collapse using only the consistency loss.
-Our method is straightforward, compatible with any model-free RL algorithm, and demonstrates state-of-the-art performance in continuous control benchmarks from DeepMind Control Suite.\looseness-1
+  In reinforcement learning (RL) for continuous control, we propose an
+  efficient representation learning method using only a latent-state
+  consistency loss. Our approach employs an encoder and a transition
+  model to map observations to latent states and predict future
+  states, respectively. We achieve high performance and prevent
+  representation collapse by normalizing, quantizing, and bounding the
+  latent representation. Our method is straightforward, compatible
+  with any model-free RL algorithm, and demonstrates state-of-the-art
+  performance in continuous control benchmarks from DeepMind Control
+  Suite.
+\end{abstract}
 
-  % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
-  % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
-  % is limited by its sample-efficiency.
-  % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
-  % handling complex/high-dimensional observation spaces.
-  Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
-  Not only can it handle complex observation spaces but it can also improve sample efficiency.
-  A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
-  It requires an encoder which maps observations to latent states and a transition model which predicts
-  next latent states.
-  It then minimizes the distance between the next state predicted by the transition model and
-  the next state predicted by the encoder.
-  Typically it is combined with other loss terms such as reward or value prediction.
-  This prevents representation collapse at the cost of learning a task-specific representation.
-  In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
-  a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
-  consistency loss.
-  We empirically show that our method prevents representation collapse as it maintains matrix rank.
-  Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
-  and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
-  % Moreover, it kk
-% This document provides a basic paper template and submission guidelines.
-% Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-% Gross violations will trigger corrections at the camera-ready phase.
-\end{abstract}
+%% \begin{abstract}
+%%   Learning representations for reinforcement learning (RL) has shown
+%%   much promise for continuous control. Representations not only can
+%%   handle complex observation spaces but also improve sample
+%%   efficiency. A particularly promising approach is to use a
+%%   self-supervised latent-state consistency loss with observation to
+%%   latent state mapping, together with next latent state, reward, and
+%%   value predictions. Reward and value predictions prevent
+%%   representation collapse at the cost of learning a task-specific
+%%   representation.  We show how to use only the latent-state
+%%   consistency loss to {\em (i)} learn a task-agnostic representation
+%%   and {\em (ii)} prevent representation collapse by normalizing and
+%%   bounding the latent representation. In experiments, our approach
+%%   which be combined with any model-free RL algorithm obtains
+%%   state-of-the-art results in continuous control benchmarks from
+%%   DeepMind Control Suite.
+%% \end{abstract}
+
+%% \begin{abstract}
+%%   % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
+%%   % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
+%%   % is limited by its sample-efficiency.
+%%   % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
+%%   % handling complex/high-dimensional observation spaces.
+%%   Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
+%%   Not only can it handle complex observation spaces but it can also improve sample efficiency.
+%%   % Recent works have shown that learning representations can improve sample efficiency in RL.
+%%   A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
+%%   It requires an encoder which maps observations to latent states and a transition model which predicts
+%%   next latent states.
+%%   It then minimizes the distance between the next state predicted by the transition model and
+%%   the next state predicted by the encoder.
+%%   Typically it is combined with other loss terms such as reward or value prediction.
+%%   This prevents representation collapse at the cost of learning a task-specific representation.
+%%   In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
+%%   a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
+%%   consistency loss.
+%%   % the leverage the latent-state
+%%   % Typically these approaches have unbounded latent spaces and are combined with auxillary loss terms such as minimizing the
+%%   % error of reward or value predictions in the latent space.
+%%   % of an encoder which maps observations to latent states and a transition model which predicts
+%%   % next latent states given a latent state and an action.
+%%   % However, most of th
+%%   We empirically show that our method prevents representation collapse as it maintains matrix rank.
+%%   Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
+%%   and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
+%%   % Moreover, it kk
+%% % This document provides a basic paper template and submission guidelines.
+%% % Abstracts must be a single paragraph, ideally between 4--6 sentences long.
+%% % Gross violations will trigger corrections at the camera-ready phase.
+%% \end{abstract}
+
+\section{Introduction}
+\label{intro}
+
+Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+However, applying RL in real-world environments is challenging. The problem is two-fold,
+{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+
+% Given such a latent space,
+% Whilst these latent states are often lower dimensional this is not always
+Many approaches have been proposed to improve sample efficiency in deep RL.
+As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
+In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+In particular, methods for compressing observations into latent states.%, aka representation learning.
+
+Learning representations for RL has been investigated for decades
+\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
+However, these approaches are usually limited to simple environments.
+Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
+attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
+Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
@@
 \end{tikzpicture}
-\caption{\textbf{Overview of \our's representation learning} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
+\caption{\textbf{\our's model architecture} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
 We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
-Our normalization results in an implicit codebook whose size we can control.
-  % Importantly, this prevents representation collapse without needing loss terms to minimize errors in reward and/or value predictions.
-  % As such, our representations are task-agnostic.
+To this end, our normalization results in an implicit codebook whose size we can control.
 We denote our normalization scheme by \cube.}
@@
 
-\section{Introduction}
-\label{intro}
-
-Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
-However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
-deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
-
-% Given such a latent space,
-% Whilst these latent states are often lower dimensional this is not always
-Many approaches have been proposed to improve sample efficiency in deep RL.
-As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
-\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-In particular, we focus on representation learning methods which compress observations into latent states.
-Nevertheless, our representation learning borrows ideas from world models as it uses multi-step predictions in a
-latent-space dynamics model to learn temporally consist representations \citep{zhaoSimplifiedTemporalConsistency2023}.
-% In particular, methods for compressing observations into latent states.%, aka representation learning.
-
-Learning representations for RL has been investigated for decades
-\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
-However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
-More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
-attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
-
-To the best of our knowledge, all prior works that learn representations with objectives that depend on
+To the best of our knowledge, all prior works learn representations with objectives that depend on
 predicting the reward and/or value in the latent space
 \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
-This is usually motivated by preventing representation collapse.
 % use a latent-state consistency loss based upon the Cosine
@@
 % The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
-However, learning representations using reward and/or value loss terms leads to the representations being task specific.
+As such, these methods learn task-specific representations.
 
@@
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
-It is based soley on the latent-state consistency loss, \ie a commonly use self-supervised loss for RL.
+It is based soley on a self-predictive loss.
 In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
 \citep{mentzerFiniteScalarQuantization2023}.
-As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
-% As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our method {\em (i)} prevents representation collapse (as it learns a high-rank representation),
-{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
-{\em (iii)} is simple to implement, and {\em (iii)} learns a task-agnostic
-representation which could be helpful in downstream tasks.
+As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
+Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} is simple to implement,
+and {\em (iii)} is task-agnostic.
 

### 2024-02-01 paper/main.tex
Updates

@@
 % \newcommand{\our}{\textsc{iFSQ-RL}\xspace}
+\newcommand{\ourrew}{\textsc{iQRL+rew}\xspace}
+\newcommand{\ourcosrew}{\textsc{iQRL+cos+rew}\xspace}
 \newcommand{\our}{\textsc{iQRL}\xspace}
@@
 \begin{abstract}
-  In reinforcement learning (RL) for continuous control, we propose an
-  efficient representation learning method using only a latent-state
-  consistency loss. Our approach employs an encoder and a transition
-  model to map observations to latent states and predict future
-  states, respectively. We achieve high performance and prevent
-  representation collapse by normalizing, quantizing, and bounding the
-  latent representation. Our method is straightforward, compatible
-  with any model-free RL algorithm, and demonstrates state-of-the-art
-  performance in continuous control benchmarks from DeepMind Control
-  Suite.
+This document provides a basic paper template and submission guidelines.
+Abstracts must be a single paragraph, ideally between 4--6 sentences long.
+Gross violations will trigger corrections at the camera-ready phase.
 \end{abstract}
-
-%% \begin{abstract}
-%%   Learning representations for reinforcement learning (RL) has shown
-%%   much promise for continuous control. Representations not only can
-%%   handle complex observation spaces but also improve sample
-%%   efficiency. A particularly promising approach is to use a
-%%   self-supervised latent-state consistency loss with observation to
-%%   latent state mapping, together with next latent state, reward, and
-%%   value predictions. Reward and value predictions prevent
-%%   representation collapse at the cost of learning a task-specific
-%%   representation.  We show how to use only the latent-state
-%%   consistency loss to {\em (i)} learn a task-agnostic representation
-%%   and {\em (ii)} prevent representation collapse by normalizing and
-%%   bounding the latent representation. In experiments, our approach
-%%   which be combined with any model-free RL algorithm obtains
-%%   state-of-the-art results in continuous control benchmarks from
-%%   DeepMind Control Suite.
-%% \end{abstract}
-
-%% \begin{abstract}
-%%   % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
-%%   % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
-%%   % is limited by its sample-efficiency.
-%%   % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
-%%   % handling complex/high-dimensional observation spaces.
-%%   Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
-%%   Not only can it handle complex observation spaces but it can also improve sample efficiency.
-%%   % Recent works have shown that learning representations can improve sample efficiency in RL.
-%%   A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
-%%   It requires an encoder which maps observations to latent states and a transition model which predicts
-%%   next latent states.
-%%   It then minimizes the distance between the next state predicted by the transition model and
-%%   the next state predicted by the encoder.
-%%   Typically it is combined with other loss terms such as reward or value prediction.
-%%   This prevents representation collapse at the cost of learning a task-specific representation.
-%%   In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
-%%   a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
-%%   consistency loss.
-%%   % the leverage the latent-state
-%%   % Typically these approaches have unbounded latent spaces and are combined with auxillary loss terms such as minimizing the
-%%   % error of reward or value predictions in the latent space.
-%%   % of an encoder which maps observations to latent states and a transition model which predicts
-%%   % next latent states given a latent state and an action.
-%%   % However, most of th
-%%   We empirically show that our method prevents representation collapse as it maintains matrix rank.
-%%   Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
-%%   and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
-%%   % Moreover, it kk
-%% % This document provides a basic paper template and submission guidelines.
-%% % Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-%% % Gross violations will trigger corrections at the camera-ready phase.
-%% \end{abstract}
 
@@
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
-It is based soley on a self-predictive loss.
+It is based solely on a self-predictive loss.
 In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
@@
 \textbf{Representation learning}
-Our representation learning is based soley on the temporal consistency loss,
+Our representation learning is based solely on the temporal consistency loss,
 However, unlike previous works, we normalize the latent vector
@@
 	\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/reward-ablation.pdf}}
-\caption{\textbf{Reward prediction is not necessary for representation learning}.
-\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:reward-ablation}
+		\centerline{\includegraphics[width=\textwidth]{./figs/task-agnostic-ablation.pdf}}
+		\caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
+We evaluate whether pre-training \our on the walk tasks in the DeepMind Control Suite help for learning the run task on the same environment.
+The pre-training proves to be beneficial on all evaluated environments.
+If the representation is task-specific (the reward prediction term is included), the pre-training can be clearly detrimental.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+		\label{fig:multi-task-pretraining}
 	\end{center}
@@
 \begin{center}
-\centerline{\includegraphics[width=\textwidth]{./figs/task-agnostic-ablation.pdf}}
-\caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
-\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:multi-task-pretraining}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/reward-ablation.pdf}}
+\caption{\textbf{Reward prediction is not necessary for representation learning}.
+We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
+\our performs on par with the best method on all environments except Fish Swim, where it is slightly behind our version of TCRL.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:reward-ablation}
 \end{center}
@@
 
-\textbf{\our's representation is helpful for multi-task RL}
-In contrast to TCRL and TD-MPC, \our learns a representation that is task-agnostic.
-As such, \our's task-agnostic representations can be used to speed-up learning in the incremental multi-task setting.
-That is, leverage information learned from solving one task to aid learning another task in the same environment.
-For example, pre-training our representation in walker-walk and then using this to speed up training in walker-walk.
-In \cref{fig:multi-task-pretraining} we show that \our benefits from pre-training the representation in walker-walk
-before training on walker-run.
-In contrast, TCRL does not see this improvement.
-This is because its representation loss contains a reward prediction term which makes its representation task-specific.
-These results demonstrate that confirms task-agnostic representation
+\textbf{Reward predict is not necessary for representation learning}
+Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (Equation~\ref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
+Instead, we rely solely on the temporal consistency loss.
+To analyze the impact of not including the reward prediction term, we perform an ablation study, where we compare our method to three variants of our method:
+\begin{itemize}
+	\item \ourrew, where we add a reward prediction head into our model architecture by re-defining the dynamics as $\hat{z}_{t+1}, \hat{r}_t = d_{\phi} (z_{t}, a_{t})$ (see Equation~\ref{eq:transition}), and include a reward prediction term (discounted MSE Loss) in the representation loss:
+	\begin{align*}
+		\mathcal{L}_\text{rew} = \sum_{h=0}^{H-1} \gamma^h \| \hat{r}_{t+h} - r_{t+h}\|_2^2,
+	\end{align*}
+	where $\tilde{r}_{t+h}$ is the ground-truth H-step reward and $\hat{r}_{t+h}$ the predicted H-step rewards.
+	\item \ourcosrew, where in addition to the reward prediction head in \ourrew, we additionally convert the self-supervised MSE loss (see Equation~\ref{eq:rep-loss}) for the temporal consistency into a cosine similarity, similarly to TCRL:
+	\begin{align*} \label{eq:rep-loss}
+		\hat{z}_{t+h} &= d_{\phi}(e_{\theta}(s_{t}), a_{t}) \\
+		\tilde{z}_{t+h} &= e_{\bar{\theta}}(s_{t+1}) \\
+		\mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
+		&= - \sum_{h=0}^{H-1} \gamma^{h} \left ( \frac{\hat{z}_{t+h}}{\| \hat{z}_{t+h} \|_2} \right )^T \left ( \frac{\tilde{z}_{t+h}}{\| \tilde{z}_{t+h} \|_2} \right )
+	\end{align*}
+	\item TCRL-ours, where in addition to \ourcosrew, we remove the FSQ normalization scheme to make the method equal to TCRL with some minor improvements to the method, such as using nstep-TD3 \cite{fujimotoAddressingFunctionApproximation2018} as the base algorithm instead of DDPG with Double Q and nstep returns, using the neural network architecture from \citet{hansen2023td}, and performing policy smoothing \cite{fujimotoAddressingFunctionApproximation2018}.
+\end{itemize}
+The results for this ablation study are shown in Figure~\ref{fig:reward-ablation}. 
+The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods with the exception of Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
+Even then, \our performs equally to the normalized variants of our method with the reward prediction.
+Our results imply that learning to predict the reward is not necessary for learning a suitable latent representation.
+The upside of not including the reward prediction head is that it makes learning the representation learning task-agnostic, which we believe to be   important for downstream applications such as learning in an incremental multi-task setting.
+
+\textbf{\our's task-agnostic representation is helpful for multi-task RL}
+In contrast to TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (see Equation~\ref{eq:rep-loss}).
+The potential upside of \our's task-agnostic representations is that they can be used to speed-up learning in the incremental multi-task setting.
+That is, we can leverage the representation learned from solving one task to aid learning another tasks in the same domain.
+For instance, we can pre-train a representation in Walker Walk, and use it to speed up the training process in Walker Run.
+The only requirement is that the action and observation spaces must match.
+In addition, we assume that the transition dynamics are unchanged, but the reward may change.
+
+We evaluate whether pre-training representations is beneficial on three environments.
+We start by pre-training \our on Walker Walk, Quadruped Walk and Humanoid Walk.
+In addition, we replicate the pre-training on Walker Walk and Quadruped Walk by adding a reward prediction term to the representation loss.
+Then, we keep the encoder and the dynamics function from the pre-training, and perform training on the target tasks, which are Walker Run, Quadruped Run, and Humanoid-Run.
+We compare the pre-trained agents to training \our on these tasks from scratch.
+The results for this experiment can be found in Figure~\ref{fig:multi-task-pretraining}.
+
+The plots in \cref{fig:multi-task-pretraining} show that when the encoder and the dynamics function have been pre-trained, the agent learns quicker than when it is traind from scratch.
+This effect is particularly visible in Quadruped Run and Humanoid Run.
+In contrast, when the reward term has been included in the pre-training objective, making the representation task-specific, the pre-training proves to be harmful on Quadruped Run.
+This results strongly indicate that the task-agnostic representations are highly beneficial in the continual and lifelong learning settings.
+
+% we show that \our can benefit from pre-training the representation in a different environment with the same dynamics. In contrast, if the representation learning objective used
+
+%pre-training objective contains
+
+%That is, leveraging information learned from solving one task to aid learning another task in the same environment.
+%For example, pre-training our representation in walker-walk and then using this to speed up training in walker-walk.
+%In \cref{fig:multi-task-pretraining} we show that \our benefits from pre-training the representation in walker-walk
+%before training on walker-run.
+%In contrast, TCRL does not see this improvement.
+%This is because its representation loss contains a reward prediction term which makes its representation task-specific.
+%These results demonstrate that confirms task-agnostic representation
 
@@
 
-\subsection{Ablations}
-
-\textbf{Reward predict is not necessary for representation learning}
-Next, we test if using reward prediction aids representation learning.
-\cref{fig:reward-ablation} shows that
-
-\textbf{Reconstruction loss has detrimental impact}
+\textbf{Reconstruction loss has a detrimental impact}
 Recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}

### 2024-02-01 paper/figs/plotting/task-agnostic-ablation/fetch_data.py

@@ -153,9 +153,9 @@ def fetch_results(run_path, run_name_list, keys):
         if not wandb_run.config["load_pretrained_agent"]:
             history["name"] = "iQRL"
         elif wandb_run.config["agent"]["reward_loss"]:
-            history["name"] = "iQRL-pretrained"
-        else:
             history["name"] = "iQRL+rew-pretrained"
+        else:
+            history["name"] = "iQRL-pretrained"
             # history["name"] = f"no-norm $d={wandb_run.config['agent']['latent_dim']}$"
         history["utd_ratio"] = wandb_run.config["agent"]["utd_ratio"]
 

### 2024-02-01 paper/main.tex

--- 2024-02-01 paper/main.tex
@@
 \end{tikzpicture}
-\caption{\textbf{\our's model architecture} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
+\caption{\textbf{Overview of \our's representation learning} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
 We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
-To this end, our normalization results in an implicit codebook whose size we can control.
+Our normalization results in an implicit codebook (whose size we can control) and helps prevent representation collapse.
+  % Importantly, this prevents representation collapse without needing loss terms to minimize errors in reward and/or value predictions.
+  % As such, our representations are task-agnostic.
   We denote our normalization scheme by \cube.}
@@
 
-To the best of our knowledge, all prior works learn representations with objectives that depend on
-predicting the reward and/or value in the latent space
+\section{Introduction}
+\label{intro}
+
+Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+However, applying RL in real-world environments is challenging. The problem is two-fold,
+{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+
+% Given such a latent space,
+% Whilst these latent states are often lower dimensional this is not always
+Many approaches have been proposed to improve sample efficiency in deep RL.
+As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
+In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+In particular, we focus on representation learning methods which compress observations into latent states.
+Nevertheless, our representation learning borrows ideas from world models as it uses multi-step predictions in a
+latent-space dynamics model to learn temporally consist representations \citep{zhaoSimplifiedTemporalConsistency2023}.
+% In particular, methods for compressing observations into latent states.%, aka representation learning.
+
+Learning representations for RL has been investigated for decades
+\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
+However, these approaches are usually limited to simple environments.
+Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
+attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
+% Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
+
+Recently, TD-MPC \citep{hansenTemporalDifferenceLearning2022} and TCRL \citep{zhaoSimplifiedTemporalConsistency2023}
+have obtained state-of-the-art performance on continuous control benchmarks by learning
+representations with self-supervised losses.
+In particular, they leverage the latent-state consistency loss.
+This requires an encoder $e_{\theta}$ which maps observations $o_{t}$ to latent states $z_{t}$ and a transition model $d_{\phi}$ which predicts
+next latent states $z_{t+1}$ given an a latent state $z_{t}$ and an action $a_{t}$.
+It then minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
+and the next state predicted by the target encoder $e_{\bar{\phi}}(z_{t+1})$.
+Self-supervised losses are susceptible to a problem known as representation collapse, where the encoder learns a constant
+representation \cite{jingUnderstandingDimensionalCollapse2021}.
+As such, it is common to combine the latent-state consistency loss with other loss terms, such as
+minimizing the reward prediction error in the latent space
 \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+This helps to prevent representation collapse at the cost of learning a task-specific representation.
+% Typically it is combined with other loss terms such as reward or value prediction.
+
+% To the best of our knowledge, all prior works that learn representations with objectives that depend on
+% predicting the reward and/or value in the latent space
+% \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+% This is usually motivated by preventing representation collapse.
+% To the best of our knowledge, all prior works that learn representations with objectives that depend on
+% predicting the reward and/or value in the latent space
+% \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
+% This is usually motivated by preventing representation collapse.
 % use a latent-state consistency loss based upon the Cosine
@@
 % The cosine similarity loss provides training stability whilst the reward prediction prevents representation collapse.
-As such, these methods learn task-specific representations.
+% However, learning representations using reward and/or value loss terms leads to the representations being task specific.
 

### 2024-02-01 paper/main.tex

--- 2024-02-01 paper/main.tex
@@
 \begin{abstract}
-This document provides a basic paper template and submission guidelines.
-Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-Gross violations will trigger corrections at the camera-ready phase.
+  % Arno's proposal
+  In reinforcement learning (RL) for continuous control, we propose  an efficient representation learning method using a latent-state consistency loss.
+This approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
+By minimizing discrepancies between these states, we achieve task-agnostic representations and prevent representation collapse using only the consistency loss.
+Our method is straightforward, compatible with any model-free RL algorithm, and demonstrates state-of-the-art performance in continuous control benchmarks from DeepMind Control Suite.\looseness-1
+
+  % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
+  % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
+  % is limited by its sample-efficiency.
+  % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
+  % handling complex/high-dimensional observation spaces.
+  Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
+  Not only can it handle complex observation spaces but it can also improve sample efficiency.
+  A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
+  It requires an encoder which maps observations to latent states and a transition model which predicts
+  next latent states.
+  It then minimizes the distance between the next state predicted by the transition model and
+  the next state predicted by the encoder.
+  Typically it is combined with other loss terms such as reward or value prediction.
+  This prevents representation collapse at the cost of learning a task-specific representation.
+  In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
+  a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
+  consistency loss.
+  We empirically show that our method prevents representation collapse as it maintains matrix rank.
+  Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
+  and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
+  % Moreover, it kk
+% This document provides a basic paper template and submission guidelines.
+% Abstracts must be a single paragraph, ideally between 4--6 sentences long.
+% Gross violations will trigger corrections at the camera-ready phase.
 \end{abstract}
-
-\section{Introduction}
-\label{intro}
-
-Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
-However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
-deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
-
-% Given such a latent space,
-% Whilst these latent states are often lower dimensional this is not always
-Many approaches have been proposed to improve sample efficiency in deep RL.
-As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
-\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-In particular, methods for compressing observations into latent states.%, aka representation learning.
-
-Learning representations for RL has been investigated for decades
-\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
-However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
-More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
-attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
@@
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
-It is based solely on a self-predictive loss.
+It is based soley on the latent-state consistency loss, \ie a commonly used self-supervised loss for continuous RL.
 In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
 \citep{mentzerFiniteScalarQuantization2023}.
-As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our representation {\em (i)} significantly speeds up learning, {\em (ii)} is simple to implement,
-and {\em (iii)} is task-agnostic.
+As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
+% As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
+Importantly, our method {\em (i)} prevents representation collapse (as it learns a high-rank representation),
+{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
+{\em (iii)} is simple to implement, and {\em (iii)} learns a task-agnostic
+representation which could be helpful in downstream tasks.
 

### 2024-02-01 paper/main.tex

@@
 \begin{abstract}
-  % Arno's proposal
-  In reinforcement learning (RL) for continuous control, we propose  an efficient representation learning method using a latent-state consistency loss.
-This approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
-By minimizing discrepancies between these states, we achieve task-agnostic representations and prevent representation collapse using only the consistency loss.
-Our method is straightforward, compatible with any model-free RL algorithm, and demonstrates state-of-the-art performance in continuous control benchmarks from DeepMind Control Suite.\looseness-1
-
-  % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
-  % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
-  % is limited by its sample-efficiency.
-  % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
-  % handling complex/high-dimensional observation spaces.
-  Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
-  Not only can it handle complex observation spaces but it can also improve sample efficiency.
-  A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
-  It requires an encoder which maps observations to latent states and a transition model which predicts
-  next latent states.
-  It then minimizes the distance between the next state predicted by the transition model and
-  the next state predicted by the encoder.
-  Typically it is combined with other loss terms such as reward or value prediction.
-  This prevents representation collapse at the cost of learning a task-specific representation.
-  In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
-  a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
-  consistency loss.
-  We empirically show that our method prevents representation collapse as it maintains matrix rank.
-  Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
-  and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
-  % Moreover, it kk
-% This document provides a basic paper template and submission guidelines.
-% Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-% Gross violations will trigger corrections at the camera-ready phase.
+  In reinforcement learning (RL) for continuous control, we propose an
+  efficient representation learning method using only a latent-state
+  consistency loss. Our approach employs an encoder and a transition
+  model to map observations to latent states and predict future
+  states, respectively. We achieve high performance and prevent
+  representation collapse by normalizing, quantizing, and bounding the
+  latent representation. Our method is straightforward, compatible
+  with any model-free RL algorithm, and demonstrates state-of-the-art
+  performance in continuous control benchmarks from DeepMind Control
+  Suite.
 \end{abstract}
+
+%% \begin{abstract}
+%%   Learning representations for reinforcement learning (RL) has shown
+%%   much promise for continuous control. Representations not only can
+%%   handle complex observation spaces but also improve sample
+%%   efficiency. A particularly promising approach is to use a
+%%   self-supervised latent-state consistency loss with observation to
+%%   latent state mapping, together with next latent state, reward, and
+%%   value predictions. Reward and value predictions prevent
+%%   representation collapse at the cost of learning a task-specific
+%%   representation.  We show how to use only the latent-state
+%%   consistency loss to {\em (i)} learn a task-agnostic representation
+%%   and {\em (ii)} prevent representation collapse by normalizing and
+%%   bounding the latent representation. In experiments, our approach
+%%   which be combined with any model-free RL algorithm obtains
+%%   state-of-the-art results in continuous control benchmarks from
+%%   DeepMind Control Suite.
+%% \end{abstract}
+
+%% \begin{abstract}
+%%   % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
+%%   % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
+%%   % is limited by its sample-efficiency.
+%%   % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
+%%   % handling complex/high-dimensional observation spaces.
+%%   Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
+%%   Not only can it handle complex observation spaces but it can also improve sample efficiency.
+%%   % Recent works have shown that learning representations can improve sample efficiency in RL.
+%%   A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
+%%   It requires an encoder which maps observations to latent states and a transition model which predicts
+%%   next latent states.
+%%   It then minimizes the distance between the next state predicted by the transition model and
+%%   the next state predicted by the encoder.
+%%   Typically it is combined with other loss terms such as reward or value prediction.
+%%   This prevents representation collapse at the cost of learning a task-specific representation.
+%%   In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
+%%   a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
+%%   consistency loss.
+%%   % the leverage the latent-state
+%%   % Typically these approaches have unbounded latent spaces and are combined with auxillary loss terms such as minimizing the
+%%   % error of reward or value predictions in the latent space.
+%%   % of an encoder which maps observations to latent states and a transition model which predicts
+%%   % next latent states given a latent state and an action.
+%%   % However, most of th
+%%   We empirically show that our method prevents representation collapse as it maintains matrix rank.
+%%   Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
+%%   and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
+%%   % Moreover, it kk
+%% % This document provides a basic paper template and submission guidelines.
+%% % Abstracts must be a single paragraph, ideally between 4--6 sentences long.
+%% % Gross violations will trigger corrections at the camera-ready phase.
+%% \end{abstract}
+
+\section{Introduction}
+\label{intro}
+
+Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+However, applying RL in real-world environments is challenging. The problem is two-fold,
+{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+
+% Given such a latent space,
+% Whilst these latent states are often lower dimensional this is not always
+Many approaches have been proposed to improve sample efficiency in deep RL.
+As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
+In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+In particular, methods for compressing observations into latent states.%, aka representation learning.
+
+Learning representations for RL has been investigated for decades
+\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
+However, these approaches are usually limited to simple environments.
+Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
+attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
+Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 

### 2024-02-01 paper/main.tex
adding to method

--- 2024-02-01 paper/main.tex
@@
 \begin{abstract}
-  In reinforcement learning (RL) for continuous control, we propose an
-  efficient representation learning method using only a latent-state
-  consistency loss. Our approach employs an encoder and a transition
-  model to map observations to latent states and predict future
-  states, respectively. We achieve high performance and prevent
-  representation collapse by normalizing, quantizing, and bounding the
-  latent representation. Our method is straightforward, compatible
-  with any model-free RL algorithm, and demonstrates state-of-the-art
-  performance in continuous control benchmarks from DeepMind Control
-  Suite.
+  % Arno's proposal
+  In reinforcement learning (RL) for continuous control, we propose  an efficient representation learning method using a latent-state consistency loss.
+This approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
+By minimizing discrepancies between these states, we achieve task-agnostic representations and prevent representation collapse using only the consistency loss.
+Our method is straightforward, compatible with any model-free RL algorithm, and demonstrates state-of-the-art performance in continuous control benchmarks from DeepMind Control Suite.\looseness-1
+
+  % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
+  % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
+  % is limited by its sample-efficiency.
+  % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
+  % handling complex/high-dimensional observation spaces.
+  Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
+  Not only can it handle complex observation spaces but it can also improve sample efficiency.
+  A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
+  It requires an encoder which maps observations to latent states and a transition model which predicts
+  next latent states.
+  It then minimizes the distance between the next state predicted by the transition model and
+  the next state predicted by the encoder.
+  Typically it is combined with other loss terms such as reward or value prediction.
+  This prevents representation collapse at the cost of learning a task-specific representation.
+  In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
+  a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
+  consistency loss.
+  We empirically show that our method prevents representation collapse as it maintains matrix rank.
+  Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
+  and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
+  % Moreover, it kk
+% This document provides a basic paper template and submission guidelines.
+% Abstracts must be a single paragraph, ideally between 4--6 sentences long.
+% Gross violations will trigger corrections at the camera-ready phase.
 \end{abstract}
-
-%% \begin{abstract}
-%%   Learning representations for reinforcement learning (RL) has shown
-%%   much promise for continuous control. Representations not only can
-%%   handle complex observation spaces but also improve sample
-%%   efficiency. A particularly promising approach is to use a
-%%   self-supervised latent-state consistency loss with observation to
-%%   latent state mapping, together with next latent state, reward, and
-%%   value predictions. Reward and value predictions prevent
-%%   representation collapse at the cost of learning a task-specific
-%%   representation.  We show how to use only the latent-state
-%%   consistency loss to {\em (i)} learn a task-agnostic representation
-%%   and {\em (ii)} prevent representation collapse by normalizing and
-%%   bounding the latent representation. In experiments, our approach
-%%   which be combined with any model-free RL algorithm obtains
-%%   state-of-the-art results in continuous control benchmarks from
-%%   DeepMind Control Suite.
-%% \end{abstract}
-
-%% \begin{abstract}
-%%   % Deep reinforcement learning (RL) has shown much promise for solving real-world continuous control problems.
-%%   % Whilst deep reinforcement learning (RL) is a competitive approach for solving real-world continuous control problems, it
-%%   % is limited by its sample-efficiency.
-%%   % Recent works have shown that learning representations can improve sample efficiency in RL, whilst also
-%%   % handling complex/high-dimensional observation spaces.
-%%   Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
-%%   Not only can it handle complex observation spaces but it can also improve sample efficiency.
-%%   % Recent works have shown that learning representations can improve sample efficiency in RL.
-%%   A particularly promising approach is to use a self-supervised loss, known as the latent-state consistency loss.
-%%   It requires an encoder which maps observations to latent states and a transition model which predicts
-%%   next latent states.
-%%   It then minimizes the distance between the next state predicted by the transition model and
-%%   the next state predicted by the encoder.
-%%   Typically it is combined with other loss terms such as reward or value prediction.
-%%   This prevents representation collapse at the cost of learning a task-specific representation.
-%%   In this work, we show that a careful consideration of the latent space enables us to {\em (i)} learn
-%%   a task-agnostic representation and {\em (ii)} prevent representation collapse, whilst using only the latent-state
-%%   consistency loss.
-%%   % the leverage the latent-state
-%%   % Typically these approaches have unbounded latent spaces and are combined with auxillary loss terms such as minimizing the
-%%   % error of reward or value predictions in the latent space.
-%%   % of an encoder which maps observations to latent states and a transition model which predicts
-%%   % next latent states given a latent state and an action.
-%%   % However, most of th
-%%   We empirically show that our method prevents representation collapse as it maintains matrix rank.
-%%   Our representation learning approach is {\em (i)} simple, {\em (ii)} can be combined with any model-free RL algorithm
-%%   and, {\em (iii)} obtains state-of-the-art results in continuous control benchmarks from DeepMind Control Suite.
-%%   % Moreover, it kk
-%% % This document provides a basic paper template and submission guidelines.
-%% % Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-%% % Gross violations will trigger corrections at the camera-ready phase.
-%% \end{abstract}
-
-\section{Introduction}
-\label{intro}
-
-Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
-However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
-deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
-
-% Given such a latent space,
-% Whilst these latent states are often lower dimensional this is not always
-Many approaches have been proposed to improve sample efficiency in deep RL.
-As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
-\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-In particular, methods for compressing observations into latent states.%, aka representation learning.
-
-Learning representations for RL has been investigated for decades
-\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
-However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
-More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
-attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
@@
 % As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our method {\em (i)} prevents representation collapse (as it learns a high-rank representation),
+Importantly, our method {\em (i)} alleviates representation collapse,
 {\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
-{\em (iii)} is simple to implement, and {\em (iii)} learns a task-agnostic
+{\em (iii)} is simple to implement, and {\em (iv)} learns a task-agnostic
 representation which could be helpful in downstream tasks.
@@
 \end{align}
-Central to our method is the latent state $z_{t}$
+The encoder $e_{\theta}$  and latent-space dynamics model $d_{\phi}$ are responsible for representation learning.
+The encoder $e_{\theta}$ maps observations $o_{t}$ to latent states $z_{t}$ and is responsible for learning a representation
+which can aid RL.
+The latent-space dynamics model $d_{\phi}$ predicts next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
+It's sole purpose is to aid representation learning by making the latent states temporally consistent.
+Note that we do not use it for model-based RL.
+Once we have the representation learned by our encoder, we map all observations to the latent space and perform model-free RL
+in this latent space.
+Throughout this paper, we use Twin Delayed Deep Deterministic Policy Gradient
+(TD3), \citep{fujimotoAddressingFunctionApproximation2018}).
+It consists of two state-action value functions $\{Q_{\psi_{1}},Q_{\psi_{2}} \}$, known as critics, and a deterministic
+actor $\pi_{\eta}$.
+Following prior works \citep{yaratsMasteringVisualContinuous2021} we use a linear exploration noise schedule
+which decays from $1$ to $0.1$ during training.
 

### 2024-02-01 paper/figs/plotting/rank-comparison/fetch_data.py
Rank plots

@@ -1,0 +1,172 @@
+# %%
+ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_data = {
+    "path": ddpg_path,
+    "data": {
+        # 'acrobot-swingup': ['1eiys0vf', '1yedk0zx', '1pbg10k5', 'zgb6ndd1', '2dh8jyh9',],
+        # "cheetah-run": ["6bzzn892"],  # "lcmbhw37"],
+        # "humanoid-walk": ["lcmbhw37"],  # "lcmbhw37"],
+        # "humanoid-run": ["jv7cfy2v"],
+        # 'fish-swim': ['1zrla7u4', '3rx4mc18', 'xwq0fu1r', 'smv9miz4', 'nu50jdp0', ],
+        # "cartpole-swingup": ["w6eykaow"],  # UTD=1
+        # "hopper-stand": ["3ozramwe"],  # UTD=1
+        # "quadruped-walk": ["8q59zkws"],  # UTD=1
+        # "cheetah-run": ["mkmvf5md"],  # UTD=4
+        # : [,],
+        "cheetah-run-d128-no-norm": [
+            "5j2it0tc",
+            "v0fz1014",
+            "v1i6mftd",
+            "yfkyy6rt",
+            "txtebz1w",
+        ],
+        "cheetah-run-d128-norm": [
+            "l4c3phty",
+            "96xm0zx5",
+            "6tul0lmk",
+            "y7c9k82p",
+            "zx6ykgj9",
+        ],
+        "hopper-stand-d=64-no-norm": [
+            "0ymb3imw",
+            "5fibx0fg",
+            "i8ou5zx4",
+            "bijgsw6i",
+            "02uvlbjb",
+        ],
+        "hopper-stand-d=64": [
+            "b978u8ed",
+            "77bswwl0",
+            "rkdmf32p",
+            "ucvgujrp",
+            "zqv19l3p",
+        ],
+        "humanoid-run": [
+            # no-norm d=1024 project=False
+            "uuyjutba",
+            "f1wdhmr1",
+            "cbrtltx4",
+            "arilzq7a",
+            "ua5c03do",
+            # norm, d=1024 projection=False
+            "dt0r1wyb",
+            "g35s9yyg",
+            "fuqpil5h",
+            "15i70x9w",
+            "agjdi84a",
+        ],
+        "acrobot-swingup": [
+            # no norm
+            "vkbkcrjm",
[…]

### 2024-02-01 paper/main.tex
adding more representation stuff

@@
 
-\our has four parameterized components.
+\textbf{Method overview}
+\our has four main components which we wish to learn:
 \begin{align}
@@
 &\text{Value: } & q_{t} &= Q_{\psi} (z_{t}, a_{t}) \label{eq:value} \\
-&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) . \label{eq:policy}
+&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) . \label{eq:policy}.
 \end{align}
@@
 
-\textbf{Bounding the latent space}
-Typically the latent state $z_{t} \in \R^{d}$
-However, in our work we bound the domain of the latent space, \ie $z_{t} \in [-L, L]^{d}$.
-
 \textbf{Representation learning}
-Our representation learning is based solely on the temporal consistency loss,
-However, unlike previous works, we normalize the latent vector
-Our representation learning consists of an encoder $e_{\theta}$ (which maps observations to
-latent states) and a transition model $d_{\phi}$ which maps
-Our representation learning adopts a self-supervised loss.
+Our representation learning is based solely on the latent-state consistency loss,
 %
@@
 %
+which minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
+and the next state predicted by the momentum encoder $e_{\bar{\phi}}(z_{t+1})$.
+The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is trained jointly with the transition model.
+The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
+of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
+The target network update rate is denoted $\tau$.
+Unlike previous works, we normalize the latent vector
+%
 \textbf{Finite Scalar Quantization}
@@
 where each entry in  $f : z \rightarrow \left[ L/2 \right] \mathrm{tanh}(z))$
+
+\textbf{Bounding the latent space}
+Typically the latent state $z_{t} \in \R^{d}$
+However, in our work we bound the domain of the latent space, \ie $z_{t} \in [-L, L]^{d}$.
 

### 2024-02-01 paper/main.tex
udpated
Update to abstract

@@
   states, respectively. We achieve high performance and prevent
-  representation collapse by normalizing, quantizing, and bounding the
-  latent representation. Our method is straightforward, compatible
-  with any model-free RL algorithm, and demonstrates state-of-the-art
+  representation collapse by normalizing the quantized latent
+  representation such that rank in the presentation is empirically
+  preserved. Our method is straightforward, compatible with any
+  model-free RL algorithm, and demonstrates state-of-the-art
   performance in continuous control benchmarks from DeepMind Control

### 2024-02-02 paper/main.tex
Matrix rank

--- 2024-02-01 paper/main.tex
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Sample-efficient Reinforcement Learning with Implicitly Quantized Representations}
+\icmltitlerunning{Sample-Efficient Reinforcement Learning with Implicitly Quantized Representations}
 
@@
 
+We analyze the impact of the latent-space normalization on the rank of the latent representations outputted by the learned encoder.
+We follow the methodology of \citet{ni2024bridging} for estimating the matrix rank.
+We plot the ranks in \cref{fig:rank-comparison-z1}.
+The latent state dimensions are 256, 128, 256, and 128 for Acrobot Swingup, Hopper Stand, Humanoid Run and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
+Without normalization, a dimensional collapse occurs, which can have significant harmful effects as the representational power of the latent state diminishes \cite{jingUnderstandingDimensionalCollapse2021}.
+
 \begin{figure*}[ht]
@@
 	\begin{center}
-		\centerline{\includegraphics[width=\textwidth]{./figs/task-agnostic-ablation.pdf}}
-		\caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
-We evaluate whether pre-training \our on the walk tasks in the DeepMind Control Suite help for learning the run task on the same environment.
-The pre-training proves to be beneficial on all evaluated environments.
-If the representation is task-specific (the reward prediction term is included), the pre-training can be clearly detrimental.
+		\centerline{\includegraphics[width=1.0\textwidth]{./figs/rank-comparison-z1.pdf}}
+		\caption{\textbf{Representation collapse without normalization}.
+We observe that the representation normalization of \our is crucial for avoiding representation collapse.
+When no normalization is applied to the latents, the matrix rank of the representations drops rapidly, which can have a harmful effect on the downstream performance.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-		\label{fig:multi-task-pretraining}
+		\label{fig:rank-comparison-z1}
 	\end{center}
@@
 \label{fig:reward-ablation}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\begin{figure*}[ht]
+	\vskip 0.2in
+	\begin{center}
+		\centerline{\includegraphics[width=\textwidth]{./figs/task-agnostic-ablation.pdf}}
+		\caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
+We evaluate whether pre-training \our on the walk tasks in the DeepMind Control Suite help for learning the run task on the same environment.
+The pre-training proves to be beneficial on all evaluated environments.
+If the representation is task-specific (the reward prediction term is included), the pre-training can be clearly detrimental.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+		\label{fig:multi-task-pretraining}
 	\end{center}
@@
 \label{conclusion}
+
+\section*{Broader Impact}
+This paper presents work whose goal is to advance the field of continuous RL that is applicable to, for instance, robotics.
+Therefore, it is possible that our methods could be employed in unforeseen harmful ways, such as autonomous drones for military or law enforcement use.
+There are even other potential societal consequences of our work, none which we feel must be specifically highlighted here.
 

### 2024-02-02 paper/main.tex

@@
 % \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
-&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2}
+&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
 \end{align}
 %
+which minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
+and the next state predicted by the momentum encoder $e_{\bar{\phi}}(z_{t+1})$.
+The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
+The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
+of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
+The target network update rate is denoted $\tau$.
+Unlike previous works, we normalize the latent vector
+
 \textbf{Finite Scalar Quantization}

### 2024-02-02 paper/main.tex
updating mehtods

--- 2024-02-01 paper/main.tex
@@
   states, respectively. We achieve high performance and prevent
-  representation collapse by normalizing the quantized latent
-  representation such that rank in the presentation is empirically
-  preserved. Our method is straightforward, compatible with any
-  model-free RL algorithm, and demonstrates state-of-the-art
+  representation collapse by normalizing, quantizing, and bounding the
+  latent representation. Our method is straightforward, compatible
+  with any model-free RL algorithm, and demonstrates state-of-the-art
   performance in continuous control benchmarks from DeepMind Control
@@
 % \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
-&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2}
+&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
 \end{align}
@@
 and the next state predicted by the momentum encoder $e_{\bar{\phi}}(z_{t+1})$.
-The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is trained jointly with the transition model.
+The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
 The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
@@
 The target network update rate is denoted $\tau$.
+
 Unlike previous works, we normalize the latent vector
-%
+
 \textbf{Finite Scalar Quantization}
@@
 \textbf{Policy and value function learning}
-
+We learn the policy and value function using TD3, \citep{fujimotoAddressingFunctionApproximation2018})
+and follow \cite{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
+by augmenting the loss with n-step returns.
+The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
+encoder $z_{t} e_{\theta}(o_{t})$ and learn the value function and policy in the latent space.
+The value function is updated by minimizing the following objective:
 \begin{align} \label{eq:value-loss}
-  \mathcal{L}_{Q}(\psi ; \mathcal{D}) &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad \forall k \in 1, 2 \\
-  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n})
+  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
+  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n}), \nonumber
 \end{align}
+where the action is
 
 \begin{align} \label{eq:policy-loss}
- \mathcal{L}_{\pi}(\eta; \mathcal{D}) = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
+ \mathcal{L}_{\pi} = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
 \end{align}

### 2024-02-02 paper/main.tex

--- 2024-02-02 paper/main.tex
@@
 The target network update rate is denoted $\tau$.
+
 Unlike previous works, we normalize the latent vector

### 2024-02-02 paper/main.tex
Capitalization and shorter title

--- 2024-02-02 paper/main.tex
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Sample-Efficient Reinforcement Learning with Implicitly Quantized Representations}
+\icmltitlerunning{Sample-Efficient RL with Implicitly Quantized Representations}
 
@@
 % \icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
-\icmltitle{Sample-efficient Reinforcement Learning \\ with Implicitly Quantized Representations}
+\icmltitle{Sample-Efficient Reinforcement Learning \\ with Implicitly Quantized Representations}
 

### 2024-02-02 paper/main.tex
Equation --> cref eq

@@
 \end{tikzpicture}
-\caption{\textbf{Overview of \our's representation learning} \our learns a representation soley with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
+\caption{\textbf{Overview of \our's representation learning} \our learns a representation solely with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
 We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
@@
 \textbf{Reward predict is not necessary for representation learning}
-Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (Equation~\ref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
+Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (\cref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
 Instead, we rely solely on the temporal consistency loss.
@@
 \begin{itemize}
-	\item \ourrew, where we add a reward prediction head into our model architecture by re-defining the dynamics as $\hat{z}_{t+1}, \hat{r}_t = d_{\phi} (z_{t}, a_{t})$ (see Equation~\ref{eq:transition}), and include a reward prediction term (discounted MSE Loss) in the representation loss:
+	\item \ourrew, where we add a reward prediction head into our model architecture by re-defining the dynamics as $\hat{z}_{t+1}, \hat{r}_t = d_{\phi} (z_{t}, a_{t})$ (see \cref{eq:transition}), and include a reward prediction term (discounted MSE Loss) in the representation loss:
 	\begin{align*}
@@
 	where $\tilde{r}_{t+h}$ is the ground-truth H-step reward and $\hat{r}_{t+h}$ the predicted H-step rewards.
-	\item \ourcosrew, where in addition to the reward prediction head in \ourrew, we additionally convert the self-supervised MSE loss (see Equation~\ref{eq:rep-loss}) for the temporal consistency into a cosine similarity, similarly to TCRL:
+	\item \ourcosrew, where in addition to the reward prediction head in \ourrew, we additionally convert the self-supervised MSE loss (see \cref{eq:rep-loss}) for the temporal consistency into a cosine similarity, similarly to TCRL:
 	\begin{align*} \label{eq:rep-loss}
@@
 \end{itemize}
-The results for this ablation study are shown in Figure~\ref{fig:reward-ablation}. 
+The results for this ablation study are shown in \cref{fig:reward-ablation}. 
 The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods with the exception of Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
@@
 \textbf{\our's task-agnostic representation is helpful for multi-task RL}
-In contrast to TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (see Equation~\ref{eq:rep-loss}).
+In contrast to TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (\cref{eq:rep-loss}).
 The potential upside of \our's task-agnostic representations is that they can be used to speed-up learning in the incremental multi-task setting.
@@
 We compare the pre-trained agents to training \our on these tasks from scratch.
-The results for this experiment can be found in Figure~\ref{fig:multi-task-pretraining}.
+The results for this experiment can be found in \cref{fig:multi-task-pretraining}.
 

### 2024-02-02 paper/main.tex
udpated

--- 2024-02-02 paper/main.tex
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Sample-Efficient RL with Implicitly Quantized Representations}
+\icmltitlerunning{Sample-Efficient Reinforcement Learning with Implicitly Quantized Representations}
 
@@
 % \icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
-\icmltitle{Sample-Efficient Reinforcement Learning \\ with Implicitly Quantized Representations}
+\icmltitle{Sample-efficient Reinforcement Learning \\ with Implicitly Quantized Representations}
 
@@
 \textbf{Policy and value function learning}
-
+We learn the policy and value function using TD3 \citep{fujimotoAddressingFunctionApproximation2018}.
+However, we follow \citet{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
+and augment the loss with n-step returns.
+The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
+encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
+The value function is updated by minimizing the following objective:
 \begin{align} \label{eq:value-loss}
-  \mathcal{L}_{Q}(\psi ; \mathcal{D}) &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad \forall k \in 1, 2 \\
-  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n})
+  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\bar{\theta}}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
+  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n}), \nonumber
 \end{align}
+where we use policy smoothing by adding clipped Gaussian noise $\epsilon \sim \mathcal{N} (0,\sigma^{2})$ to the
+action $a_{t+n} = \pi_{\eta}(z_{t+n}) + \epsilon$.
+Note that we use the target encoder to get the latent states in both the prediction and the target.
 
 \begin{align} \label{eq:policy-loss}
- \mathcal{L}_{\pi}(\eta; \mathcal{D}) = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
+ \mathcal{L}_{\pi} = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
 \end{align}

### 2024-02-02 paper/main.tex

--- 2024-02-02 paper/main.tex
@@
 \textbf{Policy and value function learning}
-We learn the policy and value function using TD3, \citep{fujimotoAddressingFunctionApproximation2018})
-and follow \cite{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
-by augmenting the loss with n-step returns.
+We learn the policy and value function using TD3 \citep{fujimotoAddressingFunctionApproximation2018}.
+However, we follow \citet{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
+and augment the loss with n-step returns.
 The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
-encoder $z_{t} e_{\theta}(o_{t})$ and learn the value function and policy in the latent space.
+encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
 The value function is updated by minimizing the following objective:
 \begin{align} \label{eq:value-loss}
-  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
+  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\bar{\theta}}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
   y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n}), \nonumber
 \end{align}
-where the action is
+where we use policy smoothing by adding clipped Gaussian noise $\epsilon \sim \mathcal{N} (0,\sigma^{2})$ to the
+action $a_{t+n} = \pi_{\eta}(z_{t+n}) + \epsilon$.
+Note that we use the target encoder to get the latent states in both the prediction and the target.
 

### 2024-02-02 paper/main.tex
added more to method

@@
 \end{align}
-where we use policy smoothing by adding clipped Gaussian noise $\epsilon \sim \mathcal{N} (0,\sigma^{2})$ to the
-action $a_{t+n} = \pi_{\eta}(z_{t+n}) + \epsilon$.
+where we use policy smoothing by adding clipped Gaussian noise $\epsilon_{t+n} \sim \mathcal{N} (0,\sigma^{2})$ to the
+action $a_{t+n} = \pi_{\bar{\eta}}(z_{t+n}) + \epsilon_{t+n}$.
 Note that we use the target encoder to get the latent states in both the prediction and the target.
-
+We then use the target value functions $Q_{\bar{\psi}_{k}}$ and the target policy $\pi_{\bar{\eta}}$ to calculate the TD target.
+Following TD3, we learn the actor's parameters by minimizing,
+%
 \begin{align} \label{eq:policy-loss}
- \mathcal{L}_{\pi} = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
+ \mathcal{L}_{\pi} = - \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\bar{\theta}}(o_{t}), a_{t}) \right].
 \end{align}
-
-Whilst our method resembles TCRL \citep{zhaoSimplifiedTemporalConsistency2023} it is important to note that
-our transition model does not predict the reward.
-As such, our method learns a task-agnostic representation.
+%
+That is, maximizing the Q-value using the clipped double Q learning trick to combat overestimation in Q learning.
+
+% Whilst our method resembles TCRL \citep{zhaoSimplifiedTemporalConsistency2023} it is important to note that
+% our transition model does not predict the reward.
+% As such, our method learns a task-agnostic representation.
 

### 2024-02-02 paper/main.tex
updated abstract

@@
 \begin{abstract}
-  In reinforcement learning (RL) for continuous control, we propose an
-  efficient representation learning method using only a latent-state
-  consistency loss. Our approach employs an encoder and a transition
-  model to map observations to latent states and predict future
-  states, respectively. We achieve high performance and prevent
-  representation collapse by normalizing the quantized latent
-  representation such that rank in the presentation is empirically
-  preserved. Our method is straightforward, compatible with any
-  model-free RL algorithm, and demonstrates state-of-the-art
-  performance in continuous control benchmarks from DeepMind Control
-  Suite.
+  Learning representations for reinforcement learning (RL) has shown much promise for continuous control.
+We propose an efficient representation learning method using only a self-supervised latent-state consistency loss.
+Our approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
+We achieve high performance and prevent representation collapse by normalizing and bounding the latent representation such that rank in the representation is empirically preserved.
+Our method is straightforward, compatible with any model-free RL algorithm, and demonstrates state-of-the-art performance in continuous control benchmarks from DeepMind Control Suite.
 \end{abstract}
+
+% \begin{abstract}
+%   In reinforcement learning (RL) for continuous control, we propose an
+%   efficient representation learning method using only a latent-state
+%   consistency loss. Our approach employs an encoder and a transition
+%   model to map observations to latent states and predict future
+%   states, respectively. We achieve high performance and prevent
+%   representation collapse by normalizing the quantized latent
+%   representation such that rank in the presentation is empirically
+%   preserved. Our method is straightforward, compatible with any
+%   model-free RL algorithm, and demonstrates state-of-the-art
+%   performance in continuous control benchmarks from DeepMind Control
+%   Suite.
+% \end{abstract}
 

### 2024-02-02 paper/main.tex
comment duplication of old intro
Updated texts

--- 2024-02-02 paper/main.tex
@@
 \textbf{Policy and value function learning}
-We learn the policy and value function using TD3 \citep{fujimotoAddressingFunctionApproximation2018}.
-However, we follow \citet{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
-and augment the loss with n-step returns.
+We learn the policy and value function using TD3, \citep{fujimotoAddressingFunctionApproximation2018})
+and follow \cite{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
+by augmenting the loss with n-step returns.
 The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
-encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
+encoder $z_{t} e_{\theta}(o_{t})$ and learn the value function and policy in the latent space.
 The value function is updated by minimizing the following objective:
 \begin{align} \label{eq:value-loss}
-  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\bar{\theta}}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
+  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
   y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n}), \nonumber
 \end{align}
-where we use policy smoothing by adding clipped Gaussian noise $\epsilon \sim \mathcal{N} (0,\sigma^{2})$ to the
-action $a_{t+n} = \pi_{\eta}(z_{t+n}) + \epsilon$.
-Note that we use the target encoder to get the latent states in both the prediction and the target.
+where the action is
 
@@
 \label{sec:experiments}
-In this section, we evaluate \our in a wide variety of continuous control tasks from the DeepMind Control (DMC) suite.
+In this section, we evaluate \our in a wide variety of continuous control tasks from the DeepMind Control (DMC) suite \cite{tassa2018deepmind}.
 We aim to answer the following questions:
@@
 \textbf{Normalizing the latent-space improves sample efficiency}
-In \cref{fig:normalization_improves_sample_efficiency},
+In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on eight problems from the DeepMind Control Suite.
+The results show that our method, \our, performs on par with TD-MPC while being significantly simpler and quicker to train, as the loss term is simpler and no planning is needed unlike in TD-MPC.
+We even outperform TCRL, which is the most similar baseline to our work.
+We ablate the effect of the FSQ-based latent-space normalization scheme in \cref{fig:normalization-ablation}.
+The normalization is overall beneficial.
 
@@
 \caption{\textbf{Ablation of normalization}.
-\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
 \label{fig:normalization-ablation}
@@
 \textbf{Reconstruction loss has a detrimental impact}
-Recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
+Learning to minimize the observation reconstruction error has been widely applied in model-based reinforcement learning \cite{sutton2018reinforcement, ha2018recurrent, hafnerLearning2019}, and an observation decoder has been a component of the most successful RL algorithms to date \cite{hafner2023mastering}.
+However, recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
 has shown that incorporating a reconstruction term into the representation loss can have a negative impact on performance.
-To provide a thorough analysis of \our, we include results where we add a reconstruction term to our
-representation loss \cref{eq:rep-loss}.
+They hypothesize that learning to reconstruct the observations is inefficient as they can have irrelevant details and visuals like shading that are uncontrollable by the agent and do not affect the tasks.
+
+To provide a thorough analysis of \our, we include results where we add a reconstruction term to our representation loss \cref{eq:rep-loss}:
+\begin{align*}
+	\hat{o}_t &= g_\xi(z_t), \\
+	\mathcal{L}_o &= \mathbb{E}_{o_t\sim\mathcal{D}} [\| \hat{o}_t - o_t \|_2^2],
+\end{align*}
+where $g_\xi$ is a learned observation decoder that takes the latent state as the input and outputs the reconstructed observation.
+The decoder $g_\xi$ is a standard MLP.
 We perform reconstruction at each time step in the horizon.
-\cref{fig:multi-task-pretraining} shows that in no environments does reconstruction aid learning.
-Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on
-overall performance.
+The results in \cref{fig:multi-task-pretraining} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
+%Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on
+%overall performance.
 
@@
 \caption{\textbf{Reconstruction loss has detrimental impact}.
-\our XXXXXX We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+Unlike many methods, such as Dreamer-v3 \cite{hafner2023mastering}, \our neither has an observation decoder nor a reconstruction term in the loss function.
+We show that adding a reconstruction loss harms the performance of \our across all evaluation environments.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
 \label{fig:multi-task-pretraining}

### 2024-02-02 paper/main.tex

@@
 
-\section{Introduction}
-\label{intro}
-
-Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
-However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
-deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
-
-% Given such a latent space,
-% Whilst these latent states are often lower dimensional this is not always
-Many approaches have been proposed to improve sample efficiency in deep RL.
-As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
-\citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-In particular, methods for compressing observations into latent states.%, aka representation learning.
-
-Learning representations for RL has been investigated for decades
-\citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
-However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
-\citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
-More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
-attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
-Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
+% \section{Introduction}
+% \label{intro}
+
+% Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+% However, applying RL in real-world environments is challenging. The problem is two-fold,
+% {\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+% deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+
+% % Given such a latent space,
+% % Whilst these latent states are often lower dimensional this is not always
+% Many approaches have been proposed to improve sample efficiency in deep RL.
+% As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
+% {\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+% \citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
+% In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+% In particular, methods for compressing observations into latent states.%, aka representation learning.
+
+% Learning representations for RL has been investigated for decades
+% \citep{abelOptimalBehaviorApproximate2016,mannorDynamicAbstractionReinforcement2004,liUnifiedTheoryState2006,andreStateAbstractionProgrammable2002,deardenAbstractionApproximateDecisiontheoretic1997,singhReinforcementLearningSoft1994,higginsDefinitionDisentangledRepresentations2018,vanhoofStableReinforcementLearning2016,watterEmbedControlLocally2015,ghoshRepresentationsStableOffPolicy2020}.
+% However, these approaches are usually limited to simple environments.
+% Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+% \citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
+% More recently, self-supervised learning (SSL) approaches (which do not reconstruct observations)
+% attempt to learn good features without labels \cite{anandUnsupervisedStateRepresentation2019}.
+% Alternative approaches leverage contrastive losses to learn representations \cite{laskinCURLContrastiveUnsupervised2020}.
 
@@
 \textbf{Policy and value function learning}
-We learn the policy and value function using TD3, \citep{fujimotoAddressingFunctionApproximation2018})
-and follow \cite{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
-by augmenting the loss with n-step returns.
+We learn the policy and value function using TD3 \citep{fujimotoAddressingFunctionApproximation2018}.
+However, we follow \citet{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
+and augment the loss with n-step returns.
 The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
-encoder $z_{t} e_{\theta}(o_{t})$ and learn the value function and policy in the latent space.
+encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
 The value function is updated by minimizing the following objective:
 \begin{align} \label{eq:value-loss}
-  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
+  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\bar{\theta}}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
   y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n}), \nonumber
 \end{align}
-where the action is
-
+where we use policy smoothing by adding clipped Gaussian noise $\epsilon_{t+n} \sim \mathcal{N} (0,\sigma^{2})$ to the
+action $a_{t+n} = \pi_{\bar{\eta}}(z_{t+n}) + \epsilon_{t+n}$.
+Note that we use the target encoder to get the latent states in both the prediction and the target.
+We then use the target value functions $Q_{\bar{\psi}_{k}}$ and the target policy $\pi_{\bar{\eta}}$ to calculate the TD target.
+Following TD3, we learn the actor's parameters by minimizing,
+%
 \begin{align} \label{eq:policy-loss}
- \mathcal{L}_{\pi} = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
+ \mathcal{L}_{\pi} = - \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\bar{\theta}}(o_{t}), a_{t}) \right].
 \end{align}
-
-Whilst our method resembles TCRL \citep{zhaoSimplifiedTemporalConsistency2023} it is important to note that
-our transition model does not predict the reward.
-As such, our method learns a task-agnostic representation.
+%
+That is, maximizing the Q-value using the clipped double Q learning trick to combat overestimation in Q learning.
+
+% Whilst our method resembles TCRL \citep{zhaoSimplifiedTemporalConsistency2023} it is important to note that
+% our transition model does not predict the reward.
+% As such, our method learns a task-agnostic representation.
 

### 2024-02-02 paper/figs/plotting/reward-ablation/fetch_data.py
Updated reward ablation

@@ -333,7 +333,6 @@ def fetch_results(run_path, run_name_list, keys, agent_name=None):
     data = []
     for run_name in run_name_list:
         if run_name in ["3duujm1p", "kl6xd92o", "s4t5viko", "flcmnkt7", "1zjl534e"]:
-            print("Tulee")
             wandb_run = api.run("aalto-ml/lifelong-td3-tc/" + run_name)
         else:
             wandb_run = api.run(run_path + run_name)

### 2024-02-02 paper/main.tex
moved stuff
moved another figure
Related work, fixed a reference

@@
 
-In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method
-(we use TD3 \citep{fujimotoAddressingFunctionApproximation2018}).
-See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
-It is based soley on the latent-state consistency loss, \ie a commonly used self-supervised loss for continuous RL.
-In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
-\citep{mentzerFiniteScalarQuantization2023}.
-As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
-% As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our method {\em (i)} alleviates representation collapse,
-{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
-{\em (iii)} is simple to implement, and {\em (iv)} learns a task-agnostic
-representation which could be helpful in downstream tasks.
-
-\section{Method}
-\label{sec:method}
-
-In this section, we detail our method, titled \textit{implicitly Quantized Reinforcement Learning} (\our).
-\our is conceptually simple, it (1) learns a representation of the observation space and then,
-(2) performs model-free RL (e.g. TD3) on this representation.
-% It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
-% to a latent space $z \in \mathcal{Z}$.
-% It first learns a representation of the observation space and then uses
-See \cref{fig:diagram-self-predictive} and \cref{alg:main_alg}.
-
-We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
-where an agent receives an observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.
-The agent then obtains a reward $r_{t} = \mathcal{R} (o_{t}, a_{t})$ and obtains the next observation
-$o_{t+1} = P(\cdot \mid o_{t}, a_{t})$.
-The discount factor is denoted $\gamma \in [0, 1]$.
-
 \begin{algorithm}[tb]
@@
 
+In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method
+(we use TD3 \citep{fujimotoAddressingFunctionApproximation2018}).
+See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
+It is based soley on the latent-state consistency loss, \ie a commonly used self-supervised loss for continuous RL.
+In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
+\citep{mentzerFiniteScalarQuantization2023}.
+As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
+% As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
+Importantly, our method {\em (i)} alleviates representation collapse,
+{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
+{\em (iii)} is simple to implement, and {\em (iv)} learns a task-agnostic
+representation which could be helpful in downstream tasks.
+
+\section{Related Work}
+\label{sec:related_work}
+
+\section{Method}
+\label{sec:method}
+
+In this section, we detail our method, titled \textit{implicitly Quantized Reinforcement Learning} (\our).
+\our is conceptually simple, it (1) learns a representation of the observation space and then,
+(2) performs model-free RL (e.g. TD3) on this representation.
+% It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
+% to a latent space $z \in \mathcal{Z}$.
+% It first learns a representation of the observation space and then uses
+See \cref{fig:diagram-self-predictive} and \cref{alg:main_alg}.
+
+We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
+where an agent receives an observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.
+The agent then obtains a reward $r_{t} = \mathcal{R} (o_{t}, a_{t})$ and obtains the next observation
+$o_{t+1} = P(\cdot \mid o_{t}, a_{t})$.
+The discount factor is denoted $\gamma \in [0, 1]$.
+
+\begin{table}[t]
+\caption{FSQ levels $\mathcal{L}$ to approximate different codebook sizes $|\mathcal{C}|$}
+\label{sample-table}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lccc}
+\toprule
+Target size $|\mathcal{C}|$ & $2^{8}$ & $2^{10}$ & $2^{12}$ \\
+\midrule
+Proposed $\mathcal{L}$ & $[8,6,5]$ & $[8,5,5,5]$ & $[7,5,5,5,5]$ \\
+% \bottomrule
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+
 \textbf{Method overview}
@@
 
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/baselines_comparison.pdf}}
+\caption{\textbf{Normalizing the latent-space improves sample efficiency} \our is significanly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
+It performs similarly to TD-MPC (orange) whilst being significantly simpler and quicker to train.
+\our even outperforms TCRL, which is the most similar baseline.
+Results are for eight DMC tasks with UTD=1.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:normalization_improves_sample_efficiency}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
 \textbf{Representation learning}
@@
 % i=1 Li
-
-\begin{table}[t]
-\caption{FSQ levels $\mathcal{L}$ to approximate different codebook sizes $|\mathcal{C}|$}
-\label{sample-table}
-\vskip 0.15in
-\begin{center}
-\begin{small}
-\begin{sc}
-\begin{tabular}{lccc}
-\toprule
-Target size $|\mathcal{C}|$ & $2^{8}$ & $2^{10}$ & $2^{12}$ \\
-\midrule
-Proposed $\mathcal{L}$ & $[8,6,5]$ & $[8,5,5,5]$ & $[7,5,5,5,5]$ \\
-% \bottomrule
-\end{tabular}
-\end{sc}
-\end{small}
-\end{center}
-\vskip -0.1in
-\end{table}
 
@@
 % In contrast, when increasing the UTD ratio for TCRL no improvements are observed.
-
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/baselines_comparison.pdf}}
-\caption{\textbf{Normalizing the latent-space improves sample efficiency} \our is significanly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
-It performs similarly to TD-MPC (orange) whilst being significantly simpler and quicker to train.
-\our even outperforms TCRL, which is the most similar baseline.
-Results are for eight DMC tasks with UTD=1.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:normalization_improves_sample_efficiency}
-\end{center}
-\vskip -0.2in
-\end{figure*}
 
@@
 \textbf{Reconstruction loss has a detrimental impact}
-Learning to minimize the observation reconstruction error has been widely applied in model-based reinforcement learning \cite{sutton2018reinforcement, ha2018recurrent, hafnerLearning2019}, and an observation decoder has been a component of the most successful RL algorithms to date \cite{hafner2023mastering}.
+Learning to minimize the observation reconstruction error has been widely applied in model-based reinforcement learning \cite{sutton2018reinforcement, haRecurrentWorldModels2018, hafnerLearning2019}, and an observation decoder has been a component of the most successful RL algorithms to date \cite{hafner2023mastering}.
 However, recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}

### 2024-02-02 paper/figs/plotting/reward-ablation/make_plot.py
Get rid of magenta

@@ -24,7 +24,7 @@ COLORS = {
     "iQRL": "#e41a1c",
     "iQRL+rew": "#377eb8",
     "iQRL+cos+rew": "#ff7f00",
-    "TCRL-ours": "magenta",
+    "TCRL-ours": "#4DAF4A",
     #"TCRL-ours-small": "#984ea3",
 }
 # %%

### 2024-02-02 agents/iqrl.py
made agent iQRL

@@ -327,7 +327,7 @@ class MLPReward(MLPResettable):
         return r
 
 
-class VQ_TC_TD3(Agent):
+class iQRL(Agent):
     def __init__(
         self,
         # TD3 config

### 2024-02-02 agents/iqrl.py
updated prints using iQRL

@@ -395,7 +395,7 @@ class iQRL(Agent):
         # encoder_reset_params_freq: int = 10000,  # reset enc params after X param updates
         compile: bool = False,
         device: str = "cuda",
-        name: str = "TC_TD3",
+        name: str = "iQRL",
     ):
         super().__init__(
             observation_space=observation_space, action_space=action_space, name=name
@@ -704,7 +704,7 @@ class iQRL(Agent):
         num_updates = int(num_new_transitions * self.td3.utd_ratio)
         info = {}
 
-        logger.info(f"Performing {num_updates} VQ-TC-TD3 updates...")
+        logger.info(f"Performing {num_updates} iQRL updates...")
         reset_flag = 0
         for i in range(num_updates):
             batch = replay_buffer.sample(self.td3.batch_size, val=False)
@@ -796,7 +796,7 @@ class iQRL(Agent):
 
                 reset_flag = 0
 
-        logger.info("Finished training VQ-TC-TD3")
+        logger.info("Finished training iQRL")
 
         return info
 
@@ -824,7 +824,7 @@ class iQRL(Agent):
         num_updates = int(num_new_transitions * self.td3.utd_ratio)
         info = {}
 
-        logger.info(f"Performing {num_updates} VQ-TC-TD3 updates...")
+        logger.info(f"Performing {num_updates} iQRL updates...")
         reset_flag = 0
         for i in range(num_updates):
             batch = replay_buffer.sample(self.td3.batch_size, val=False)
@@ -960,7 +960,7 @@ class iQRL(Agent):
 
                 reset_flag = 0
 
-        logger.info("Finished training VQ-TC-TD3")
+        logger.info("Finished training iQRL")
 
         return info
 

### 2024-02-02 agents/__init__.py
neatened

@@ -1,12 +1,3 @@
 #!/usr/bin/env python3
-# from .tae_ddpg import TAEDDPG
-# from .td3 import TD3
-# from .critics import Critic, MLPCritic
-# from .actors import Actor, MLPActor
-# from .agents import Agent, ActorCritic, LatentActorCritic, AEDDPG, DDPG, VQDDPG
 from .td3 import TD3
-from .ddpg_ae import DDPG_AE
-from .td_mpc import TD_MPC
-from .tc_td3 import TC_TD3
-from .vq_ddpg import VectorQuantizedDDPG
 from .iqrl import iQRL

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py
changed x-axis scales on figures

@@ -60,7 +60,7 @@ def plot(df, key="episode_reward"):
         hue_order = data.agent.unique()
 
         if idx == 3:
-            sns.lineplot(
+            g = sns.lineplot(
                 # x=int("env_step" / 1000),
                 x="env_step",
                 # x="env_step",
@@ -76,7 +76,7 @@ def plot(df, key="episode_reward"):
             )
             ax.legend().set_title(None)
         else:
-            sns.lineplot(
+            g = sns.lineplot(
                 # x="episode",
                 x="env_step",
                 y=key,
@@ -88,6 +88,14 @@ def plot(df, key="episode_reward"):
                 legend=False,
                 ax=ax,
             )
+        if env == "quadruped-walk":
+            g.set(xlim=(0, 500000))
+        if env == "walker-walk":
+            g.set(xlim=(0, 250000))
+        if env == "dog-walk":
+            g.set(xlim=(0, 1500000))
+        if env == "humanoid-walk":
+            g.set(xlim=(0, 3000000))
 
         ax.set_title(" ".join([ele.capitalize() for ele in env.split("-")]))
         ax.set_xlabel("Environment Steps (1e3)")

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py
moved legend

@@ -59,7 +59,7 @@ def plot(df, key="episode_reward"):
         ax = axs[row, col]
         hue_order = data.agent.unique()
 
-        if idx == 3:
+        if idx == 4:
             g = sns.lineplot(
                 # x=int("env_step" / 1000),
                 x="env_step",

### 2024-02-02 paper/main.tex
updated figs
added codebook stuff

@@
    \FOR{$i$ {\bfseries to} $N_{\text{episodes}}$}
-    \STATE Collect data in environment
+    \STATE Collect data in environment:
     \STATE $\qquad \mathcal{D} \leftarrow \mathcal{D} \cup \{o_{t}, a_{t}, o_{t+1}, r_{t+1}\}^{T}_{t=0}$
     \FOR{$i=1$ {\bfseries to} $T$}
-        \STATE Update representation
+        \STATE Update representation:
         \STATE $\qquad [\theta, \phi] \leftarrow [\theta, \phi] + \alpha \nabla \left( \mathcal{L}_{\text{rep}}(\theta, \phi; \mathcal{D}) \right)$  \COMMENT{\cref{eq:rep-loss}}
@@
         \IF{i \% 2 == 0}
-          \STATE Update actor less frequently than critic
+          \STATE Update actor less frequently than critic:
           \STATE $\quad \quad \eta \leftarrow \psi + \alpha \nabla \left( \mathcal{L}_{\pi}(\psi; \mathcal{D}) \right)$  \COMMENT{\cref{eq:policy-loss}}
@@
         \ENDIF
-        \STATE Update target networks
+        \STATE Update target networks:
         \STATE $\quad \quad [\bar{\theta}, \bar{\phi}, \bar{\psi}] \leftarrow \tau [\bar{\theta}, \bar{\phi}, \bar{\psi}] + (1-\tau) [{\theta}, {\phi}, {\psi}]$
@@
 
-\begin{table}[t]
-\caption{FSQ levels $\mathcal{L}$ to approximate different codebook sizes $|\mathcal{C}|$}
-\label{sample-table}
-\vskip 0.15in
-\begin{center}
-\begin{small}
-\begin{sc}
-\begin{tabular}{lccc}
-\toprule
-Target size $|\mathcal{C}|$ & $2^{8}$ & $2^{10}$ & $2^{12}$ \\
-\midrule
-Proposed $\mathcal{L}$ & $[8,6,5]$ & $[8,5,5,5]$ & $[7,5,5,5,5]$ \\
-% \bottomrule
-\end{tabular}
-\end{sc}
-\end{small}
-\end{center}
-\vskip -0.1in
-\end{table}
-
 \textbf{Method overview}
@@
 &\text{Value: } & q_{t} &= Q_{\psi} (z_{t}, a_{t}) \label{eq:value} \\
-&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) . \label{eq:policy}.
+&\text{Policy: } & a_{t} &\sim \pi_{\eta} (z_{t}) . \label{eq:policy}
 \end{align}
@@
 %
-which minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
-and the next state predicted by the momentum encoder $e_{\bar{\phi}}(z_{t+1})$.
+which minimizes the distance between the next state predicted by the transition model $\hat{z}_{t+1} = d_{\phi}(e_{\phi}(z_{t}), a_{t})$
+and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(z_{t+1})$.
 The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
@@
 of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
+% As such, gradient to not flow through the target network
 The target network update rate is denoted $\tau$.
 
-Unlike previous works, we normalize the latent vector
-
-\textbf{Finite Scalar Quantization}
-We follow Finite Scalar Quantization (FSQ, \cite{mentzerFiniteScalarQuantization2023})
-and bound each dimension of our latent space.
-Given the output of our MLP encoder $z \in \R^{d}$,
-
-where each entry in  $f : z \rightarrow \left[ L/2 \right] \mathrm{tanh}(z))$
-
-\textbf{Bounding the latent space}
-Typically the latent state $z_{t} \in \R^{d}$
-However, in our work we bound the domain of the latent space, \ie $z_{t} \in [-L, L]^{d}$.
+\begin{table}[t]
+\caption{FSQ levels $\mathcal{L}$ to approximate different codebook sizes $|\mathcal{C}|$}
+\label{tab:fsq-levels}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lccc}
+\toprule
+Target size $|\mathcal{C}|$ & $2^{6}$ & $2^{8}$ & $2^{10}$ \\
+\midrule
+Proposed $\mathcal{L}$ & $[8,8]$ & $[8,6,5]$ & $[8,5,5,5]$ \\
+% Target size $|\mathcal{C}|$ & $2^{8}$ & $2^{10}$ & $2^{12}$ \\
+% \midrule
+% Proposed $\mathcal{L}$ & $[8,6,5]$ & $[8,5,5,5]$ & $[7,5,5,5,5]$ \\
+% \bottomrule
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+
+\textbf{Normalization with implicit quantization}
+% However, this can lead to exploding gradients and
+Motivated by preventing representation collapse we seek a representation with high matrix rank.
+To this end, we bound our latent space following the approach from
+Finite Scalar Quantization \citep[FSQ,][]{mentzerFiniteScalarQuantization2023}.
+Note that the latent state $z_{t}$ is typically a vector of unbounded real numbers, \ie $z_{t} \in \R^{d}$.
+% each (scalar) entry in the representation $z_{t}$ is independently quantized to the nearest integer by rounding
+Their important observation is that carefully bounding each channel gives rise to an implicit codebook of a chosen size.
+Let's consider a single element of the latent state $z$ and give it $c$ channels so it's of shape $[c]$.
+% Here $d$ is the output dimension of the NN.
+Mapping each entry $z_{i}$ to $L$ values (\eg, via $z_{i} \rightarrow [L/2] \text{tanh}(z_{i})$ and rounding to integers),
+obtains a quantized $\tilde{z} \in \mathcal{C}$, where $\tilde{z}$ takes one of $L^{c}$ unique possible vectors.
+This gives rise to an implicit codebook $\mathcal{C}$ given by the product of the per-channel codebook sets.
+% The size of the codebook $|\mathcal{C}| = L^{c}$ is given by the product of the
+% For example, specifying $c=3$ channels
+This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values.
+Giving a codebook of size $|\mathcal{C}| = L_{i}^{d_{i}}$.
+% Since quantization is performed by rounding to integers
+
+FSQ has the following hyperparameters. We must specify the number of channels $c$ and the number of levels per channel
+$\mathcal{L} = [L_{1},\ldots,L_{d}]$.
+\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obatin codebooks of
+different sizes.
+FSQ, relies on a simple, fixed grid partition in a much
+lower-dimensional space.
+In practice, we found codebooks of size $|\mathcal{C}|=2^{6}$ sufficient for most environments in the DeepMind Control suite.
+However, for the dog-walk environment we used a codebook of size $|\mathcal{C}|=2^{12}$.
+
+% Fig. 1 shows FSQ for d=3, L=3, implying a codebook
+% C = {(−1, −1, −1),(−1, −1, 0),(−1, −1, 1), . . . ,(1, 1, 1)}, where |C| = L
+% d = 27.
+
+% However, thi w
+% bounding the range of the quantizer
+% However, in our work we bound the domain of the latent space, \ie $z_{t} \in [-L, L]^{d}$.
+% We follow Finite Scalar Quantization (FSQ, \cite{mentzerFiniteScalarQuantization2023})
+% and bound each dimension of our latent space.
+% Given the output of our MLP encoder $z \in \R^{d}$,
+
+% where each entry in  $f : z \rightarrow \left[ L/2 \right] \mathrm{tanh}(z))$
+
+% Unlike previous works, we normalize the latent vector
+
+% \textbf{Bounding the latent space}
 
@@
 
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
+\caption{\textbf{Ablation of normalization}.
+We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:normalization-ablation}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
 \section{Experiments}
@@
 Without normalization, a dimensional collapse occurs, which can have significant harmful effects as the representational power of the latent state diminishes \cite{jingUnderstandingDimensionalCollapse2021}.
-
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
-\caption{\textbf{Ablation of normalization}.
-We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:normalization-ablation}
-\end{center}
-\vskip -0.2in
-\end{figure*}
 
@@
 \label{conclusion}
+We have presented \our, a simple technique for learning task-agnostic representations which can be combined with
+any model-free RL algorithm.
+
+\textbf{Future work}
+To the best of our knowledge, we are the first to present a method that learns a task-agnostic representation
+using only latent-state consistency. As such, we hope our method for learning task-agnostic representations paves the
+way for more research in multi-task and lifelong learning.
+Given that our method obtains an implicit codebook an obvious future direction is exploring if we can leverage the
+discretized latent-space.
+
+\textbf{Limitations}
 

### 2024-02-02 paper/main.tex
moved figures
fix

@@
 
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
+\caption{\textbf{Ablation of normalization}.
+We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:normalization-ablation}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
 \textbf{Normalization with implicit quantization}
@@
 In practice, we found codebooks of size $|\mathcal{C}|=2^{6}$ sufficient for most environments in the DeepMind Control suite.
-However, for the dog-walk environment we used a codebook of size $|\mathcal{C}|=2^{12}$.
+However, for the dog-walk environment we used a codebook of size $|\mathcal{C}|=2^{9}$.
 
@@
 % i=1 Li
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=0.9\textwidth]{./figs/reward-ablation.pdf}}
+\caption{\textbf{Reward prediction is not necessary for representation learning}.
+We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
+\our performs on par with the best method on all environments except Fish Swim, where it is slightly behind our version of TCRL.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:reward-ablation}
+\end{center}
+\vskip -0.2in
+\end{figure*}
+
+\begin{figure*}[ht]
+	\vskip 0.2in
+	\begin{center}
+		\centerline{\includegraphics[width=1.0\textwidth]{./figs/rank-comparison-z1.pdf}}
+		\caption{\textbf{Representation collapse without normalization}.
+We observe that the representation normalization of \our is crucial for avoiding representation collapse.
+When no normalization is applied to the latents, the matrix rank of the representations drops rapidly, which can have a harmful effect on the downstream performance.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+		\label{fig:rank-comparison-z1}
+	\end{center}
+	\vskip -0.2in
+\end{figure*}
 
@@
 
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
-\caption{\textbf{Ablation of normalization}.
-We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:normalization-ablation}
-\end{center}
-\vskip -0.2in
-\end{figure*}
-
 \section{Experiments}
@@
 	\begin{center}
-		\centerline{\includegraphics[width=1.0\textwidth]{./figs/rank-comparison-z1.pdf}}
-		\caption{\textbf{Representation collapse without normalization}.
-We observe that the representation normalization of \our is crucial for avoiding representation collapse.
-When no normalization is applied to the latents, the matrix rank of the representations drops rapidly, which can have a harmful effect on the downstream performance.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-		\label{fig:rank-comparison-z1}
-	\end{center}
-	\vskip -0.2in
-\end{figure*}
-
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/reward-ablation.pdf}}
-\caption{\textbf{Reward prediction is not necessary for representation learning}.
-We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
-\our performs on par with the best method on all environments except Fish Swim, where it is slightly behind our version of TCRL.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:reward-ablation}
-\end{center}
-\vskip -0.2in
-\end{figure*}
-
-\begin{figure*}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=\textwidth]{./figs/task-agnostic-ablation.pdf}}
+		\centerline{\includegraphics[width=0.8\textwidth]{./figs/task-agnostic-ablation.pdf}}
 		\caption{\textbf{\our's task-agnostic representation helps for multi-task RL}.
@@
 
-\section{Full results}
+\section{Further results}
 

### 2024-02-02 paper/main.tex
Sec. 1

@@
 
-Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+Reinforcement learning  \citep[RL, \eg][]{sutton2018reinforcement} has shown much promise in games, animation and robotics.
 However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
-deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+{\em (i)}~handling high dimensional/complex observation spaces is non-trivial, and {\em (ii)}~deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
 
@@
 As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)}~world models that predict the future
 \citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
@@
 However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
+Some approaches used unsupervised learning \citep[\eg VAE,][with reconstruction loss]{kingmaAutoEncoding2014} to acquire representations
 \citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
@@
 In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method
-(we use TD3 \citep{fujimotoAddressingFunctionApproximation2018}).
+\citep[we use TD3][]{fujimotoAddressingFunctionApproximation2018}.
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
@@
 \citep{mentzerFiniteScalarQuantization2023}.
-As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
+As a result, our latent space is bounded and associated with an \emph{implicit} codebook, whose size we can control.
 % As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our method {\em (i)} alleviates representation collapse,
-{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
-{\em (iii)} is simple to implement, and {\em (iv)} learns a task-agnostic
+Importantly, our method {\em (i)}~alleviates representation collapse,
+{\em (ii)}~demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
+{\em (iii)}~is simple to implement, and {\em (iv)}~learns a task-agnostic
 representation which could be helpful in downstream tasks.
@@
 In this section, we detail our method, titled \textit{implicitly Quantized Reinforcement Learning} (\our).
-\our is conceptually simple, it (1) learns a representation of the observation space and then,
-(2) performs model-free RL (e.g. TD3) on this representation.
+\our is conceptually simple, it {\em (i)}~learns a representation of the observation space and then,
+{\em (2)}~performs model-free RL (\eg, TD3) on this representation.
 % It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
@@
 
-We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
+We consider Markov Decision Processes \citep[MDPs,][]{bellmanMarkovianDecisionProcess1957a} $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
 where an agent receives an observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.

### 2024-02-02 paper/main.tex
Sec. 1
Minor

@@
 \end{tikzpicture}
-\caption{\textbf{Overview of \our's representation learning} \our learns a representation solely with the latent-state consistency loss, \ie a self-predictive loss, see \cref{eq:rep-loss}.
-We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it.
-Our normalization results in an implicit codebook (whose size we can control) and helps prevent representation collapse.
+\vskip -0.1in
+\caption{\textbf{Overview of \our's representation learning} \our learns a representation solely with the latent-state consistency loss, \ie a self-predictive loss (see \cref{eq:rep-loss}).
+We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it (denoted by \cube).
+Our normalization results in an implicit codebook (whose size we can control) and helps prevent representation collapse.}
   % Importantly, this prevents representation collapse without needing loss terms to minimize errors in reward and/or value predictions.
   % As such, our representations are task-agnostic.
-  We denote our normalization scheme by \cube.}
+  %We denote our normalization scheme by \cube.}
 \label{fig:diagram-self-predictive}
@@
         \STATE \quad \quad $\psi \leftarrow \psi + \alpha \nabla \left( \mathcal{L}_{Q}(\psi; \mathcal{D}) \right)$ \COMMENT{\cref{eq:value-loss}}
-        \IF{i \% 2 == 0}
+        \IF{$i$ \% 2 == 0}
           \STATE Update actor less frequently than critic:
@@
 % Proposed $\mathcal{L}$ & $[8,6,5]$ & $[8,5,5,5]$ & $[7,5,5,5,5]$ \\
-% \bottomrule
+\bottomrule
 \end{tabular}

### 2024-02-02 paper/main.tex
Minor

@@
   % Styles (this tunes colours etc.)
-  \tikzstyle{mynode}=[fill=white,draw=black!80,rounded corners=1pt,font=\tiny,inner sep=0,align=center]
+  \tikzstyle{mynode}=[fill=black!5,draw=black!20,rounded corners=1pt,font=\tiny,inner sep=0,align=center,line width=1pt]
   \tikzstyle{trap}=[mynode,trapezium,text width=6.5em, minimum height=2em,
@@
   \draw[darr] (zh2) -- (z2);
+  \draw[arr,-] (z2) -- ($(e2.east)!(z2)!(e2.east)$);
 
@@
 \caption{\textbf{Overview of \our's representation learning} \our learns a representation solely with the latent-state consistency loss, \ie a self-predictive loss (see \cref{eq:rep-loss}).
-We follow Finite Scalar Quantization (FSQ) \cite{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it (denoted by \cube).
+We follow Finite Scalar Quantization \citep[FSQ,][]{mentzerFiniteScalarQuantization2023} and bound the domain of our latent representation, \ie we normalize it (denoted by \cube).
 Our normalization results in an implicit codebook (whose size we can control) and helps prevent representation collapse.}
@@
 However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}~handling high dimensional/complex observation spaces is non-trivial, and {\em (ii)}~deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+{\em (i)}~handling high dimensional/complex observation spaces is non-trivial, and {\em (ii)}~deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.\looseness-1
 
@@
 However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning \citep[\eg VAE,][with reconstruction loss]{kingmaAutoEncoding2014} to acquire representations
+Some approaches used unsupervised learning \citep[\eg\ VAE,][with reconstruction loss]{kingmaAutoEncoding2014} to acquire representations
 \citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.

### 2024-02-02 paper/main.tex
Related work: WIP

--- 2024-02-02 paper/main.tex
@@
 
-Reinforcement learning  \citep[RL, \eg][]{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+Reinforcement learning (RL) \citep{sutton2018reinforcement} has shown much promise in games, animation and robotics.
 However, applying RL in real-world environments is challenging. The problem is two-fold,
-{\em (i)}~handling high dimensional/complex observation spaces is non-trivial, and {\em (ii)}~deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
+{\em (i)}, handling high dimensional/complex observation spaces is non-trivial and {\em (ii)}
+deep RL typically requires millions of data points which can be unpractical (or costly), \ie it is very sample inefficient.
 
@@
 As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)}~world models that predict the future
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)} world models that predict the future
 \citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
@@
 However, these approaches are usually limited to simple environments.
-Some approaches used unsupervised learning \citep[\eg VAE,][with reconstruction loss]{kingmaAutoEncoding2014} to acquire representations
+Some approaches used unsupervised learning (\eg VAE \citep{kingmaAutoEncoding2014} with reconstruction loss) to acquire representations
 \citep{finnDeepSpatialAutoencoders2016,higginsDARLAImprovingZeroShot2017,langeAutonomousReinforcementLearning2012,watterEmbedControlLocally2015}.
@@
 In this paper, we propose a simple representation learning technique that can be combined with any model-free RL method
-\citep[we use TD3][]{fujimotoAddressingFunctionApproximation2018}.
+(we use TD3 \citep{fujimotoAddressingFunctionApproximation2018}).
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
@@
 \citep{mentzerFiniteScalarQuantization2023}.
-As a result, our latent space is bounded and associated with an \emph{implicit} codebook, whose size we can control.
+As a result, our latent space is bounded and associated with an \textit{implicit} codebook, whose size we can control.
 % As a result, our latent space is associated with an \textit{implicit} codebook, whose size we can control.
-Importantly, our method {\em (i)}~alleviates representation collapse,
-{\em (ii)}~demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
-{\em (iii)}~is simple to implement, and {\em (iv)}~learns a task-agnostic
+Importantly, our method {\em (i)} alleviates representation collapse,
+{\em (ii)} demonstrates state-of-the-art sample efficiency in continuous control benchmarks,
+{\em (iii)} is simple to implement, and {\em (iv)} learns a task-agnostic
 representation which could be helpful in downstream tasks.
@@
 
+Model-based RL (MBRL) can significantly improve on the sample efficiency of the model-free methods by reasoning about the world and maximizing the fraction of the information stream utilized for learning by utilizing the observations instead of simply trying to learn a policy or a value function to maximize the return \cite{haRecurrentWorldModels2018}.
+The world model can be used for planning \cite{allen1983planning, basye1992decision}.
+A prominent idea has been to to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder (VAE) \cite{kingmaAutoEncoding2014, igl2018deep}.
+These models rely on maximizing the conditional observation likelihood $p(o_t | s_t$), which is the reconstruction objective.
+The latent space of the model can be used for both planning with model-predictive control \cite{hafnerLearning2019} or policy learning in the imagination of the world model \cite{hafner2019dream}.
+\citet{finn2017deep} and \citet{ebert2018visual} show that it is possible to plan with models that have been taught to predict videos directly without a VAE-based latent space.
+Incorporating the model uncertainty into planning through ensembling or sampling can also have significant positive impact on the performance of model-based RL agents \cite{deisenrothPILCO2011, chua2018deep, janner2019trust, yu2020mopo, clavera2020model}.
+
+However, trying to maximize the ELBO or video reconstruction error to learn a world model can mislead the model to focus on learning task-irrelevant information, which distorts the optimization landscape.
+When the task is complex, planning with reconstruction-based world models struggles \cite{lutter2021learning}.
+Instead of relying on a reconstruction-based loss function, \citet{schwarzer2020data} propose a cosine similarity loss between predicted future latents and the true future latents, and performing Q-learning in the learned latent space.
+\citet{ma2021contrastive} propose relying on maximizing the mutual information between latent states and observations to learn a world model by deep reinforcement learning.
+They take the modify the classical world model learning objective by replacing the observation reconstruction loss with a contrastive learning term, which helps combat the model focusing on learning task-irrelevant information.
+\citet{kipf2019contrastive} use contrastive learning to learn a structured world model.
+These contrastive objectives significantly improve sample efficiency, but they are not only limited to model-based reinforcement learning.
+\citet{laskinCURLContrastiveUnsupervised2020} propose replacing the transition function with data augmentations to get a contrastive target, and performing model-free learning using the representation produced by the encoder trained with the contrastive target.
+
 \section{Method}
@@
 In this section, we detail our method, titled \textit{implicitly Quantized Reinforcement Learning} (\our).
-\our is conceptually simple, it {\em (i)}~learns a representation of the observation space and then,
-{\em (2)}~performs model-free RL (\eg, TD3) on this representation.
+\our is conceptually simple, it (1) learns a representation of the observation space and then,
+(2) performs model-free RL (e.g. TD3) on this representation.
 % It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
@@
 
-We consider Markov Decision Processes \citep[MDPs,][]{bellmanMarkovianDecisionProcess1957a} $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
+We consider Markov Decision Processes (MDPs, \citet{bellmanMarkovianDecisionProcess1957a}) $\mathcal{M} = (\mathcal{O}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$,
 where an agent receives an observation $o_{t} \in \mathcal{O}$ at time step $t$ and performs an action $a_{t} \in \mathcal{A}$.
@@
 
-\textbf{Normalizing the latent-space improves sample efficiency}
+\textbf{\our is simple, fast, and performant}
 In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on eight problems from the DeepMind Control Suite.
@@
 
-We analyze the impact of the latent-space normalization on the rank of the latent representations outputted by the learned encoder.
+\textbf{\our does not suffer from rank collapse} We analyze the impact of the latent-space normalization on the rank of the latent representations outputted by the learned encoder.
 We follow the methodology of \citet{ni2024bridging} for estimating the matrix rank.
@@
 
+\begin{figure*}[ht]
+	\vskip 0.2in
+	\begin{center}
+		\centerline{\includegraphics[width=1.0\textwidth]{./figs/reconstruction-loss-ablation.pdf}}
+		\caption{\textbf{Reconstruction loss has detrimental impact}.
+Unlike many methods, such as Dreamer-v3 \cite{hafner2023mastering}, \our neither has an observation decoder nor a reconstruction term in the loss function.
+We show that adding a reconstruction loss harms the performance of \our across all evaluation environments.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+		\label{fig:reconstruction-loss-ablation}
+	\end{center}
+	\vskip -0.2in
+\end{figure*}
+
 \textbf{Reconstruction loss has a detrimental impact}
@@
 \cref{tab:hyperparameters} lists all of the hyperparameters for training \our.
-For further details please check the code which is available at XXXXXXX.
+For further details please check the code which is available in the submitted supplementary material and will be made public upon acceptance.
 
@@
 
-\section{Further results}
-
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/reconstruction-loss-ablation.pdf}}
-\caption{\textbf{Reconstruction loss has detrimental impact}.
-Unlike many methods, such as Dreamer-v3 \cite{hafner2023mastering}, \our neither has an observation decoder nor a reconstruction term in the loss function.
-We show that adding a reconstruction loss harms the performance of \our across all evaluation environments.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:multi-task-pretraining}
-\end{center}
-\vskip -0.2in
-\end{figure*}
-
 % \begin{figure*}[ht]

### 2024-02-02 paper/figs/plotting/reconstruction-ablation/make_plot.py
updates

@@ -33,7 +33,7 @@ main_envs = [
     "hopper-stand",
     "fish-swim",
     "quadruped-run",
-    "humanoid-run",
+    # "humanoid-run",
     # "humanoid-walk",
     "dog-walk",
     # "dog-run",
@@ -61,13 +61,16 @@ def plot(df, key="episode_reward"):
     envs = np.sort(df.env.unique())
     ncol = 4
     # assert envs.shape[0] % ncol == 0
-    nrow = len(main_envs) // ncol
+    # nrow = len(main_envs) // ncol
     # nrow = 1
+    nrow = 2
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
     df = df.rename(columns=rename)
+
+    axs[-1, -1].axis("off")
 
     for idx, env in enumerate(main_envs):
         data = df[df["env"] == env]
@@ -80,7 +83,7 @@ def plot(df, key="episode_reward"):
         # hue_order = data.utd_ratio.unique()
         # breakpoint()
 
-        if idx == 0:
+        if idx == 7:
             sns.lineplot(
                 x="env_step",
                 # x="episode",

### 2024-02-02 paper/main.tex
Updates

@@
 These models rely on maximizing the conditional observation likelihood $p(o_t | s_t$), which is the reconstruction objective.
-The latent space of the model can be used for both planning with model-predictive control \cite{hafnerLearning2019} or policy learning in the imagination of the world model \cite{hafner2019dream}.
+The latent space of the model can be used for both planning with model-predictive control \cite{rubinstein1997optimization, hafnerLearning2019} or policy learning in the imagination of the world model \cite{hafner2019dream}.
 \citet{finn2017deep} and \citet{ebert2018visual} show that it is possible to plan with models that have been taught to predict videos directly without a VAE-based latent space.
@@
 
-However, trying to maximize the ELBO or video reconstruction error to learn a world model can mislead the model to focus on learning task-irrelevant information, which distorts the optimization landscape.
-When the task is complex, planning with reconstruction-based world models struggles \cite{lutter2021learning}.
+Maximize the ELBO or video reconstruction error to learn a world model can mislead the model to focus on learning task-irrelevant information, which distorts the optimization landscape.
+When the task is complex, planning with reconstruction-based world models is unreliable \cite{lutter2021learning} and struggles to outperform model-free methods \cite{kostrikov2020image, yarats2021mastering}.
+Planning can be done through models that learn to predicting rewards and values for multi-step rollouts \cite {tamar2016value, silver2017predictron, oh2017value, schrittwieser2020mastering}.
+However, these methods do not utilize the observations for model learning at all, which can make them sample in-efficient.
+
+How to use observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
+\citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latents.
 Instead of relying on a reconstruction-based loss function, \citet{schwarzer2020data} propose a cosine similarity loss between predicted future latents and the true future latents, and performing Q-learning in the learned latent space.
-\citet{ma2021contrastive} propose relying on maximizing the mutual information between latent states and observations to learn a world model by deep reinforcement learning.
-They take the modify the classical world model learning objective by replacing the observation reconstruction loss with a contrastive learning term, which helps combat the model focusing on learning task-irrelevant information.
+\citet{ma2021contrastive} modify the classical world model learning objective by replacing the observation reconstruction loss with a contrastive learning term, which maximizes the mutual information between latent states and observations and helps combat the model focusing on learning task-irrelevant information.
 \citet{kipf2019contrastive} use contrastive learning to learn a structured world model.
-These contrastive objectives significantly improve sample efficiency, but they are not only limited to model-based reinforcement learning.
+The contrastive objectives for model-based RL significantly improve sample efficiency, but they are not only limited to model-based reinforcement learning.
 \citet{laskinCURLContrastiveUnsupervised2020} propose replacing the transition function with data augmentations to get a contrastive target, and performing model-free learning using the representation produced by the encoder trained with the contrastive target.
+
+TD-MPC \cite{hansenTemporalDifferenceLearning2022} use a consistency loss to learn representations that used for planning with Model Predictive Path Integral together with reward and value functions learned through temporal difference methods \cite{williams2015model}.
+TD-MPC2 \cite{hansen2023td} presents a series of improvements to TD-MPC to make the world model show scaling benefits and perform excellent in a multi-task setting.
+They train world models up to 317 million parameters and show monotone improvement in the normalized score across multiple environments and model sizes.
+The TD-MPC family of methods has been shown to be applicable to real-world robotics \cite{lancaster2023modem} and learn representations that are suitable for transfer learning after adaptation \cite{yang2023movie} or fine-tuning \cite{feng2023finetuning}.
+\citet{zhu2023repo} demonstrate that the observation encoder can be made more resilient to spurious correlations by restricting the information flow from the observations to the latent representation, and the encoder can be easily aligned at test time in a reward-free manner.
+The loss function of TD-MPC can also be used to learn a manager in a hierarchical RL setting \cite{chitnis2023iql}.
+
+\citet{zhaoSimplifiedTemporalConsistency2023} show that the planning component of TD-MPC is not strictly necessary for high performance, and applying model-free learning on top of the self-consistent representations is sufficient for performance competitive with state-of-the-art.
+\citet{yuan2022euclid} propose using intrinsic rewards for pre-training temporally consistent models, before using the rewards for adapting the world model for downstream tasks.
+We build on top of TCRL to show that we can combat the representation collapse through a normalization term and learn task-agnostic representations by dropping the reward term of the TCRL loss and only utilizing a multistep temporal consistency loss.
 
@@
     \item How does \our compare to state-of-the-art model-free and model-based algorithms?
+    \item Does the FSQ-based normalization of \our help combat rank collapse of the learned latent-space representation?
     \item Is learning a representation with only latent-state consistency really better than including reward predictions and value predictions?
@@
 
-\textbf{Future work}
+\textbf{Limitations and Future work}
 To the best of our knowledge, we are the first to present a method that learns a task-agnostic representation
@@
 discretized latent-space.
-
-\textbf{Limitations}
+So far, we have only applied our representation learning loss to state-based tasks.
+Extending the method to handle even higher-dimensional visual tasks is an interesting direction.
+Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
+Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
+Exploring whether explicitly quantifying the model uncertainty by e.g.
+using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
 

### 2024-02-02 paper/main.tex
App
Minor

@@
 Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
-Exploring whether explicitly quantifying the model uncertainty by e.g.
-using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
+Exploring whether explicitly quantifying the model uncertainty by \eg\ using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
 
@@
 
-\begin{table}[t]
+\begin{table}[h]
 \caption{Hyperparameters for \our.}
@@
 Latent dimension ($d$) & $1024$ & \\
-% \bottomrule
+ \bottomrule
 \end{tabular}

### 2024-02-02 paper/main.tex
Fig. 1
Minor updates to appendix
Code mentioned

@@
   % Draw nodes
-  \node[blob] (z0) at (0\nodedist,0) {$\vz_{\sub}$};
+  \node[blob] (z0) at (0\nodedist,0) {$\vz_{\sub}\vphantom{\hat\vz_{\subs{1}}}$};
   \node[blob] (z1) at (1\nodedist,0) {$\hat\vz_{\subs{1}}$};
@@
 
-\textbf{Limitations and Future work}
+\textbf{Limitations and future work}
 To the best of our knowledge, we are the first to present a method that learns a task-agnostic representation
@@
 
+% Arno recommends to have this here as a last thing (leaves a positive feeling)
+A reference implementation of the methods and benchmarking setup will be made available on GitHub upon acceptance of the paper (see supplementary material for now).
+
 \section*{Broader Impact}
@@
 
-\section{Hyperparameters}
+\section{Hyperparameters, Implementation, and Training Infrastructure}
 \cref{tab:hyperparameters} lists all of the hyperparameters for training \our.
 For further details please check the code which is available in the submitted supplementary material and will be made public upon acceptance.
+We implemented our codebase with PyTorch and used the Adam optimizer for training the models.
+We used five seeds for all figures.
+All of our models can be trained on a single GPU.
+We used Nvidia A100s and AMD Instinct MI250X GPUs to run our experiments.
 

### 2024-02-02 paper/main.tex
Code mentioned

@@
 % Arno recommends to have this here as a last thing (leaves a positive feeling)
-A reference implementation of the methods and benchmarking setup will be made available on GitHub upon acceptance of the paper (see supplementary material for now).
+A reference implementation of the methods will be made available on GitHub upon acceptance of the paper (see supplementary material for now).
 

### 2024-02-02 paper/main.tex
minor

@@
 The world model can be used for planning \cite{allen1983planning, basye1992decision}.
-A prominent idea has been to to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder (VAE) \cite{kingmaAutoEncoding2014, igl2018deep}.
+A prominent idea has been to to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder  \citep[VAE,][]{kingmaAutoEncoding2014, igl2018deep}.
 These models rely on maximizing the conditional observation likelihood $p(o_t | s_t$), which is the reconstruction objective.

### 2024-02-02 paper/main.tex
remove bullet
Related work

--- 2024-02-02 paper/main.tex
@@
 
-Maximize the ELBO or video reconstruction error to learn a world model can mislead the model to focus on learning task-irrelevant information, which distorts the optimization landscape.
+Maximizing the ELBO or video reconstruction error to learn a world model can mislead the model to focus on learning task-irrelevant information, which distorts the optimization landscape \cite{zhang2018natural,zintgraf2021varibad}.
 When the task is complex, planning with reconstruction-based world models is unreliable \cite{lutter2021learning} and struggles to outperform model-free methods \cite{kostrikov2020image, yarats2021mastering}.
+The problems are aggravated when the environment is partially observable \cite{morad2023popgym}.
 Planning can be done through models that learn to predicting rewards and values for multi-step rollouts \cite {tamar2016value, silver2017predictron, oh2017value, schrittwieser2020mastering}.
@@
 
-How to use observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
+How to use high-dimensional and noisy observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
+A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning}.
+A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate, ni2024bridging}.
 \citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latents.
@@
 \citet{kipf2019contrastive} use contrastive learning to learn a structured world model.
+\citet{ghugare2022simplifying} argue that aligning the auxiliary objectives and RL objective can be difficult, and derive a bound directly on the overall RL objective, while is compatible with self-consistency.
 The contrastive objectives for model-based RL significantly improve sample efficiency, but they are not only limited to model-based reinforcement learning.
@@
 Without normalization, a dimensional collapse occurs, which can have significant harmful effects as the representational power of the latent state diminishes \cite{jingUnderstandingDimensionalCollapse2021}.
+Note that in contrast to \citet{ni2024bridging}, we observe that using a detached momentum encoder as the target encoder is insufficient to prevent a rank collapse in practice.
+We hypothesize that this is due to trivial representations being enough to minimize the prediction error, when a reward term is not present in the loss \cite{tang2023understanding}.
 
@@
 		\label{fig:multi-task-pretraining}
+	\end{center}
+	\vskip -0.2in
+\end{figure*}
+
+\begin{figure*}[ht]
+	\vskip 0.2in
+	\begin{center}
+		\centerline{\includegraphics[width=1.0\textwidth]{./figs/reconstruction-loss-ablation.pdf}}
+		\caption{\textbf{Reconstruction loss has detrimental impact}.
+Unlike many methods, such as Dreamer-v3 \cite{hafner2023mastering}, \our neither has an observation decoder nor a reconstruction term in the loss function.
+We show that adding a reconstruction loss harms the performance of \our across all evaluation environments.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+		\label{fig:reconstruction-loss-ablation}
 	\end{center}
@@
 
-\begin{figure*}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=1.0\textwidth]{./figs/reconstruction-loss-ablation.pdf}}
-		\caption{\textbf{Reconstruction loss has detrimental impact}.
-Unlike many methods, such as Dreamer-v3 \cite{hafner2023mastering}, \our neither has an observation decoder nor a reconstruction term in the loss function.
-We show that adding a reconstruction loss harms the performance of \our across all evaluation environments.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-		\label{fig:reconstruction-loss-ablation}
-	\end{center}
-	\vskip -0.2in
-\end{figure*}
-
 \textbf{Reconstruction loss has a detrimental impact}
@@
 
-\textbf{Limitations and future work}
+\textbf{Limitations and Future work}
 To the best of our knowledge, we are the first to present a method that learns a task-agnostic representation
@@
 Exploring whether explicitly quantifying the model uncertainty by \eg\ using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
-
-% Arno recommends to have this here as a last thing (leaves a positive feeling)
-A reference implementation of the methods will be made available on GitHub upon acceptance of the paper (see supplementary material for now).
 

### 2024-02-02 paper/main.tex

@@
   % Minimization
-  \node[blue!50,below of=z1,align=center,font=\scriptsize,xshift=4em] (min) {minimize: $\|\bar\vz_{\bullet}-\hat\vz_{\bullet}\|^{2}_{2}$};
+  \node[blue!50,below of=z1,align=center,font=\scriptsize,xshift=4em] (min) {minimize: $\|\bar\vz-\hat\vz\|^{2}_{2}$};
   \draw[blue!50,line width=1pt,shorten <=1mm,shorten >=1mm,->] (min) to[bend right=30] ($(z1)!.5!(zh1)$);
@@
 Exploring whether explicitly quantifying the model uncertainty by \eg\ using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
+
+% Arno recommends to have this here as a last thing (leaves a positive feeling)
+A reference implementation of the methods will be made available on GitHub upon acceptance of the paper (see supplementary material for now).
 

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py
updated figures

@@ -35,7 +35,7 @@ main_envs = [
     "walker-walk",
     "humanoid-walk",
     "dog-walk",
-    "dog-run",
+    # "dog-run",
 ]
 
 
@@ -43,14 +43,17 @@ def plot(df, key="episode_reward"):
     envs = np.sort(df.env.unique())
     ncol = 4
     # assert envs.shape[0] % ncol == 0
-    nrow = len(main_envs) // ncol
+    # nrow = len(main_envs) // ncol
+    nrow = 2
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
+    # df = df[df["env_step"] < 1000000]
 
-    # breakpoint()
+    all_data = []
     # df["env_step"] = df["env_step"] / 1000
     for idx, env in enumerate(main_envs):
+        idx += 1
         data = df[df["env"] == env]
         # breakpoint()
         # data[data["agent"] == "iFSQ-RL"] = data[data["agent"] == "iFSQ-RL"].iloc[::2]
@@ -58,6 +61,16 @@ def plot(df, key="episode_reward"):
         col = idx % ncol
         ax = axs[row, col]
         hue_order = data.agent.unique()
+
+        min_ep_length = 1e10
+        for agent in data.agent:
+            ep_length = np.max(df[df["agent"] == agent]["env_step"])
+            # print(f"ep_length {ep_length} for env {env}")
+            min_ep_length = min(min_ep_length, ep_length)
+            # min_ep_length=np.min(min_ep)
+        # print(f"min_ep_length {min_ep_length} for env {env}")
+        all_data += [df[df["env_step"] < min_ep_length]]
+        # breakpoint()
 
         if idx == 4:
             g = sns.lineplot(
@@ -93,13 +106,34 @@ def plot(df, key="episode_reward"):
         if env == "walker-walk":
             g.set(xlim=(0, 250000))
         if env == "dog-walk":
-            g.set(xlim=(0, 1500000))
+            g.set(xlim=(0, 750000))
         if env == "humanoid-walk":
             g.set(xlim=(0, 3000000))
 
         ax.set_title(" ".join([ele.capitalize() for ele in env.split("-")]))
         ax.set_xlabel("Environment Steps (1e3)")
         ax.set_ylabel("Episode Return")
+
+    df = pd.concat(all_data)
[…]

### 2024-02-02 paper/figs/plotting/reconstruction-ablation/fetch_data.py

@@ -69,6 +69,20 @@ ddpg_data = {
             "ucvgujrp",
             "zqv19l3p",
         ],
+        "cartpole-swingup-d=64-rec": [
+            "ibu5qn0t",
+            "bb079aef",
+            "ilhv49du",
+            "e15gra4x",
+            "u1rmgrtb",
+        ],
+        "cartpole-swingup-d=64": [
+            "dul9ksc7",
+            "aygtwe76",
+            "chmq024k",
+            "mwhja16u",
+            "lmkngxtl",
+        ],
         # "hopper-stand": [
         #     ##### Project=True
         #     # # reconstruction

### 2024-02-02 paper/main.tex
small changes

--- 2024-02-02 paper/main.tex
@@
   % Draw nodes
-  \node[blob] (z0) at (0\nodedist,0) {$\vz_{\sub}\vphantom{\hat\vz_{\subs{1}}}$};
+  \node[blob] (z0) at (0\nodedist,0) {$\vz_{\sub}$};
   \node[blob] (z1) at (1\nodedist,0) {$\hat\vz_{\subs{1}}$};
@@
 The world model can be used for planning \cite{allen1983planning, basye1992decision}.
-A prominent idea has been to to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder  \citep[VAE,][]{kingmaAutoEncoding2014, igl2018deep}.
+A prominent idea has been to to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder (VAE) \cite{kingmaAutoEncoding2014, igl2018deep}.
 These models rely on maximizing the conditional observation likelihood $p(o_t | s_t$), which is the reconstruction objective.
@@
 \centerline{\includegraphics[width=1.0\textwidth]{./figs/baselines_comparison.pdf}}
-\caption{\textbf{Normalizing the latent-space improves sample efficiency} \our is significanly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
+\caption{\textbf{\our is simple, fast, and performant} \our is significantly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
 It performs similarly to TD-MPC (orange) whilst being significantly simpler and quicker to train.
@@
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+
 \label{fig:normalization_improves_sample_efficiency}
@@
 \label{sec:experiments}
-In this section, we evaluate \our in a wide variety of continuous control tasks from the DeepMind Control (DMC) suite \cite{tassa2018deepmind}.
+In this section, we evaluate \our in a variety of continuous control tasks from the DeepMind Control (DMC) suite \cite{tassa2018deepmind}.
 We aim to answer the following questions:
@@
 \textbf{\our is simple, fast, and performant}
-In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on eight problems from the DeepMind Control Suite.
+In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based baselines TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on eight problems from the DeepMind Control Suite.
 The results show that our method, \our, performs on par with TD-MPC while being significantly simpler and quicker to train, as the loss term is simpler and no planning is needed unlike in TD-MPC.
@@
 any model-free RL algorithm.
-
-\textbf{Limitations and future work}
+In our experiments we compared against both state-of-the-art model-free RL methods and state-of--the-art model-based RL methods.
+
+Our results demonstrate state-of-the-art
+based on our empirical results, we suggest further
+
+\textbf{Limitations and Future work}
 To the best of our knowledge, we are the first to present a method that learns a task-agnostic representation
 using only latent-state consistency. As such, we hope our method for learning task-agnostic representations paves the
-way for more research in multi-task and lifelong learning.
+way for more research into multi-task and lifelong learning.
 Given that our method obtains an implicit codebook an obvious future direction is exploring if we can leverage the
@@
 Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
-Exploring whether explicitly quantifying the model uncertainty by \eg\ using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
-
-% Arno recommends to have this here as a last thing (leaves a positive feeling)
-A reference implementation of the methods will be made available on GitHub upon acceptance of the paper (see supplementary material for now).
+Exploring whether explicitly quantifying the model uncertainty by e.g.
+using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
 
@@
 
-\section{Hyperparameters, Implementation, and Training Infrastructure}
+\section{Hyperparameters}
 \cref{tab:hyperparameters} lists all of the hyperparameters for training \our.
 For further details please check the code which is available in the submitted supplementary material and will be made public upon acceptance.
-We implemented our codebase with PyTorch and used the Adam optimizer for training the models.
-We used five seeds for all figures.
-All of our models can be trained on a single GPU.
-We used Nvidia A100s and AMD Instinct MI250X GPUs to run our experiments.
-
-\begin{table}[h]
+
+\begin{table}[t]
 \caption{Hyperparameters for \our.}
@@
 Latent dimension ($d$) & $1024$ & \\
- \bottomrule
+% \bottomrule
 \end{tabular}

### 2024-02-02 paper/main.tex

--- 2024-02-02 paper/main.tex
@@
 any model-free RL algorithm.
+In our experiments we compared against both state-of-the-art model-free RL methods and state-of--the-art model-based RL methods.
+
+Our results demonstrate state-of-the-art
+based on our empirical results, we suggest further
 
@@
 using only latent-state consistency. As such, we hope our method for learning task-agnostic representations paves the
-way for more research in multi-task and lifelong learning.
+way for more research into multi-task and lifelong learning.
 Given that our method obtains an implicit codebook an obvious future direction is exploring if we can leverage the

### 2024-02-02 paper/main.tex
Fixed typos

--- 2024-02-02 paper/main.tex
@@
 \icmlauthor{Aidan Scannell}{aalto,fcai}
+\icmlauthor{Kalle Kujanpää}{aalto,fcai}
+\icmlauthor{Yi Zhao}{aalto}
 \icmlauthor{Arno Solin}{aalto}
@@
 
-Reinforcement learning  \citep[RL, \eg][]{sutton2018reinforcement} has shown much promise in games, animation and robotics.
+Reinforcement learning  \citep[RL, \eg][]{sutton2018reinforcement} has shown much promise in games, animation, and robotics.
 However, applying RL in real-world environments is challenging. The problem is two-fold,
@@
 As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxillary tasks on the agent's observations and {\em (ii)}~world models that predict the future
+{\em (i)} auxiliary tasks on the agent's observations and {\em (ii)}~world models that predict the future
 \citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
-In particular, we focus on representation learning methods which compress observations into latent states.
+In this work, we are interested in the first class of methods, which use auxiliary tasks to improve sample efficiency.
+In particular, we focus on representation learning methods that compress observations into latent states.
 Nevertheless, our representation learning borrows ideas from world models as it uses multi-step predictions in a
-latent-space dynamics model to learn temporally consist representations \citep{zhaoSimplifiedTemporalConsistency2023}.
+latent-space dynamics model to learn temporally consistent representations \citep{zhaoSimplifiedTemporalConsistency2023}.
 % In particular, methods for compressing observations into latent states.%, aka representation learning.

### 2024-02-02 paper/main.tex
Minor mods

@@
 How to use high-dimensional and noisy observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
-A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning}.
-A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate, ni2024bridging}.
+A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning.
+ni2024bridging}.
+A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate}.
 \citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latents.
@@
 \textbf{\our's task-agnostic representation is helpful for multi-task RL}
-In contrast to TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (\cref{eq:rep-loss}).
+In contrast to \eg TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (\cref{eq:rep-loss}).
 The potential upside of \our's task-agnostic representations is that they can be used to speed-up learning in the incremental multi-task setting.

### 2024-02-02 paper/main.tex
Minor

--- 2024-02-02 paper/main.tex
@@
 Nevertheless, our representation learning borrows ideas from world models as it uses multi-step predictions in a
-latent-space dynamics model to learn temporally consistent representations \citep{zhaoSimplifiedTemporalConsistency2023}.
+latent-space dynamics model to learn temporally consistent representations \citep{zhaoSimplifiedTemporalConsistency2023}.\looseness-1
 % In particular, methods for compressing observations into latent states.%, aka representation learning.

### 2024-02-02 paper/main.tex
Language fixes

--- 2024-02-02 paper/main.tex
@@
 This requires an encoder $e_{\theta}$ which maps observations $o_{t}$ to latent states $z_{t}$ and a transition model $d_{\phi}$ which predicts
-next latent states $z_{t+1}$ given an a latent state $z_{t}$ and an action $a_{t}$.
+the next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
 It then minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
@@
 See \cref{fig:diagram-self-predictive} for an overview of our (extremely simple) representation learning method.
-It is based soley on the latent-state consistency loss, \ie a commonly used self-supervised loss for continuous RL.
+It is based solely on the latent-state consistency loss, \ie a commonly used self-supervised loss for continuous RL.
 In contrast to prior works, we normalize our latent representation using the technique from Finite Scalar Quantization
@@
 {\em (iii)}~is simple to implement, and {\em (iv)}~learns a task-agnostic
-representation which could be helpful in downstream tasks.
+representation that could be helpful in downstream tasks.
 
@@
 
-Model-based RL (MBRL) can significantly improve on the sample efficiency of the model-free methods by reasoning about the world and maximizing the fraction of the information stream utilized for learning by utilizing the observations instead of simply trying to learn a policy or a value function to maximize the return \cite{haRecurrentWorldModels2018}.
+Model-based RL (MBRL) can significantly improve the sample efficiency of the model-free methods by reasoning about the world and maximizing the fraction of the information stream utilized for learning by utilizing the observations instead of simply trying to learn a policy or a value function to maximize the return \cite{haRecurrentWorldModels2018}.
 The world model can be used for planning \cite{allen1983planning, basye1992decision}.
-A prominent idea has been to to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder  \citep[VAE,][]{kingmaAutoEncoding2014, igl2018deep}.
+A prominent idea has been to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder  \citep[VAE,][]{kingmaAutoEncoding2014, igl2018deep}.
 These models rely on maximizing the conditional observation likelihood $p(o_t | s_t$), which is the reconstruction objective.
@@
 \citet{finn2017deep} and \citet{ebert2018visual} show that it is possible to plan with models that have been taught to predict videos directly without a VAE-based latent space.
-Incorporating the model uncertainty into planning through ensembling or sampling can also have significant positive impact on the performance of model-based RL agents \cite{deisenrothPILCO2011, chua2018deep, janner2019trust, yu2020mopo, clavera2020model}.
+Incorporating the model uncertainty into planning through ensembling or sampling can also have a significant positive impact on the performance of model-based RL agents \cite{deisenrothPILCO2011, chua2018deep, janner2019trust, yu2020mopo, clavera2020model}.
 
@@
 The problems are aggravated when the environment is partially observable \cite{morad2023popgym}.
-Planning can be done through models that learn to predicting rewards and values for multi-step rollouts \cite {tamar2016value, silver2017predictron, oh2017value, schrittwieser2020mastering}.
-However, these methods do not utilize the observations for model learning at all, which can make them sample in-efficient.
+Planning can be done through models that learn to predict rewards and values for multi-step rollouts \cite {tamar2016value, silver2017predictron, oh2017value, schrittwieser2020mastering}.
+However, these methods do not utilize the observations for model learning at all, which can make them inefficient.
 
@@
 A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate}.
-\citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latents.
-Instead of relying on a reconstruction-based loss function, \citet{schwarzer2020data} propose a cosine similarity loss between predicted future latents and the true future latents, and performing Q-learning in the learned latent space.
+\citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latent representations.
+Instead of relying on a reconstruction-based loss function, \citet{schwarzer2020data} propose a cosine similarity loss between predicted future latents and the true future latents and Q-learning in the learned latent space.
 \citet{ma2021contrastive} modify the classical world model learning objective by replacing the observation reconstruction loss with a contrastive learning term, which maximizes the mutual information between latent states and observations and helps combat the model focusing on learning task-irrelevant information.
 \citet{kipf2019contrastive} use contrastive learning to learn a structured world model.
-\citet{ghugare2022simplifying} argue that aligning the auxiliary objectives and RL objective can be difficult, and derive a bound directly on the overall RL objective, while is compatible with self-consistency.
+\citet{ghugare2022simplifying} argue that aligning the auxiliary objectives and RL objective can be difficult.
+The authors derive a bound directly on the overall RL objective, which is compatible with self-consistency.
 The contrastive objectives for model-based RL significantly improve sample efficiency, but they are not only limited to model-based reinforcement learning.
-\citet{laskinCURLContrastiveUnsupervised2020} propose replacing the transition function with data augmentations to get a contrastive target, and performing model-free learning using the representation produced by the encoder trained with the contrastive target.
-
-TD-MPC \cite{hansenTemporalDifferenceLearning2022} use a consistency loss to learn representations that used for planning with Model Predictive Path Integral together with reward and value functions learned through temporal difference methods \cite{williams2015model}.
-TD-MPC2 \cite{hansen2023td} presents a series of improvements to TD-MPC to make the world model show scaling benefits and perform excellent in a multi-task setting.
-They train world models up to 317 million parameters and show monotone improvement in the normalized score across multiple environments and model sizes.
-The TD-MPC family of methods has been shown to be applicable to real-world robotics \cite{lancaster2023modem} and learn representations that are suitable for transfer learning after adaptation \cite{yang2023movie} or fine-tuning \cite{feng2023finetuning}.
+\citet{laskinCURLContrastiveUnsupervised2020} propose replacing the transition function with data augmentations to get a contrastive target and performing model-free learning using the representation produced by the encoder trained with the contrastive target.
+
+TD-MPC \cite{hansenTemporalDifferenceLearning2022} use a consistency loss to learn representations for planning with Model Predictive Path Integral together with reward and value functions learned through temporal difference methods \cite{williams2015model}.
+TD-MPC2 \cite{hansen2023td} presents a series of improvements to TD-MPC to make the world model show scaling benefits and achieve state-of-the-art performance in a multi-task setting.
+They train world models with up to 317 million parameters and show monotone improvement in the normalized score across multiple environments and model sizes.
+The TD-MPC family of methods is applicable to real-world robotics \cite{lancaster2023modem} and learn representations that are suitable for transfer learning after adaptation \cite{yang2023movie} or fine-tuning \cite{feng2023finetuning}.
 \citet{zhu2023repo} demonstrate that the observation encoder can be made more resilient to spurious correlations by restricting the information flow from the observations to the latent representation, and the encoder can be easily aligned at test time in a reward-free manner.
@@
 \citet{zhaoSimplifiedTemporalConsistency2023} show that the planning component of TD-MPC is not strictly necessary for high performance, and applying model-free learning on top of the self-consistent representations is sufficient for performance competitive with state-of-the-art.
-\citet{yuan2022euclid} propose using intrinsic rewards for pre-training temporally consistent models, before using the rewards for adapting the world model for downstream tasks.
+\citet{yuan2022euclid} propose using intrinsic rewards for pre-training temporally consistent models before using the rewards for adapting the world model for downstream tasks.
 We build on top of TCRL to show that we can combat the representation collapse through a normalization term and learn task-agnostic representations by dropping the reward term of the TCRL loss and only utilizing a multistep temporal consistency loss.

### 2024-02-02 paper/main.tex
update figs and corrected some notation

--- 2024-02-02 paper/main.tex
@@
 \icmlauthor{Aidan Scannell}{aalto,fcai}
-\icmlauthor{Kalle Kujanpää}{aalto,fcai}
-\icmlauthor{Yi Zhao}{aalto}
 \icmlauthor{Arno Solin}{aalto}
@@
 
-Reinforcement learning  \citep[RL, \eg][]{sutton2018reinforcement} has shown much promise in games, animation, and robotics.
+Reinforcement learning  \citep[RL, \eg][]{sutton2018reinforcement} has shown much promise in games, animation and robotics.
 However, applying RL in real-world environments is challenging. The problem is two-fold,
@@
 As \citet{laskinCURLContrastiveUnsupervised2020} highlighted, these can generally be divided into two streams of research:
-{\em (i)} auxiliary tasks on the agent's observations and {\em (ii)}~world models that predict the future
+{\em (i)} auxillary tasks on the agent's observations and {\em (ii)}~world models that predict the future
 \citep{haRecurrentWorldModels2018,hafnerLearning2019,hansenTemporalDifferenceLearning2022}.
-In this work, we are interested in the first class of methods, which use auxiliary tasks to improve sample efficiency.
-In particular, we focus on representation learning methods that compress observations into latent states.
+In this work, we are interested in the first class of methods, which use auxillary tasks to improve sample efficiency.
+In particular, we focus on representation learning methods which compress observations into latent states.
 Nevertheless, our representation learning borrows ideas from world models as it uses multi-step predictions in a
-latent-space dynamics model to learn temporally consistent representations \citep{zhaoSimplifiedTemporalConsistency2023}.\looseness-1
+latent-space dynamics model to learn temporally consist representations \citep{zhaoSimplifiedTemporalConsistency2023}.
 % In particular, methods for compressing observations into latent states.%, aka representation learning.
@@
 How to use high-dimensional and noisy observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
-A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning.
-ni2024bridging}.
-A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate}.
+A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning}.
+A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate, ni2024bridging}.
 \citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latents.
@@
 % \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
-&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
+&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(\hat{z}_{t}, a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
 \end{align}
 %
-which minimizes the distance between the next state predicted by the transition model $\hat{z}_{t+1} = d_{\phi}(e_{\phi}(z_{t}), a_{t})$
-and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(z_{t+1})$.
-The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
+which minimizes the distance between the next state predicted by the transition model $\hat{z}_{t+1} = d_{\phi}(\hat{z}_{t}, a_{t})$
+and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(s_{t+1})$.
+The latent states are obtained with multi-step predictions in the latent space $\hat{z}_{t+1} = d(\hat{z}_{t}, a_{t})$
+with $\hat{z}_{0} = e_{\theta}(s_{0})$.
+The initial mapping to the latent space $\hat{z}_{0} = e_{\theta}(s_{0})$ uses the online encoder which
+is being trained jointly with the transition model $d_{\phi}(\hat{z_{t}}, a_{t})$.
 The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
@@
 \cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obatin codebooks of
-different sizes.
+different sizes \citep{mentzerFiniteScalarQuantization2023}.
 FSQ, relies on a simple, fixed grid partition in a much
@@
 The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
-encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
+target encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
 The value function is updated by minimizing the following objective:
@@
 	\end{align*}
-	where $\tilde{r}_{t+h}$ is the ground-truth H-step reward and $\hat{r}_{t+h}$ the predicted H-step rewards.
+	where $r_{t+h}$ is the ground-truth H-step reward and $\hat{r}_{t+h}$ the predicted H-step rewards.
 	\item \ourcosrew, where in addition to the reward prediction head in \ourrew, we additionally convert the self-supervised MSE loss (see \cref{eq:rep-loss}) for the temporal consistency into a cosine similarity, similarly to TCRL:
 	\begin{align*} \label{eq:rep-loss}
-		\hat{z}_{t+h} &= d_{\phi}(e_{\theta}(s_{t}), a_{t}) \\
+		\hat{z}_{t+h} &= d_{\phi}(\hat{z}_{t}, a_{t}) \\
 		\tilde{z}_{t+h} &= e_{\bar{\theta}}(s_{t+1}) \\
@@
 Our results imply that learning to predict the reward is not necessary for learning a suitable latent representation.
-The upside of not including the reward prediction head is that it makes learning the representation learning task-agnostic, which we believe to be   important for downstream applications such as learning in an incremental multi-task setting.
+The upside of not including the reward prediction head is that it makes the representation task-agnostic, which we believe to be   important for downstream applications such as learning in an incremental multi-task setting.
 
 \textbf{\our's task-agnostic representation is helpful for multi-task RL}
-In contrast to \eg TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (\cref{eq:rep-loss}).
+In contrast to TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (\cref{eq:rep-loss}).
 The potential upside of \our's task-agnostic representations is that they can be used to speed-up learning in the incremental multi-task setting.

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py

@@ -130,7 +130,7 @@ def plot(df, key="episode_reward"):
         legend="auto",
         ax=axs[0, 0],
     )
-    axs[0, 0].set_title("Avg. Over 8 DMC Tasks")
+    axs[0, 0].set_title("Avg. Over 7 DMC Tasks")
     axs[0, 0].set_xlabel("Environment Steps (1e3)")
     axs[0, 0].set_ylabel("Episode Return")
 

### 2024-02-02 paper/main.tex

--- 2024-02-02 paper/main.tex
@@
 Our results imply that learning to predict the reward is not necessary for learning a suitable latent representation.
-The upside of not including the reward prediction head is that it makes learning the representation learning task-agnostic, which we believe to be   important for downstream applications such as learning in an incremental multi-task setting.
+The upside of not including the reward prediction head is that it makes the representation task-agnostic, which we believe to be   important for downstream applications such as learning in an incremental multi-task setting.
 

### 2024-02-02 paper/main.tex
Fixing typos

--- 2024-02-02 paper/main.tex
@@
 which can aid RL.
-The latent-space dynamics model $d_{\phi}$ predicts next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
-It's sole purpose is to aid representation learning by making the latent states temporally consistent.
+The latent-space dynamics model $d_{\phi}$ predicts the next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
+Its sole purpose is to aid representation learning by making the latent states temporally consistent.
 Note that we do not use it for model-based RL.
@@
 actor $\pi_{\eta}$.
-Following prior works \citep{yaratsMasteringVisualContinuous2021} we use a linear exploration noise schedule
+Following prior works \citep{yaratsMasteringVisualContinuous2021}, we use a linear exploration noise schedule
 which decays from $1$ to $0.1$ during training.
@@
 \our even outperforms TCRL, which is the most similar baseline.
-Results are for eight DMC tasks with UTD=1.
+Results are for seven DMC tasks with UTD=1.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(z_{t+1})$.
-The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
-The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
+The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder, which is being trained jointly with the transition model.
+The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder, which uses an exponential moving average (EMA)
 of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
@@
 % For example, specifying $c=3$ channels
-This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values.
-Giving a codebook of size $|\mathcal{C}| = L_{i}^{d_{i}}$.
+This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values, giving a codebook of size $|\mathcal{C}| = L_{i}^{d_{i}}$.
 % Since quantization is performed by rounding to integers
 
-FSQ has the following hyperparameters. We must specify the number of channels $c$ and the number of levels per channel
+FSQ has the following hyperparameters: we must specify the number of channels $c$ and the number of levels per channel
 $\mathcal{L} = [L_{1},\ldots,L_{d}]$.
-\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obatin codebooks of
+\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obtain codebooks of
 different sizes.
@@
 In practice, we found codebooks of size $|\mathcal{C}|=2^{6}$ sufficient for most environments in the DeepMind Control suite.
-However, for the dog-walk environment we used a codebook of size $|\mathcal{C}|=2^{9}$.
+However, for the Dog Walk environment we used a codebook of size $|\mathcal{C}|=2^{9}$.
 
@@
 We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
-\our performs on par with the best method on all environments except Fish Swim, where it is slightly behind our version of TCRL.
+\our performs on par with the best method on all environments except Fish Swim, where it is slightly inferior to our version of TCRL.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 We observe that the representation normalization of \our is crucial for avoiding representation collapse.
-When no normalization is applied to the latents, the matrix rank of the representations drops rapidly, which can have a harmful effect on the downstream performance.
+When no normalization is applied to the latent variables, the matrix rank of the representations drops rapidly, which can hurt the downstream performance.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 We then use the target value functions $Q_{\bar{\psi}_{k}}$ and the target policy $\pi_{\bar{\eta}}$ to calculate the TD target.
-Following TD3, we learn the actor's parameters by minimizing,
+Following TD3, we learn the actor's parameters by minimizing
 %
@@
 %
-That is, maximizing the Q-value using the clipped double Q learning trick to combat overestimation in Q learning.
+That is, we maximize the Q-value using the clipped double Q learning trick to combat overestimation in Q learning.
 
@@
 \label{sec:experiments}
-In this section, we evaluate \our in a variety of continuous control tasks from the DeepMind Control (DMC) suite \cite{tassa2018deepmind}.
+In this section, we evaluate \our in a variety of continuous control tasks from the DeepMind Control (DMC) Suite \cite{tassa2018deepmind}.
 We aim to answer the following questions:
@@
     \item Is learning a representation with only latent-state consistency really better than including reward predictions and value predictions?
-    \item Can \our's task-agnostic representation aid task-adaptation with its pre-trained representation?
+    \item Can \our's task-agnostic representation aid task adaptation with its pre-trained representation?
 \end{itemize}
@@
 \textbf{\our is simple, fast, and performant}
-In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based baselines TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on eight problems from the DeepMind Control Suite.
-The results show that our method, \our, performs on par with TD-MPC while being significantly simpler and quicker to train, as the loss term is simpler and no planning is needed unlike in TD-MPC.
+In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based baselines TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on seven problems from the DeepMind Control Suite.
+The results show that our method, \our, performs on par with TD-MPC while being significantly simpler and quicker to train, as the loss term consists of fewer terms and no planning is needed, unlike in TD-MPC.
 We even outperform TCRL, which is the most similar baseline to our work.
@@
 We plot the ranks in \cref{fig:rank-comparison-z1}.
-The latent state dimensions are 256, 128, 256, and 128 for Acrobot Swingup, Hopper Stand, Humanoid Run and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
+The latent state dimensions are 256, 128, 256, and 128 for Acrobot Swingup, Hopper Stand, Humanoid Run, and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
 Without normalization, a dimensional collapse occurs, which can have significant harmful effects as the representational power of the latent state diminishes \cite{jingUnderstandingDimensionalCollapse2021}.
 Note that in contrast to \citet{ni2024bridging}, we observe that using a detached momentum encoder as the target encoder is insufficient to prevent a rank collapse in practice.
-We hypothesize that this is due to trivial representations being enough to minimize the prediction error, when a reward term is not present in the loss \cite{tang2023understanding}.
+We hypothesize that this is due to trivial representations being enough to minimize the prediction error when a reward term is not present in the loss \cite{tang2023understanding}.
 
@@
 The pre-training proves to be beneficial on all evaluated environments.
-If the representation is task-specific (the reward prediction term is included), the pre-training can be clearly detrimental.
+If the representation is task-specific, \ie the reward prediction term is included, the pre-training can be clearly detrimental.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (\cref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
-Instead, we rely solely on the temporal consistency loss.
+Instead, we rely solely on the self-predictive temporal consistency loss.
 To analyze the impact of not including the reward prediction term, we perform an ablation study, where we compare our method to three variants of our method:
@@
 The results for this ablation study are shown in \cref{fig:reward-ablation}. 
-The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods with the exception of Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
+The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization due to not suffering from latent collapse in this specific environment.
 Even then, \our performs equally to the normalized variants of our method with the reward prediction.

### 2024-02-02 paper/main.tex
small updates

--- 2024-02-02 paper/main.tex
@@
 How to use high-dimensional and noisy observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
-A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning.
-ni2024bridging}.
+A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning, ni2024bridging}.
 A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate}.
@@
 \textbf{\our's task-agnostic representation is helpful for multi-task RL}
-In contrast to \eg TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss formula (\cref{eq:rep-loss}).
-The potential upside of \our's task-agnostic representations is that they can be used to speed-up learning in the incremental multi-task setting.
+In contrast to \eg TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss (\cref{eq:rep-loss}).
+The potential upside of \our's task-agnostic representation is it can be used to speed-up learning in the incremental multi-task setting.
 That is, we can leverage the representation learned from solving one task to aid learning another tasks in the same domain.
@@
 
-The plots in \cref{fig:multi-task-pretraining} show that when the encoder and the dynamics function have been pre-trained, the agent learns quicker than when it is traind from scratch.
+The plots in \cref{fig:multi-task-pretraining} show that when the encoder and the dynamics function have been pre-trained, the agent learns quicker than when it is trained from scratch.
 This effect is particularly visible in Quadruped Run and Humanoid Run.
 In contrast, when the reward term has been included in the pre-training objective, making the representation task-specific, the pre-training proves to be harmful on Quadruped Run.
-This results strongly indicate that the task-agnostic representations are highly beneficial in the continual and lifelong learning settings.
+These results indicate that \our's task-agnostic representation is beneficial in the continual and lifelong learning settings.
+% However, given the breadth of
 

### 2024-02-02 paper/main.tex

--- 2024-02-02 paper/main.tex
@@
 % \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
-&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
+&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(\hat{z}_{t}, a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
 \end{align}
 %
-which minimizes the distance between the next state predicted by the transition model $\hat{z}_{t+1} = d_{\phi}(e_{\phi}(z_{t}), a_{t})$
-and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(z_{t+1})$.
-The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder, which is being trained jointly with the transition model.
-The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder, which uses an exponential moving average (EMA)
+which minimizes the distance between the next state predicted by the transition model $\hat{z}_{t+1} = d_{\phi}(\hat{z}_{t}, a_{t})$
+and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(s_{t+1})$.
+The latent states are obtained with multi-step predictions in the latent space $\hat{z}_{t+1} = d(\hat{z}_{t}, a_{t})$
+with $\hat{z}_{0} = e_{\theta}(s_{0})$.
+The initial mapping to the latent space $\hat{z}_{0} = e_{\theta}(s_{0})$ uses the online encoder which
+is being trained jointly with the transition model $d_{\phi}(\hat{z_{t}}, a_{t})$.
+The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
 of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
@@
 $\mathcal{L} = [L_{1},\ldots,L_{d}]$.
-\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obtain codebooks of
-different sizes.
+\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obtain codebooks of different sizes \citep{mentzerFiniteScalarQuantization2023}.
 FSQ, relies on a simple, fixed grid partition in a much
@@
 The results for this ablation study are shown in \cref{fig:reward-ablation}. 
-The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization due to not suffering from latent collapse in this specific environment.
+The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except in Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization due to not suffering from rank collapse in this specific environment.
 Even then, \our performs equally to the normalized variants of our method with the reward prediction.
 Our results imply that learning to predict the reward is not necessary for learning a suitable latent representation.
-The upside of not including the reward prediction head is that it makes learning the representation learning task-agnostic, which we believe to be important for downstream applications such as learning in an incremental multi-task setting.
+The upside of not including the reward prediction head is that it makes the representation task-agnostic, which we believe to be important for downstream applications such as learning in an incremental multi-task setting.
 

### 2024-02-02 paper/main.tex
Deleted a lie

@@
 The results for this ablation study are shown in \cref{fig:reward-ablation}. 
-The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except in Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization due to not suffering from rank collapse in this specific environment.
+The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except in Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
 Even then, \our performs equally to the normalized variants of our method with the reward prediction.

### 2024-02-02 paper/main.tex
Language fixes and updated rank comparison plot

@@
 We plot the ranks in \cref{fig:rank-comparison-z1}.
-The latent state dimensions are 256, 128, 256, and 128 for Acrobot Swingup, Hopper Stand, Humanoid Run, and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
+The latent state dimensions are 256, 256, 256, and 128 for Acrobot Swingup, Fish Swim, Humanoid Run, and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
 Without normalization, a dimensional collapse occurs, which can have significant harmful effects as the representational power of the latent state diminishes \cite{jingUnderstandingDimensionalCollapse2021}.
@@
 The results for this ablation study are shown in \cref{fig:reward-ablation}. 
-The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except in Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
+The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except in Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization due to being able to maintain the latent matrix rank in this specific environment (see \cref{fig:rank-comparison-z1}).
 Even then, \our performs equally to the normalized variants of our method with the reward prediction.
@@
 In contrast to \eg TCRL \cite{zhaoSimplifiedTemporalConsistency2023} and TD-MPC \cite{hansenTemporalDifferenceLearning2022}, \our learns a representation that is task-agnostic, that is, there are no task-specific components such as reward or value prediction in the representation loss (\cref{eq:rep-loss}).
-The potential upside of \our's task-agnostic representation is it can be used to speed-up learning in the incremental multi-task setting.
-That is, we can leverage the representation learned from solving one task to aid learning another tasks in the same domain.
-For instance, we can pre-train a representation in Walker Walk, and use it to speed up the training process in Walker Run.
+The potential upside of \our's task-agnostic representation is it can be used to speed up learning in the incremental multi-task setting.
+That is, we can leverage the representation learned from solving one task to aid in learning other tasks in the same domain.
+For instance, we can pre-train a representation in Walker Walk and use it to speed up the training process in Walker Run.
 The only requirement is that the action and observation spaces must match.
@@
 
-We evaluate whether pre-training representations is beneficial on three environments.
-We start by pre-training \our on Walker Walk, Quadruped Walk and Humanoid Walk.
+We evaluate whether pre-trained representations are beneficial in three environments.
+We start by pre-training \our on Walker Walk, Quadruped Walk, and Humanoid Walk.
 In addition, we replicate the pre-training on Walker Walk and Quadruped Walk by adding a reward prediction term to the representation loss.
-Then, we keep the encoder and the dynamics function from the pre-training, and perform training on the target tasks, which are Walker Run, Quadruped Run, and Humanoid-Run.
+Then, we retain the encoder and dynamics function from the pre-training and train \our on the target tasks Walker Run, Quadruped Run, and Humanoid Run.
 We compare the pre-trained agents to training \our on these tasks from scratch.
-The results for this experiment can be found in \cref{fig:multi-task-pretraining}.
+The results for this experiment are shown in \cref{fig:multi-task-pretraining}.
 
@@
 In contrast, when the reward term has been included in the pre-training objective, making the representation task-specific, the pre-training proves to be harmful on Quadruped Run.
-These results indicate that \our's task-agnostic representation is beneficial in the continual and lifelong learning settings.
+These results indicate that \our's task-agnostic representation is beneficial in continual and lifelong learning settings.
 % However, given the breadth of
@@
 \textbf{Reconstruction loss has a detrimental impact}
-Learning to minimize the observation reconstruction error has been widely applied in model-based reinforcement learning \cite{sutton2018reinforcement, haRecurrentWorldModels2018, hafnerLearning2019}, and an observation decoder has been a component of the most successful RL algorithms to date \cite{hafner2023mastering}.
+Learning to minimize the observation reconstruction error has been widely applied in model-based reinforcement learning \cite{sutton2018reinforcement, haRecurrentWorldModels2018, hafnerLearning2019}, and an observation decoder has been a component of many of the most successful RL algorithms to date \cite{hafner2023mastering}.
 However, recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
-has shown that incorporating a reconstruction term into the representation loss can have a negative impact on performance.
-They hypothesize that learning to reconstruct the observations is inefficient as they can have irrelevant details and visuals like shading that are uncontrollable by the agent and do not affect the tasks.
+has shown that incorporating a reconstruction term into the representation loss can have a negative impact on performance, as learning to reconstruct the observations is inefficient due to the observations containing irrelevant details and visuals like shading that are uncontrollable by the agent and do not affect the tasks.
 
@@
 We perform reconstruction at each time step in the horizon.
-The results in \cref{fig:multi-task-pretraining} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
+The results in \cref{fig:multi-task-pretraining} show that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
 %Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on
@@
 any model-free RL algorithm.
-In our experiments we compared against both state-of-the-art model-free RL methods and state-of--the-art model-based RL methods.
+In our experiments, we compared \our to both state-of-the-art model-free RL methods and state-of-the-art model-based RL methods.
 
@@
 way for more research into multi-task and lifelong learning.
-Given that our method obtains an implicit codebook an obvious future direction is exploring if we can leverage the
+Given that our method obtains an implicit codebook, an obvious future direction is exploring whether we can leverage the
 discretized latent-space.
-So far, we have only applied our representation learning loss to state-based tasks.
-Extending the method to handle even higher-dimensional visual tasks is an interesting direction.
+So far, we have only applied our representation loss to state-based tasks.
+Extending the method to handle even higher-dimensional visual tasks is an attractive direction.
 Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
-Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
-Exploring whether explicitly quantifying the model uncertainty by \eg\ using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
+Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample-efficient.
+Exploring whether explicitly quantifying the model uncertainty by \eg using larger ensembles would help to adapt our algorithm to a real-world robotics setting is also a promising direction for future work.
 
@@
 Therefore, it is possible that our methods could be employed in unforeseen harmful ways, such as autonomous drones for military or law enforcement use.
-There are even other potential societal consequences of our work, none which we feel must be specifically highlighted here.
+There are even other potential societal consequences of our work, none of which we feel must be specifically highlighted here.
 
@@
 \cref{tab:hyperparameters} lists all of the hyperparameters for training \our.
-For further details please check the code which is available in the submitted supplementary material and will be made public upon acceptance.
+For further details, please check the code, which is available in the submitted supplementary material and will be made public upon acceptance.
 We implemented our codebase with PyTorch and used the Adam optimizer for training the models.
 We used five seeds for all figures.
-All of our models can be trained on a single GPU.
+All of our experiments have been run on a single GPU.
 We used Nvidia A100s and AMD Instinct MI250X GPUs to run our experiments.

### 2024-02-02 paper/figs/plotting/rank-comparison/fetch_data.py

@@ -69,6 +69,14 @@ ddpg_data = {
             "4xfe6bez",
             "95l445ci",
         ],
+        "fish-swim-no-norm": [
+            "a2r96nmy",
+            "1cezxqig",
+            "15r32rsc",
+            "rxyyhje0",
+            "8pho3asn",
+        ],
+        "fish-swim": ["3duujm1p", "kl6xd92o", "s4t5viko", "flcmnkt7", "1zjl534e"],
         # 'humanoid-walk': ['2xa4zh7i', 'u01r5cd8', '2dm45ezi', '2ywqbphi', '33kix5gc', ],
         # "dog-walk": ["gtrhfxn3"],
         # "dog-run": ["ukzsy2t5"],

### 2024-02-02 paper/main.tex
updated conclusion plus small edits

--- 2024-02-02 paper/main.tex
@@
 which can aid RL.
-The latent-space dynamics model $d_{\phi}$ predicts the next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
-Its sole purpose is to aid representation learning by making the latent states temporally consistent.
+The latent-space dynamics model $d_{\phi}$ predicts next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
+It's sole purpose is to aid representation learning by making the latent states temporally consistent.
 Note that we do not use it for model-based RL.
@@
 actor $\pi_{\eta}$.
-Following prior works \citep{yaratsMasteringVisualContinuous2021}, we use a linear exploration noise schedule
+Following prior works \citep{yaratsMasteringVisualContinuous2021} we use a linear exploration noise schedule
 which decays from $1$ to $0.1$ during training.
@@
 \our even outperforms TCRL, which is the most similar baseline.
-Results are for seven DMC tasks with UTD=1.
+Results are for eight DMC tasks with UTD=1.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 % For example, specifying $c=3$ channels
-This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values, giving a codebook of size $|\mathcal{C}| = L_{i}^{d_{i}}$.
+This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values.
+Giving a codebook of size $|\mathcal{C}| = L_{i}^{d_{i}}$.
 % Since quantization is performed by rounding to integers
 
-FSQ has the following hyperparameters: we must specify the number of channels $c$ and the number of levels per channel
+FSQ has the following hyperparameters. We must specify the number of channels $c$ and the number of levels per channel
 $\mathcal{L} = [L_{1},\ldots,L_{d}]$.
-\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obtain codebooks of different sizes \citep{mentzerFiniteScalarQuantization2023}.
+\cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obatin codebooks of
+different sizes \citep{mentzerFiniteScalarQuantization2023}.
 FSQ, relies on a simple, fixed grid partition in a much
@@
 In practice, we found codebooks of size $|\mathcal{C}|=2^{6}$ sufficient for most environments in the DeepMind Control suite.
-However, for the Dog Walk environment we used a codebook of size $|\mathcal{C}|=2^{9}$.
+However, for the dog-walk environment we used a codebook of size $|\mathcal{C}|=2^{9}$.
 
@@
 We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
-\our performs on par with the best method on all environments except Fish Swim, where it is slightly inferior to our version of TCRL.
+\our performs on par with the best method on all environments except Fish Swim, where it is slightly behind our version of TCRL.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 We observe that the representation normalization of \our is crucial for avoiding representation collapse.
-When no normalization is applied to the latent variables, the matrix rank of the representations drops rapidly, which can hurt the downstream performance.
+When no normalization is applied to the latents, the matrix rank of the representations drops rapidly, which can have a harmful effect on the downstream performance.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 We then use the target value functions $Q_{\bar{\psi}_{k}}$ and the target policy $\pi_{\bar{\eta}}$ to calculate the TD target.
-Following TD3, we learn the actor's parameters by minimizing
+Following TD3, we learn the actor's parameters by minimizing,
 %
@@
 %
-That is, we maximize the Q-value using the clipped double Q learning trick to combat overestimation in Q learning.
+That is, maximizing the Q-value using the clipped double Q learning trick to combat overestimation in Q learning.
 
@@
 \label{sec:experiments}
-In this section, we evaluate \our in a variety of continuous control tasks from the DeepMind Control (DMC) Suite \cite{tassa2018deepmind}.
+In this section, we evaluate \our in a variety of continuous control tasks from the DeepMind Control (DMC) suite \cite{tassa2018deepmind}.
 We aim to answer the following questions:
@@
     \item Is learning a representation with only latent-state consistency really better than including reward predictions and value predictions?
-    \item Can \our's task-agnostic representation aid task adaptation with its pre-trained representation?
+    \item Can \our's task-agnostic representation aid task-adaptation with its pre-trained representation?
 \end{itemize}
@@
 \textbf{\our is simple, fast, and performant}
-In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based baselines TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on seven problems from the DeepMind Control Suite.
-The results show that our method, \our, performs on par with TD-MPC while being significantly simpler and quicker to train, as the loss term consists of fewer terms and no planning is needed, unlike in TD-MPC.
+In \cref{fig:normalization_improves_sample_efficiency}, we compare \our to the model-free baselines Soft-Actor Critic (SAC) \cite{haarnojaSoft2018} and Randomized Ensemble Double Q Learning (REDQ) \cite{chenRandomizedEnsembledDouble2021}, and the model-based baselines TD-MPC \cite{hansenTemporalDifferenceLearning2022} and TCRL \cite{zhaoSimplifiedTemporalConsistency2023} on eight problems from the DeepMind Control Suite.
+The results show that our method, \our, performs on par with TD-MPC while being significantly simpler and quicker to train, as the loss term is simpler and no planning is needed unlike in TD-MPC.
 We even outperform TCRL, which is the most similar baseline to our work.
@@
 We plot the ranks in \cref{fig:rank-comparison-z1}.
-The latent state dimensions are 256, 128, 256, and 128 for Acrobot Swingup, Hopper Stand, Humanoid Run, and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
+The latent state dimensions are 256, 128, 256, and 128 for Acrobot Swingup, Hopper Stand, Humanoid Run and Cheetah Run, respectively, and the plots show that the matrix rank stays close to the latent state dimension when our FSQ normalization scheme is applied.
 Without normalization, a dimensional collapse occurs, which can have significant harmful effects as the representational power of the latent state diminishes \cite{jingUnderstandingDimensionalCollapse2021}.
 Note that in contrast to \citet{ni2024bridging}, we observe that using a detached momentum encoder as the target encoder is insufficient to prevent a rank collapse in practice.
-We hypothesize that this is due to trivial representations being enough to minimize the prediction error when a reward term is not present in the loss \cite{tang2023understanding}.
+We hypothesize that this is due to trivial representations being enough to minimize the prediction error, when a reward term is not present in the loss \cite{tang2023understanding}.
 
@@
 The pre-training proves to be beneficial on all evaluated environments.
-If the representation is task-specific, \ie the reward prediction term is included, the pre-training can be clearly detrimental.
+If the representation is task-specific (the reward prediction term is included), the pre-training can be clearly detrimental.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
@@
 Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (\cref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
-Instead, we rely solely on the self-predictive temporal consistency loss.
+Instead, we rely solely on the temporal consistency loss.
 To analyze the impact of not including the reward prediction term, we perform an ablation study, where we compare our method to three variants of our method:
@@
 	\begin{align*} \label{eq:rep-loss}
-		\hat{z}_{t+h} &= d_{\phi}(\hat{z}_{t}, a_{t}) \\
-		\tilde{z}_{t+h} &= e_{\bar{\theta}}(s_{t+1}) \\
 		\mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
@@
 	\end{align*}
+  where as before we have $\hat{z}_{t+h} = d_{\phi}(\hat{z}_{t}, a_{t})$ and $\tilde{z}_{t+h} = e_{\bar{\theta}}(s_{t+1})$.
 	\item TCRL-ours, where in addition to \ourcosrew, we remove the FSQ normalization scheme to make the method equal to TCRL with some minor improvements to the method, such as using nstep-TD3 \cite{fujimotoAddressingFunctionApproximation2018} as the base algorithm instead of DDPG with Double Q and nstep returns, using the neural network architecture from \citet{hansen2023td}, and performing policy smoothing \cite{fujimotoAddressingFunctionApproximation2018}.
@@
 The results for this ablation study are shown in \cref{fig:reward-ablation}. 
-The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods except in Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
+The plots show that our method, \our, without a reward prediction term in the loss, has equal or superior performance to the comparison methods with the exception of Fish Swim, where the unnormalized TCRL-ours outperforms all the methods with normalization.
 Even then, \our performs equally to the normalized variants of our method with the reward prediction.
@@
 We perform reconstruction at each time step in the horizon.
-The results in \cref{fig:multi-task-pretraining} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
+The results in \cref{fig:reconstruction-loss-ablation} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
 %Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on
@@
 \label{conclusion}
-We have presented \our, a simple technique for learning task-agnostic representations which can be combined with
-any model-free RL algorithm.
-In our experiments we compared against both state-of-the-art model-free RL methods and state-of--the-art model-based RL methods.
-
-Our results demonstrate state-of-the-art
-based on our empirical results, we suggest further
+We have presented \our, a simple technique for learning representations which demonstrates state-of-the-art performance in
+continuous control tasks.
+Importantly, our normalization and bounding of the latent space empirically preserves the representation's matrix rank,
+indicating that it alleviates representation collapse.
+Importantly, our method is {\em (i)} straightforward, {\em (ii)} compatible with any model-free RL algorithm,
+and {\em (iii)} learns a task-agnostic representation.
+% Moreover, it can be combined with any model-free RL algorithm
+% In our experiments we compared against both state-of-the-art model-free RL methods and state-of--the-art model-based RL methods.
+% Our results demonstrate state-of-the-art based on our empirical results, we suggest further
 
@@
 Extending the method to handle even higher-dimensional visual tasks is an interesting direction.
-Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
+% Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
 Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.

### 2024-02-02 paper/main.tex

--- 2024-02-02 paper/main.tex
@@
 We perform reconstruction at each time step in the horizon.
-The results in \cref{fig:multi-task-pretraining} show that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
+The results in \cref{fig:reconstruction-loss-ablation} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
 %Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on
@@
 \label{conclusion}
-We have presented \our, a simple technique for learning task-agnostic representations which can be combined with
-any model-free RL algorithm.
-In our experiments, we compared \our to both state-of-the-art model-free RL methods and state-of-the-art model-based RL methods.
-
-Our results demonstrate state-of-the-art
-based on our empirical results, we suggest further
+We have presented \our, a simple technique for learning representations which demonstrates state-of-the-art performance in
+continuous control tasks.
+Importantly, our normalization and bounding of the latent space empirically preserves the representation's matrix rank,
+indicating that it alleviates representation collapse.
+Importantly, our method is {\em (i)} straightforward, {\em (ii)} compatible with any model-free RL algorithm,
+and {\em (iii)} learns a task-agnostic representation.
+% Moreover, it can be combined with any model-free RL algorithm
+% In our experiments we compared against both state-of-the-art model-free RL methods and state-of--the-art model-based RL methods.
+% Our results demonstrate state-of-the-art based on our empirical results, we suggest further
 
@@
 way for more research into multi-task and lifelong learning.
+% Given that our method obtains an implicit codebook an obvious future direction is exploring if we can leverage the
+% discretized latent-space.
+So far, we have only applied our representation learning loss to state-based tasks.
+Extending the method to handle even higher-dimensional visual tasks is an interesting direction.
+% Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
+Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
+Exploring whether explicitly quantifying the model uncertainty by \eg\ using larger ensembles would help to adapt our method to a real-world robotics setting is also a promising direction for future work.
 Given that our method obtains an implicit codebook, an obvious future direction is exploring whether we can leverage the
-discretized latent-space.
-So far, we have only applied our representation loss to state-based tasks.
-Extending the method to handle even higher-dimensional visual tasks is an attractive direction.
-Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
+discretized latent-space. So far, we have only applied our representation loss to state-based tasks. Extending the method to handle even higher-dimensional visual tasks is an attractive direction.
+% Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
 Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample-efficient.

### 2024-02-02 paper/main.tex
Minor typo fix
Minor updates to appendix

@@
 Given that our method obtains an implicit codebook, an obvious future direction is exploring whether we can leverage the
-discretized latent-space. So far, we have only applied our representation loss to state-based tasks. Extending the method to handle even higher-dimensional visual tasks is an attractive direction.
+discretized latent space. So far, we have only applied our representation loss to state-based tasks. Extending the method to handle even higher-dimensional visual tasks is an attractive direction.
 % Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
@@
 \section{Hyperparameters, Implementation, and Training Infrastructure}
-\cref{tab:hyperparameters} lists all of the hyperparameters for training \our.
-For further details, please check the code, which is available in the submitted supplementary material and will be made public upon acceptance.
-We implemented our codebase with PyTorch and used the Adam optimizer for training the models.
-We used five seeds for all figures.
-All of our experiments have been run on a single GPU.
+\cref{tab:hyperparameters} lists all of the hyperparameters for training \our and the ablations of our method.
+For further details of the implementation, model architectures, and training please check the code, which is available in the submitted supplementary material and will be made public upon acceptance to guarantee seamless reproducibility.
+We implemented our codebase with PyTorch and used the AdamW optimizer for training the models.
+We used five seeds for all figures and plotted the 95 \% confidence intervals as the shaded area.
 We used Nvidia A100s and AMD Instinct MI250X GPUs to run our experiments.
+All our experiments have been run on a single GPU with a single-digit number of CPU workers.
 

### 2024-02-02 paper/main.tex
re ordered figures

--- 2024-02-02 paper/main.tex
@@
 \citet{finn2017deep} and \citet{ebert2018visual} show that it is possible to plan with models that have been taught to predict videos directly without a VAE-based latent space.
-Incorporating the model uncertainty into planning through ensembling or sampling can also have a significant positive impact on the performance of model-based RL agents \cite{deisenrothPILCO2011, chua2018deep, janner2019trust, yu2020mopo, clavera2020model}.
+% Incorporating the model uncertainty into planning through ensembling or sampling can also have a significant positive impact on the performance of model-based RL agents \cite{deisenrothPILCO2011, chua2018deep, janner2019trust, yu2020mopo, clavera2020model}.
 
@@
 Planning can be done through models that learn to predict rewards and values for multi-step rollouts \cite {tamar2016value, silver2017predictron, oh2017value, schrittwieser2020mastering}.
-However, these methods do not utilize the observations for model learning at all, which can make them inefficient.
+However, these methods do not utilize the observations for model learning, making them inefficient.
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/baselines_comparison.pdf}}
+\caption{\textbf{\our is simple, fast, and performant} \our is significantly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
+It performs similarly to TD-MPC (orange) whilst being significantly simpler and quicker to train.
+\our even outperforms TCRL, which is the most similar baseline.
+Results are for seven DMC tasks with UTD=1.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+
+\label{fig:normalization_improves_sample_efficiency}
+\end{center}
+\vskip -0.2in
+\end{figure*}
 
@@
 
-TD-MPC \cite{hansenTemporalDifferenceLearning2022} use a consistency loss to learn representations for planning with Model Predictive Path Integral together with reward and value functions learned through temporal difference methods \cite{williams2015model}.
+TD-MPC \cite{hansenTemporalDifferenceLearning2022} use a consistency loss to learn representations for planning with Model Predictive Path Integral control together with reward and value functions learned through temporal difference methods \cite{williams2015model}.
 TD-MPC2 \cite{hansen2023td} presents a series of improvements to TD-MPC to make the world model show scaling benefits and achieve state-of-the-art performance in a multi-task setting.
@@
 See \cref{fig:diagram-self-predictive} and \cref{alg:main_alg}.
+
+\begin{figure*}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
+\caption{\textbf{Ablation of normalization}.
+We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
+We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
+\label{fig:normalization-ablation}
+\end{center}
+\vskip -0.2in
+\end{figure*}
 
@@
 
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/baselines_comparison.pdf}}
-\caption{\textbf{\our is simple, fast, and performant} \our is significantly more sample efficient than the model-free baselines SAC (blue) and REDQ (purple).
-It performs similarly to TD-MPC (orange) whilst being significantly simpler and quicker to train.
-\our even outperforms TCRL, which is the most similar baseline.
-Results are for seven DMC tasks with UTD=1.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-
-\label{fig:normalization_improves_sample_efficiency}
-\end{center}
-\vskip -0.2in
-\end{figure*}
-
 \textbf{Representation learning}
@@
 \begin{center}
-\centerline{\includegraphics[width=1.0\textwidth]{./figs/normalization-ablation.pdf}}
-\caption{\textbf{Ablation of normalization}.
-We analyze the impact of our FSQ-based normalization scheme on the sample-efficiency of \our.
+\centerline{\includegraphics[width=0.9\textwidth]{./figs/reward-ablation.pdf}}
+\caption{\textbf{Reward prediction is not necessary for representation learning}.
+We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
+\our performs on par with the best method on all environments except Fish Swim, where it is slightly inferior to our version of TCRL.
 We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:normalization-ablation}
+\label{fig:reward-ablation}
 \end{center}
@@
 % i=1 Li
-
-\begin{figure*}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=0.9\textwidth]{./figs/reward-ablation.pdf}}
-\caption{\textbf{Reward prediction is not necessary for representation learning}.
-We compare \our to variants of our method with a reward prediction term included in the representation learning loss.
-\our performs on par with the best method on all environments except Fish Swim, where it is slightly inferior to our version of TCRL.
-We plot the mean (solid line) and the $95\%$ confidence intervals (shaded) across 5 random seeds, where each seed averages over 10 evaluation episodes.}
-\label{fig:reward-ablation}
-\end{center}
-\vskip -0.2in
-\end{figure*}
 
@@
 In contrast, when the reward term has been included in the pre-training objective, making the representation task-specific, the pre-training proves to be harmful on Quadruped Run.
-These results indicate that \our's task-agnostic representation is beneficial in continual and lifelong learning settings.
+This indicates that \our's task-agnostic representation is beneficial in continual and lifelong learning settings.
 % However, given the breadth of
@@
 We perform reconstruction at each time step in the horizon.
-The results in \cref{fig:reconstruction-loss-ablation} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the overall performance, thus lending further support to the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
+The results in \cref{fig:reconstruction-loss-ablation} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the performance.
+This supports the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.
 %Moreover, in the quadruped-run and dog-walk environments incorporating a reconstruction term has a detrimental impact on

### 2024-02-02 paper/main.tex
updates

@@
 
-\textbf{Reward predict is not necessary for representation learning}
+\textbf{Reward prediction is not necessary for representation learning}
 Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (\cref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py

@@ -72,7 +72,7 @@ def plot(df, key="episode_reward"):
         all_data += [df[df["env_step"] < min_ep_length]]
         # breakpoint()
 
-        if idx == 4:
+        if idx == 2:
             g = sns.lineplot(
                 # x=int("env_step" / 1000),
                 x="env_step",
@@ -127,7 +127,7 @@ def plot(df, key="episode_reward"):
         hue="agent",
         # hue_order=hue_order,
         palette=COLORS,
-        legend="auto",
+        legend=False,
         ax=axs[0, 0],
     )
     axs[0, 0].set_title("Avg. Over 7 DMC Tasks")

### 2024-02-02 paper/main.tex
typos

@@
 
-How to use high-dimensional and noisy observations efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
+How can high-dimensional and noisy observations be used efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
 A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning, ni2024bridging}.
@@
 We perform reconstruction at each time step in the horizon.
-The results in \cref{fig:reconstruction-loss-ablation} shows that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the performance.
+The results in \cref{fig:reconstruction-loss-ablation} show that in no environments does reconstruction aid learning, and in some environments, such as the difficult Dog Walk and Quadruped Run, including the reconstruction term has a significant detrimental effect on the performance.
 This supports the observations of \citet{zhaoSimplifiedTemporalConsistency2023} and \citet{hansenTemporalDifferenceLearning2022} about the lack of need for a reconstruction target in continuous control tasks.

### 2024-02-02 paper/main.tex
reinforcement learning -> RL

@@
 The authors derive a bound directly on the overall RL objective, which is compatible with self-consistency.
-The contrastive objectives for model-based RL significantly improve sample efficiency, but they are not only limited to model-based reinforcement learning.
+The contrastive objectives for model-based RL significantly improve sample efficiency, but they are not only limited to model-based RL.
 \citet{laskinCURLContrastiveUnsupervised2020} propose replacing the transition function with data augmentations to get a contrastive target and performing model-free learning using the representation produced by the encoder trained with the contrastive target.
@@
 \textbf{Reward prediction is not necessary for representation learning}
-Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learning loss (\cref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
+Unlike prior methods such as TCLR \cite{zhaoSimplifiedTemporalConsistency2023}, TD-MPC \cite{hansenTemporalDifferenceLearning2022}, and Dreamer-V2 \cite{hafnerMasteringAtariDiscrete2022}, our representation learningloss (\cref{eq:rep-loss}) does not include a term for learning to predict the reward or the value.
 Instead, we rely solely on the self-predictive temporal consistency loss.
@@
 		\mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
-		&= - \sum_{h=0}^{H-1} \gamma^{h} \left ( \frac{\hat{z}_{t+h}}{\| \hat{z}_{t+h} \|_2} \right )^T \left ( \frac{\tilde{z}_{t+h}}{\| \tilde{z}_{t+h} \|_2} \right )
+		&= - \sum_{h=0}^{H-1} \gamma^{h} \left ( \frac{\hat{z}_{t+h}}{\| \hat{z}_{t+h} \|_2} \right )^T \left ( \frac{\tilde{z}_{t+h}}{\| \tilde{z}_{t+h} \|_2} \right ).
 	\end{align*}
-  where as before we have $\hat{z}_{t+h} = d_{\phi}(\hat{z}_{t}, a_{t})$ and $\tilde{z}_{t+h} = e_{\bar{\theta}}(s_{t+1})$.
+  % where as before we have $\hat{z}_{t+h} = d_{\phi}(\hat{z}_{t}, a_{t})$ and $\tilde{z}_{t+h} = e_{\bar{\theta}}(s_{t+1})$.
 	\item TCRL-ours, where in addition to \ourcosrew, we remove the FSQ normalization scheme to make the method equal to TCRL with some minor improvements to the method, such as using nstep-TD3 \cite{fujimotoAddressingFunctionApproximation2018} as the base algorithm instead of DDPG with Double Q and nstep returns, using the neural network architecture from \citet{hansen2023td}, and performing policy smoothing \cite{fujimotoAddressingFunctionApproximation2018}.
@@
 \textbf{Reconstruction loss has a detrimental impact}
-Learning to minimize the observation reconstruction error has been widely applied in model-based reinforcement learning \cite{sutton2018reinforcement, haRecurrentWorldModels2018, hafnerLearning2019}, and an observation decoder has been a component of many of the most successful RL algorithms to date \cite{hafner2023mastering}.
+Learning to minimize the observation reconstruction error has been widely applied in model-based RL \cite{sutton2018reinforcement, haRecurrentWorldModels2018, hafnerLearning2019}, and an observation decoder has been a component of many of the most successful RL algorithms to date \cite{hafner2023mastering}.
 However, recent work \citep{zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022}
@@
 \begin{align*}
-	\hat{o}_t &= g_\xi(z_t), \\
 	\mathcal{L}_o &= \mathbb{E}_{o_t\sim\mathcal{D}} [\| \hat{o}_t - o_t \|_2^2],
 \end{align*}
-where $g_\xi$ is a learned observation decoder that takes the latent state as the input and outputs the reconstructed observation.
+where $\hat{o}_t = g_\xi(z_t)$ is a learned observation decoder that takes the latent state as the input and outputs the reconstructed observation.
+The decoder $g_\xi$ is a standard MLP.
+We perform reconstruction at each time step in the horizon.
+% $g_\xi$ is a learned observation decoder that takes the latent state as the input and outputs the reconstructed observation.
 The decoder $g_\xi$ is a standard MLP.

### 2024-02-02 paper/main.tex
updates

@@
 \label{sec:related_work}
-
-Model-based RL (MBRL) can significantly improve the sample efficiency of the model-free methods by reasoning about the world and maximizing the fraction of the information stream utilized for learning by utilizing the observations instead of simply trying to learn a policy or a value function to maximize the return \cite{haRecurrentWorldModels2018}.
-The world model can be used for planning \cite{allen1983planning, basye1992decision}.
+In this section, we provide an overview of sample efficiency in RL.
+Importantly, we recap how representation learning enabled model-free RL methods to obtain
+state-of-the-art sample efficiency.
+
+\textbf{Model-based RL}
+Model-based RL (MBRL) is often said to be more sample efficient than model-free methods.
+% This is because it can reason about the world and maximize the fraction of the information stream utilized for learning by utilizing the observations
+This is because it learns a model in which it can reason about the world,
+instead of simply trying to learn a policy or a value function to maximize the return \cite{haRecurrentWorldModels2018}.
+% The world model can be used for planning \cite{allen1983planning, basye1992decision}.
 A prominent idea has been to optimize the evidence lower bound of observation and reward sequences to learn world models that operate on the latent space of a learned Variational Autoencoder  \citep[VAE,][]{kingmaAutoEncoding2014, igl2018deep}.
@@
 
+\textbf{Self-predictive abstractions}
 How can high-dimensional and noisy observations be used efficiently in model-based RL without resorting to reconstruction or direct prediction of future frames?
 A common solution has been to attach an auxiliary loss to the RL objective and perform representation learning \cite{tomar2021learning, ni2024bridging}.
-A common approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate}.
+A promising approach for learning suitable representations is the use of self-predictive abstractions, where the model is trained to predict future latent states through an auxiliary loss \cite{subramanian2022approximate}.
 \citet{ye2021mastering} introduce a self-supervised consistency loss on the learned latent representations.
-Instead of relying on a reconstruction-based loss function, \citet{schwarzer2020data} propose a cosine similarity loss between predicted future latents and the true future latents and Q-learning in the learned latent space.
-\citet{ma2021contrastive} modify the classical world model learning objective by replacing the observation reconstruction loss with a contrastive learning term, which maximizes the mutual information between latent states and observations and helps combat the model focusing on learning task-irrelevant information.
+Instead of relying on a reconstruction-based loss function, \citet{schwarzer2020data} propose a cosine similarity loss between predicted future latents and the true future latents and then perform Q-learning in the
+learned latent space.\looseness-2
+% \citet{ma2021contrastive} modify the classical world model learning objective by replacing the observation reconstruction loss with a contrastive learning term, which maximizes the mutual information between latent states and observations, helping combat the model focusing on learning task-irrelevant information.
 \citet{kipf2019contrastive} use contrastive learning to learn a structured world model.
@@
 
+\textbf{Latent-state consistency for model-based RL}
 TD-MPC \cite{hansenTemporalDifferenceLearning2022} use a consistency loss to learn representations for planning with Model Predictive Path Integral control together with reward and value functions learned through temporal difference methods \cite{williams2015model}.
@@
 \citet{zhu2023repo} demonstrate that the observation encoder can be made more resilient to spurious correlations by restricting the information flow from the observations to the latent representation, and the encoder can be easily aligned at test time in a reward-free manner.
-The loss function of TD-MPC can also be used to learn a manager in a hierarchical RL setting \cite{chitnis2023iql}.
-
+% The loss function of TD-MPC can also be used to learn a manager in a hierarchical RL setting \cite{chitnis2023iql}.
+
+\textbf{Latent-state consistency for model-free RL}
 \citet{zhaoSimplifiedTemporalConsistency2023} show that the planning component of TD-MPC is not strictly necessary for high performance, and applying model-free learning on top of the self-consistent representations is sufficient for performance competitive with state-of-the-art.
@@
 		\mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
-		&= - \sum_{h=0}^{H-1} \gamma^{h} \left ( \frac{\hat{z}_{t+h}}{\| \hat{z}_{t+h} \|_2} \right )^T \left ( \frac{\tilde{z}_{t+h}}{\| \tilde{z}_{t+h} \|_2} \right ).
+		&= - \sum_{h=0}^{H-1} \gamma^{h} \left ( \frac{\hat{z}_{t+h}}{\| \hat{z}_{t+h} \|_2} \right )^T \left ( \frac{\tilde{z}_{t+h}}{\| \tilde{z}_{t+h} \|_2} \right ),
 	\end{align*}
-  % where as before we have $\hat{z}_{t+h} = d_{\phi}(\hat{z}_{t}, a_{t})$ and $\tilde{z}_{t+h} = e_{\bar{\theta}}(s_{t+1})$.
+  where as before we have $\hat{z}_{t+h} = d_{\phi}(\hat{z}_{t}, a_{t})$ and $\tilde{z}_{t+h} = e_{\bar{\theta}}(s_{t+1})$.
 	\item TCRL-ours, where in addition to \ourcosrew, we remove the FSQ normalization scheme to make the method equal to TCRL with some minor improvements to the method, such as using nstep-TD3 \cite{fujimotoAddressingFunctionApproximation2018} as the base algorithm instead of DDPG with Double Q and nstep returns, using the neural network architecture from \citet{hansen2023td}, and performing policy smoothing \cite{fujimotoAddressingFunctionApproximation2018}.
@@
 \begin{align*}
-	\mathcal{L}_o &= \mathbb{E}_{o_t\sim\mathcal{D}} [\| \hat{o}_t - o_t \|_2^2],
+	\mathcal{L}_o &= \mathbb{E}_{o_t\sim\mathcal{D}} [\| \hat{o}_t - o_t \|_2^2], \\
+\hat{o}_t &= g_\xi(z_t),
 \end{align*}
-where $\hat{o}_t = g_\xi(z_t)$ is a learned observation decoder that takes the latent state as the input and outputs the reconstructed observation.
+where $g_\xi$ is a learned observation decoder that takes the latent state as the input and outputs the reconstructed observation.
 The decoder $g_\xi$ is a standard MLP.
@@
 continuous control tasks.
-Importantly, our normalization and bounding of the latent space empirically preserves the representation's matrix rank,
+Our normalization and bounding of the latent space empirically preserves the representation's matrix rank,
 indicating that it alleviates representation collapse.
+As such, \our learns a task agnostic representation and our experiments demonstrate that it shows
+promise for incremental multi-task learning.
 Importantly, our method is {\em (i)} straightforward, {\em (ii)} compatible with any model-free RL algorithm,
@@
 discretized latent space. So far, we have only applied our representation loss to state-based tasks. Extending the method to handle even higher-dimensional visual tasks is an attractive direction.
-% Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
-Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample-efficient.
-Exploring whether explicitly quantifying the model uncertainty by \eg using larger ensembles would help to adapt our algorithm to a real-world robotics setting is also a promising direction for future work.
+Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
+Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
+% Exploring whether explicitly quantifying the model uncertainty by \eg using larger ensembles would help to adapt our algorithm to a real-world robotics setting is also a promising direction for future work.
 

### 2024-02-02 paper/main.tex
updated

@@
 We propose an efficient representation learning method using only a self-supervised latent-state consistency loss.
-Our approach employs an encoder and a transition model to map observations to latent states and predict future states, respectively.
+Our approach employs an encoder and a dynamics model to map observations to latent states and predict future states, respectively.
 We achieve high performance and prevent representation collapse by normalizing and bounding the latent representation such that rank in the representation is empirically preserved.
@@
 In particular, they leverage the latent-state consistency loss.
-This requires an encoder $e_{\theta}$ which maps observations $o_{t}$ to latent states $z_{t}$ and a transition model $d_{\phi}$ which predicts
+This requires an encoder $e_{\theta}$ which maps observations $o_{t}$ to latent states $z_{t}$ and a dynamics model $d_{\phi}$ which predicts
 the next latent states $z_{t+1}$ given a latent state $z_{t}$ and an action $a_{t}$.
-It then minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
+It then minimizes the distance between the next state predicted by the dynamics model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
 and the next state predicted by the target encoder $e_{\bar{\phi}}(z_{t+1})$.
@@
 \citep{zhangLearningInvariantRepresentations2020,zhaoSimplifiedTemporalConsistency2023,hansenTemporalDifferenceLearning2022,geladaDeepMDPLearningContinuous2019,rezaei-shoshtariContinuousMDPHomomorphisms2022}.
-This helps to prevent representation collapse at the cost of learning a task-specific representation.
+This helps to prevent representation collapse at the cost of learning a task-specific representation.\looseness-1
 % Typically it is combined with other loss terms such as reward or value prediction.
@@
 \begin{algorithmic}
-   \STATE {\bfseries Input:} Encoder $e_{\theta}$, transition $d_{\phi}$, critics $\{Q_{\psi_{1}}, Q_{\psi_{2}} \}$, policy $\pi_{\eta}$, learning rate $\alpha$, target network update rate $\tau$
+   \STATE {\bfseries Input:} Encoder $e_{\theta}$, dynamics $d_{\phi}$, critics $\{Q_{\psi_{1}}, Q_{\psi_{2}} \}$, policy $\pi_{\eta}$, learning rate $\alpha$, target network update rate $\tau$
    \FOR{$i$ {\bfseries to} $N_{\text{episodes}}$}
@@
 %
-which minimizes the distance between the next state predicted by the transition model $\hat{z}_{t+1} = d_{\phi}(\hat{z}_{t}, a_{t})$
+which minimizes the distance between the next state predicted by the dynamics model $\hat{z}_{t+1} = d_{\phi}(\hat{z}_{t}, a_{t})$
 and the next state predicted by the momentum encoder $\bar{z}_{t+1} = e_{\bar{\phi}}(s_{t+1})$.
@@
 The initial mapping to the latent space $\hat{z}_{0} = e_{\theta}(s_{0})$ uses the online encoder which
-is being trained jointly with the transition model $d_{\phi}(\hat{z_{t}}, a_{t})$.
+is being trained jointly with the dynamics model $d_{\phi}(\hat{z_{t}}, a_{t})$.
 The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
@@
 indicating that it alleviates representation collapse.
-As such, \our learns a task agnostic representation and our experiments demonstrate that it shows
+As such, \our learns a task-agnostic representation and our experiments demonstrate that it shows
 promise for incremental multi-task learning.
@@
 and {\em (iii)} learns a task-agnostic representation.
+% We hope that our representation learning technique is
 % Moreover, it can be combined with any model-free RL algorithm

### 2024-02-02 paper/main.tex
fixes

@@
 promise for incremental multi-task learning.
+Our experiments further demonstrate that \our is extremely sample efficient whilst being fast to train,
+which we believe is a strong selling point.
 Importantly, our method is {\em (i)} straightforward, {\em (ii)} compatible with any model-free RL algorithm,
@@
 Furthermore, we have neither analyzed the scaling behavior of our world models nor how well \our performs in a multi-task setting.
-Utilizing the knowledge learned during with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
+Utilizing the knowledge learned with Internet-scale pre-training with large robotic datasets could make our method even more sample efficient.
 % Exploring whether explicitly quantifying the model uncertainty by \eg using larger ensembles would help to adapt our algorithm to a real-world robotics setting is also a promising direction for future work.
@@
 Horizon $(H)$ & $5$ & Horizon for representation learning \\
-Latent dimension ($d$) & $1024$ & \\
+Latent dimension ($d$) & $128$ & \\
  \bottomrule

### 2024-02-02 paper/main.tex
minor typos

@@
 \our is conceptually simple, it {\em (i)}~learns a representation of the observation space and then,
-{\em (2)}~performs model-free RL (\eg, TD3) on this representation.
+{\em (ii)}~performs model-free RL (\eg, TD3) on this representation.
 % It the first representation learning phase, it learns an encoder which maps observations $o \in \mathcal{O}$
@@
 Noise clip & $0.3$ & \\
-N-step TD & $1$ & \\
+N-step TD & $1$ or $3$ & \\
 Policy noise & $0.2$ & \\

### 2024-02-02 paper/figs/plotting/baseline-comparison/fetch_data.py
updated to use iqrl.csv and fixed x-axis scale

@@ -1,5 +1,5 @@
 # %%
-ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_path = "aalto-ml/iqrl-icml/"
 ddpg_data = {
     "path": ddpg_path,
     "data": {
@@ -13,7 +13,14 @@ ddpg_data = {
         # GOOD performance d=128
         "quadruped-walk": ["8zgkcuhs", "egk6qxco", "oh1xhyqt", "hm3ihqzo", "p4slb4q9"],
         "cheetah-run": ["8x29ibzc", "9k7936z8", "e7afvv6h", "81pbgf97", "qfqv9wz3"],
-        "walker-walk": ["8z9o8vp6", "roz89u0n", "h1by1w6s", "zfu00w22", "jvw4myex"],
+        # "walker-walk": ["8z9o8vp6", "roz89u0n", "h1by1w6s", "zfu00w22", "jvw4myex"], # only ran for 500k
+        "walker-walk": [
+            "xfzks8lu",
+            "6zlwsbuy",
+            "hqki86zv",
+            "j13po5c8",
+            "j4l1oh62",
+        ],  # 1M steps
         # "cheetah-run": ["l4c3phty", "96xm0zx5", "6tul0lmk", "y7c9k82p", "zx6ykgj9"],
         # POOR performance d=512 project=False
         "acrobot-swingup": ["c6qmc8yx", "spwng8k4", "wgurvdan", "4xfe6bez", "95l445ci"],
@@ -165,7 +172,7 @@ data = pd.concat(
 
 # %%
 
-data.to_csv("./ifsq-rl.csv")
+data.to_csv("./iqrl.csv")
 
 
 # %%

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py

@@ -48,36 +48,21 @@ def plot(df, key="episode_reward"):
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
-    # df = df[df["env_step"] < 1000000]
 
-    all_data = []
-    # df["env_step"] = df["env_step"] / 1000
+    df["episode"] = df.apply(lambda row: int(row["env_step"] / 1000), axis=1)
+
     for idx, env in enumerate(main_envs):
         idx += 1
         data = df[df["env"] == env]
-        # breakpoint()
-        # data[data["agent"] == "iFSQ-RL"] = data[data["agent"] == "iFSQ-RL"].iloc[::2]
         row = idx // ncol
         col = idx % ncol
         ax = axs[row, col]
         hue_order = data.agent.unique()
 
-        min_ep_length = 1e10
-        for agent in data.agent:
-            ep_length = np.max(df[df["agent"] == agent]["env_step"])
-            # print(f"ep_length {ep_length} for env {env}")
-            min_ep_length = min(min_ep_length, ep_length)
-            # min_ep_length=np.min(min_ep)
-        # print(f"min_ep_length {min_ep_length} for env {env}")
-        all_data += [df[df["env_step"] < min_ep_length]]
-        # breakpoint()
-
         if idx == 2:
             g = sns.lineplot(
-                # x=int("env_step" / 1000),
-                x="env_step",
                 # x="env_step",
-                # x="episode",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -90,8 +75,8 @@ def plot(df, key="episode_reward"):
             ax.legend().set_title(None)
         else:
             g = sns.lineplot(
-                # x="episode",
-                x="env_step",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -102,25 +87,23 @@ def plot(df, key="episode_reward"):
                 ax=ax,
             )
         if env == "quadruped-walk":
-            g.set(xlim=(0, 500000))
+            g.set(xlim=(0, 500))
         if env == "walker-walk":
-            g.set(xlim=(0, 250000))
+            g.set(xlim=(0, 250))
[…]

### 2024-02-02 paper/figs/plotting/norm-ablation/fetch_data.py

@@ -1,5 +1,5 @@
 # %%
-ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_path = "aalto-ml/iqrl-icml/"
 ddpg_data = {
     "path": ddpg_path,
     "data": {
@@ -346,7 +346,7 @@ data = pd.concat(
 
 # %%
 
-data.to_csv("./ifsq-rl.csv")
+data.to_csv("./iqrl.csv")
 
 
 # %%

### 2024-02-02 paper/figs/plotting/norm-ablation/make_plot.py

@@ -70,6 +70,8 @@ def plot(df, key="episode_reward"):
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
+    df["episode"] = df.apply(lambda row: int(row["env_step"] / 1000), axis=1)
+
     df = df.rename(columns=rename)
 
     for idx, env in enumerate(main_envs):
@@ -81,12 +83,11 @@ def plot(df, key="episode_reward"):
         # hue_order = data.agent.unique()
         hue_order = data.name.unique()
         # hue_order = data.utd_ratio.unique()
-        # breakpoint()
 
         if idx == 0:
             sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -99,10 +100,9 @@ def plot(df, key="episode_reward"):
             )
             ax.legend().set_title(None)
         else:
-            # breakpoint()
             sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -139,7 +139,7 @@ df = [
     # pd.read_csv(f"{data_path}/tdmpc_main.csv"),
     # df_redq,
     # pd.read_csv(f"{data_path}/sac_main.csv"),
-    pd.read_csv(f"{data_path}/ifsq-rl.csv"),
+    pd.read_csv(f"{data_path}/iqrl.csv"),
 ]
 plot(pd.concat(df))
 

### 2024-02-02 paper/figs/plotting/rank-comparison/fetch_data.py

@@ -1,5 +1,5 @@
 # %%
-ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_path = "aalto-ml/iqrl-icml/"
 ddpg_data = {
     "path": ddpg_path,
     "data": {
@@ -174,7 +174,7 @@ data = pd.concat(
 
 # %%
 
-data.to_csv("./ifsq-rl.csv")
+data.to_csv("./iqrl.csv")
 
 
 # %%

### 2024-02-02 paper/figs/plotting/rank-comparison/make_plot.py

@@ -77,6 +77,8 @@ def plot(df, i=0):
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
+    df["episode"] = df.apply(lambda row: int(row["env_step"] / 1000), axis=1)
+
     df = df.rename(columns=rename)
 
     for idx, env in enumerate(main_envs):
@@ -92,8 +94,8 @@ def plot(df, i=0):
 
         if idx == 0:
             sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -108,8 +110,8 @@ def plot(df, i=0):
         else:
             # breakpoint()
             sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -146,10 +148,10 @@ df = [
     # pd.read_csv(f"{data_path}/tdmpc_main.csv"),
     # df_redq,
     # pd.read_csv(f"{data_path}/sac_main.csv"),
-    pd.read_csv(f"{data_path}/ifsq-rl.csv"),
+    pd.read_csv(f"{data_path}/iqrl.csv"),
 ]
-#plot(pd.concat(df), 0)
+# plot(pd.concat(df), 0)
 plot(pd.concat(df), 1)
-#plot(pd.concat(df), 2)
+# plot(pd.concat(df), 2)
 
 # %%

### 2024-02-02 paper/figs/plotting/reconstruction-ablation/fetch_data.py

@@ -1,5 +1,5 @@
 # %%
-ddpg_path = "aalto-ml/lifelong-td3-tc/"
+ddpg_path = "aalto-ml/iqrl-icml/"
 ddpg_data = {
     "path": ddpg_path,
     "data": {
@@ -327,7 +327,7 @@ data = pd.concat(
 
 # %%
 
-data.to_csv("./ifsq-rl.csv")
+data.to_csv("./iqrl.csv")
 
 
 # %%

### 2024-02-02 paper/figs/plotting/reconstruction-ablation/make_plot.py

@@ -69,6 +69,8 @@ def plot(df, key="episode_reward"):
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
+    df["episode"] = df.apply(lambda row: int(row["env_step"] / 1000), axis=1)
+
     df = df.rename(columns=rename)
 
     # axs[-1, -1].axis("off")
@@ -82,12 +84,11 @@ def plot(df, key="episode_reward"):
         # hue_order = data.agent.unique()
         hue_order = data.name.unique()
         # hue_order = data.utd_ratio.unique()
-        # breakpoint()
 
         if idx == 3:
             g = sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -100,10 +101,9 @@ def plot(df, key="episode_reward"):
             )
             ax.legend().set_title(None)
         else:
-            # breakpoint()
             g = sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -115,7 +115,7 @@ def plot(df, key="episode_reward"):
                 ax=ax,
             )
         if env == "cartpole-swingup":
-            g.set(xlim=(20000, 100000))
+            g.set(xlim=(20, 100))
 
         ax.set_title(" ".join([ele.capitalize() for ele in env.split("-")]))
         ax.set_xlabel("Environment Steps (1e3)")
@@ -142,7 +142,7 @@ df = [
     # pd.read_csv(f"{data_path}/tdmpc_main.csv"),
     # df_redq,
     # pd.read_csv(f"{data_path}/sac_main.csv"),
-    pd.read_csv(f"{data_path}/ifsq-rl.csv"),
+    pd.read_csv(f"{data_path}/iqrl.csv"),
 ]
 plot(pd.concat(df))
 

### 2024-02-02 paper/figs/plotting/reward-ablation/fetch_data.py

@@ -4,270 +4,274 @@ ddpg_data = {
     "path": ddpg_path,
     "data": {
         "quadruped-run": [
-            #iFSQ-RL
-#            "9mbrf8o3",
-#            "lffrf7xy",
-#            "zcloyygj",
-#            "4lmtpd04",
-#            "bud72vyt",
+            # iFSQ-RL
+            #            "9mbrf8o3",
+            #            "lffrf7xy",
+            #            "zcloyygj",
+            #            "4lmtpd04",
+            #            "bud72vyt",
             "225j3iqa",
             "gsh5g5ie",
             "wkwqz2hp",
             "1fqkia25",
             "bamy18xg",
             # iFSQL-RL+rew
-#            "6jwu9xzm",
-#            "3thseoe1",
-#            "02tn6r73",
-#            "gonirtj6",
-#            "e3ak3zco",
+            #            "6jwu9xzm",
+            #            "3thseoe1",
+            #            "02tn6r73",
+            #            "gonirtj6",
+            #            "e3ak3zco",
             "5fd3wano",
             "zhw57oda",
             "27qykyze",
             "ci0ik0ub",
             "0hlxppg4",
             # iFSQ-RL+cos+rew
-#            "sovpuj12",
-#            "epkk4mw8",
-#            "5mpggr4p",
-#            "ams8eapm",
-#            "6bg43249",
+            #            "sovpuj12",
+            #            "epkk4mw8",
+            #            "5mpggr4p",
+            #            "ams8eapm",
+            #            "6bg43249",
             "l5mr1ykt",
             "3o91tqjt",
             "tcecjfc5",
             "7giv4t1b",
             "chdxnkyl",
             # TCRL-ours
-#            "ijm4oaud",
-#            "3bm9cclp",
-#            "aa3eaznd",
-#            "px4fskkk",
-#            "m62ducyj",
+            #            "ijm4oaud",
+            #            "3bm9cclp",
+            #            "aa3eaznd",
+            #            "px4fskkk",
+            #            "m62ducyj",
             "8nnny8uj",
             "7j4amx4h",
             "ngnrnyes",
             "9ddmzk2n",
             "ekef5rwc",
             # TCRL-ours-small
-#            "73kkvwmy",
-#            "12krzm4k",
-#            "mp0d0djz",
-#            "bhixjgby",
-#            "6x1suc5x",
+            #            "73kkvwmy",
+            #            "12krzm4k",
+            #            "mp0d0djz",
[…]

### 2024-02-02 paper/figs/plotting/reward-ablation/make_plot.py

@@ -25,7 +25,7 @@ COLORS = {
     "iQRL+rew": "#377eb8",
     "iQRL+cos+rew": "#ff7f00",
     "TCRL-ours": "#4DAF4A",
-    #"TCRL-ours-small": "#984ea3",
+    # "TCRL-ours-small": "#984ea3",
 }
 # %%
 main_envs = [
@@ -71,6 +71,8 @@ def plot(df, key="episode_reward"):
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
+    df["episode"] = df.apply(lambda row: int(row["env_step"] / 1000), axis=1)
+
     df = df.rename(columns=rename)
 
     for idx, env in enumerate(main_envs):
@@ -86,10 +88,8 @@ def plot(df, key="episode_reward"):
 
         if idx == 0:
             sns.lineplot(
-                # x=int("env_step" / 1000),
-                x="env_step",
                 # x="env_step",
-                # x="episode",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -102,8 +102,8 @@ def plot(df, key="episode_reward"):
             ax.legend().set_title(None)
         else:
             sns.lineplot(
-                # x="episode",
-                x="env_step",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -114,39 +114,39 @@ def plot(df, key="episode_reward"):
                 ax=ax,
             )
 
-#        if idx == 0:
-#            print(data)
-#            print(ax)
-#            exit(0)
-#            sns.lineplot(
-#                x="env_step",
-#                # x="episode",
-#                y=key,
-#                data=data,
-#                errorbar=("ci", 95),
-#                hue="name",
-#                # style="utd_ratio",
-#                hue_order=hue_order,
-#                palette=COLORS,
-#                legend="auto",
-#                ax=ax,
-#            )
-#            ax.legend().set_title(None)
-#        else:
-#            # breakpoint()
-#            sns.lineplot(
-#                x="env_step",
-#                # x="episode",
-#                y=key,
-#                data=data,
-#                errorbar=("ci", 95),
-#                hue="name",
-#                # style="utd_ratio",
-#                hue_order=hue_order,
-#                palette=COLORS,
-#                legend=False,
-#                ax=ax,
-#            )
+        #        if idx == 0:
+        #            print(data)
[…]

### 2024-02-02 paper/figs/plotting/task-agnostic-ablation/fetch_data.py

@@ -29,7 +29,7 @@ ddpg_data = {
             "f5d66x17",
             "jp26qyvm",
             "n3cb46el",
-            "z4dideh5"
+            "z4dideh5",
         ],
         "quadruped-run": [
             # Pre-training"
@@ -45,11 +45,11 @@ ddpg_data = {
             "my89a2ui",
             "0bdbllyg",
             # Task-specific pre-training
-#            "uykvengt",
-#            "e6u4u0et",
-#            "phlr610t",
-#            "32c5rcc6",
-#            "2hyd7evc",
+            #            "uykvengt",
+            #            "e6u4u0et",
+            #            "phlr610t",
+            #            "32c5rcc6",
+            #            "2hyd7evc",
             "lc8kap78",
             "fs0fug53",
             "rypu71yu",
@@ -70,11 +70,11 @@ ddpg_data = {
             "cbp72vz9",
             "oi0gnnd1",
             # Task-specific pre-training
-#            "oxisl0xv",
-#            "wv0a1i6l",
-#            "2mstpcgd",
-#            "olnugqts",
-#            "1iegxig2",
+            #            "oxisl0xv",
+            #            "wv0a1i6l",
+            #            "2mstpcgd",
+            #            "olnugqts",
+            #            "1iegxig2",
             "g029b1dc",
             "2dy32443",
             "1ogr1lye",
@@ -179,7 +179,7 @@ data = pd.concat(
 
 # %%
 
-data.to_csv("./ifsq-rl.csv")
+data.to_csv("./iqrl.csv")
 
 
 # %%

### 2024-02-02 paper/figs/plotting/task-agnostic-ablation/make_plot.py

@@ -70,6 +70,8 @@ def plot(df, key="episode_reward"):
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
 
+    df["episode"] = df.apply(lambda row: int(row["env_step"] / 1000), axis=1)
+
     df = df.rename(columns=rename)
 
     for idx, env in enumerate(main_envs):
@@ -85,8 +87,8 @@ def plot(df, key="episode_reward"):
 
         if idx == 0:
             sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -101,8 +103,8 @@ def plot(df, key="episode_reward"):
         else:
             # breakpoint()
             sns.lineplot(
-                x="env_step",
-                # x="episode",
+                # x="env_step",
+                x="episode",
                 y=key,
                 data=data,
                 errorbar=("ci", 95),
@@ -139,7 +141,7 @@ df = [
     # pd.read_csv(f"{data_path}/tdmpc_main.csv"),
     # df_redq,
     # pd.read_csv(f"{data_path}/sac_main.csv"),
-    pd.read_csv(f"{data_path}/ifsq-rl.csv"),
+    pd.read_csv(f"{data_path}/iqrl.csv"),
 ]
 plot(pd.concat(df))
 

### 2024-02-02 paper/figs/plotting/norm-ablation/make_plot.py
neaten

@@ -32,8 +32,8 @@ main_envs = [
     "acrobot-swingup",
     "cheetah-run",
     # "walker-walk",
-    "walker-run",
     "hopper-stand",
+    "walker-run",
     "fish-swim",
     "quadruped-run",
     "humanoid-walk",
@@ -84,7 +84,7 @@ def plot(df, key="episode_reward"):
         hue_order = data.name.unique()
         # hue_order = data.utd_ratio.unique()
 
-        if idx == 0:
+        if idx == 3:
             sns.lineplot(
                 # x="env_step",
                 x="episode",

### 2024-02-02 paper/figs/plotting/reconstruction-ablation/make_plot.py

@@ -116,6 +116,8 @@ def plot(df, key="episode_reward"):
             )
         if env == "cartpole-swingup":
             g.set(xlim=(20, 100))
+        if env == "dog-walk":
+            g.set(xlim=(0, 780))
 
         ax.set_title(" ".join([ele.capitalize() for ele in env.split("-")]))
         ax.set_xlabel("Environment Steps (1e3)")

### 2024-02-02 paper/figs/plotting/baseline-comparison/fetch_data.py
updated runs for walker-run (to get 1M steps)

@@ -15,11 +15,11 @@ ddpg_data = {
         "cheetah-run": ["8x29ibzc", "9k7936z8", "e7afvv6h", "81pbgf97", "qfqv9wz3"],
         # "walker-walk": ["8z9o8vp6", "roz89u0n", "h1by1w6s", "zfu00w22", "jvw4myex"], # only ran for 500k
         "walker-walk": [
-            "xfzks8lu",
-            "6zlwsbuy",
-            "hqki86zv",
-            "j13po5c8",
-            "j4l1oh62",
+            "zgtw2w02",
+            "hix7b6di",
+            "o17mz4v1",
+            "lqvypdgw",
+            "ilb76d20",
         ],  # 1M steps
         # "cheetah-run": ["l4c3phty", "96xm0zx5", "6tul0lmk", "y7c9k82p", "zx6ykgj9"],
         # POOR performance d=512 project=False

### 2024-02-02 paper/figs/plotting/baseline-comparison/make_plot.py
updated

@@ -99,6 +99,11 @@ def plot(df, key="episode_reward"):
         ax.set_xlabel("Environment Steps (1e3)")
         ax.set_ylabel("Episode Return")
 
+    # df["episode_reward"] = df.apply(
+    #     lambda row: int(row["episode_reward"] / len(main_envs)), axis=1
+    # )
+
+    # Only plot for first 1M steps
     df = df[df["episode"] < 1000]
     # df = df[df["env"] != "walker-walk"]
     g = sns.lineplot(

### 2024-02-02 paper/main.tex
typos

@@
 % For example, specifying $c=3$ channels
-This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values, giving a codebook of size $|\mathcal{C}| = L_{i}^{d_{i}}$.
+This idea generalizes to the case where the $i^{\text{th}}$ channel is mapped to $L_{i}$ values, giving a codebook of size $|\mathcal{C}| = L_{i}^{c_{i}}$.
 % Since quantization is performed by rounding to integers
@@
 FSQ has the following hyperparameters: we must specify the number of channels $c$ and the number of levels per channel
-$\mathcal{L} = [L_{1},\ldots,L_{d}]$.
+$\mathcal{L} = [L_{1},\ldots,L_{c}]$.
 \cref{tab:fsq-levels} shows the recommended number of channels and number of levels per channel to obtain codebooks of different sizes \citep{mentzerFiniteScalarQuantization2023}.
@@
 In practice, we found codebooks of size $|\mathcal{C}|=2^{6}$ sufficient for most environments in the DeepMind Control suite.
-However, for the Dog Walk environment we used a codebook of size $|\mathcal{C}|=2^{9}$.
+However, for the Dog Walk environment we used a codebook of size $|\mathcal{C}|=2^{8}$.
 
@@
 \textbf{Training} & & \\
+Action repeat & 2 & \\
+Eval every episodes & $5$ & \\
 Max episode length & 1000 &  Max episode length \\
@@
              & $15000$ (hard) & \\
+Num eval episodes & $10$ & \\
 Random episodes & $10$ & Number of random episodes at start \\
-Action repeat & 2 & \\
-Eval every episodes & $5$ & \\
-Num eval episodes & $10$ & \\
 \hline
@@
 Encoder momentum coefficient ($\tau$) & 0.005 & \\
-FSQ levels & $[8, 5, 5, 5]$ & \\
+FSQ levels & $[8, 8]$ ($[8,6,5]$ in Dog) & \\
 Horizon $(H)$ & $5$ & Horizon for representation learning \\

### 2024-02-05 agents/iqrl.py
neatened

@@ -1,6 +1,6 @@
 #!/usr/bin/env python3
 import logging
-from typing import Any, Callable, List, Optional, Tuple
+from typing import List, Optional, Tuple
 
 import agents
 import gymnasium as gym
@@ -8,88 +8,13 @@ import helper as h
 import numpy as np
 import torch
 import torch.nn as nn
-import torch.nn.functional as F
 import wandb
-from custom_types import Action, Agent, EvalMode, T0
+from custom_types import Agent, EvalMode, T0
 from gymnasium.spaces import Box, Space
-from helper import SimNorm, soft_update_params
-from torch.linalg import cond, matrix_rank
 from utils import ReplayBuffer, ReplayBufferSamples
-from vector_quantize_pytorch import FSQ
 
 
 logger = logging.getLogger(__name__)
-
-
-class MLPResettable(nn.Module):
-    def __init__(self, mlp):
-        super().__init__()
-        self.mlp = mlp
-
-    def forward(self, x):
-        return self.mlp(x)
-
-    def reset(self, reset_type: str = "last-layer"):
-        if reset_type in "full":
-            h.orthogonal_init(self.parameters())
-        elif reset_type in "last-layer":
-            params = list(self.parameters())
-            h.orthogonal_init(params[-2:])
-        else:
-            raise NotImplementedError
-
-
-class Projection(MLPResettable):
-    def __init__(
-        self,
-        mlp_dims: List[int],
-        levels: Optional[List[int]] = None,
-        out_dim: int = 1024,
-        latent_dim: int = 1024,
-    ):
-        self.levels = levels
-        if levels is not None:
-            self.latent_dim = (latent_dim, len(levels))
-            in_dim = np.array(self.latent_dim).prod()
-        else:
-            self.latent_dim = latent_dim
-            in_dim = self.latent_dim
-        self.act_fn = None
-
-        mlp = h.mlp(in_dim=in_dim, mlp_dims=mlp_dims, out_dim=out_dim, act_fn=None)
-
-        super().__init__(mlp=mlp)
-
-        self.reset(reset_type="full")
-
-    def forward(self, x):
-        if self.levels is not None:
-            if x.ndim > 2:
-                x = torch.flatten(x, -2, -1)
-        z = self.mlp(x)
-        # print(f"z {z.shape}")
-        return z
-
-
-class MLPReward(MLPResettable):
[…]

### 2024-02-05 agents/td3.py

@@ -122,7 +122,7 @@ class TD3(Agent):
         ] = None,  # reset params after this many param updates
         reset_type: str = "last_layer",  #  "full" or "last-layer"
         # nstep: int = 1,
-        discount: float = 0.99,
+        gamma: float = 0.99,
         tau: float = 0.005,
         act_with_target: bool = False,  # if True act with target network
         logging_freq: int = 499,
@@ -150,7 +150,7 @@ class TD3(Agent):
         self.reset_params_freq = reset_params_freq
         self.reset_type = reset_type
         self.nstep = nstep
-        self.discount = discount
+        self.gamma = gamma
         self.tau = tau
         self.act_with_target = act_with_target
         self.logging_freq = logging_freq
@@ -232,18 +232,18 @@ class TD3(Agent):
         dones = torch.zeros_like(batch.dones[0], dtype=torch.bool)
         rewards = torch.zeros_like(batch.rewards[0])
         timeout_or_dones = torch.zeros_like(batch.dones[0], dtype=torch.bool)
-        next_state_discounts = torch.ones_like(batch.dones[0])
+        next_state_gammas = torch.ones_like(batch.dones[0])
         next_obs = torch.zeros_like(batch.observations[0])
         for t in range(self.nstep):
             next_obs = torch.where(
                 timeout_or_dones[..., None], next_obs, batch.next_observations[t]
             )
-            next_state_discounts *= torch.where(timeout_or_dones, 1, self.discount)
+            next_state_gammas *= torch.where(timeout_or_dones, 1, self.gamma)
             dones = torch.where(timeout_or_dones, dones, batch.dones[t])
             rewards += torch.where(
                 timeout_or_dones[..., None],
                 0,
-                self.discount**t * batch.rewards[t],
+                self.gamma**t * batch.rewards[t],
             )
             timeout_or_dones = torch.logical_or(
                 timeout_or_dones, torch.logical_or(dones, batch.timeouts[t])
@@ -255,7 +255,7 @@ class TD3(Agent):
             dones=dones,
             timeouts=timeout_or_dones,
             rewards=rewards,
-            next_state_discounts=next_state_discounts,
+            next_state_gammas=next_state_gammas,
         )
 
         # Update critic
@@ -305,8 +305,8 @@ class TD3(Agent):
             min_q_next_target = torch.min(q1_next_target, q2_next_target)
             next_q_value = data.rewards.flatten() + (
[…]

### 2024-02-05 helper.py

@@ -226,3 +226,21 @@ def calc_rank(name, z):
         info.update({f"{name}-rank-{j}": rank.item()})
     info.update({f"{name}-cond-num": condition.item()})
     return info
+
+
+class MLPResettable(nn.Module):
+    def __init__(self, mlp):
+        super().__init__()
+        self.mlp = mlp
+
+    def forward(self, x):
+        return self.mlp(x)
+
+    def reset(self, reset_type: str = "last-layer"):
+        if reset_type in "full":
+            orthogonal_init(self.parameters())
+        elif reset_type in "last-layer":
+            params = list(self.parameters())
+            orthogonal_init(params[-2:])
+        else:
+            raise NotImplementedError

### 2024-02-05 train.py

@@ -100,7 +100,7 @@ def train(cfg):
         torch.device(cfg.device),
         nstep=nstep,
         #        handle_timeout_termination=False,
-        discount=cfg.agent.discount,
+        gamma=cfg.agent.gamma,
         train_validation_split=cfg.train_validation_split,
     )
 

### 2024-02-05 utils/buffers.py

@@ -26,7 +26,7 @@ class ReplayBufferSamples(NamedTuple):
     dones: th.Tensor
     timeouts: th.Tensor
     rewards: th.Tensor
-    next_state_discounts: th.Tensor
+    next_state_gammas: th.Tensor
 
 
 class BaseBuffer(ABC):
@@ -229,7 +229,7 @@ class ReplayBuffer(BaseBuffer):
         nstep: int = 1,
         optimize_memory_usage: bool = False,
         handle_timeout_termination: bool = True,
-        discount: Optional[float] = 1.0,
+        gamma: Optional[float] = 1.0,
         train_validation_split: Optional[float] = None,
     ):
         super().__init__(
@@ -241,7 +241,7 @@ class ReplayBuffer(BaseBuffer):
 
         # Set n-step returns
         self.nstep = nstep
-        self.discount = discount
+        self.gamma = gamma
         self.train_validation_split = train_validation_split
 
         # Check that the replay buffer can fit into the memory
@@ -440,8 +440,8 @@ class ReplayBuffer(BaseBuffer):
                 self.next_observations[batch_inds, env_indices, :], env
             )
 
-        next_state_discounts = (
-            np.ones_like(self.dones[batch_inds, env_indices]) * self.discount
+        next_state_gammas = (
+            np.ones_like(self.dones[batch_inds, env_indices]) * self.gamma
         )
 
         data = (
@@ -459,7 +459,7 @@ class ReplayBuffer(BaseBuffer):
             self._normalize_reward(
                 self.rewards[batch_inds, env_indices].reshape(-1, 1), env
             ),
-            next_state_discounts.astype(np.float32),
+            next_state_gammas.astype(np.float32),
         )
         data_traj = ()
         for i in range(len(data)):
@@ -486,7 +486,7 @@ class ReplayBuffer(BaseBuffer):
         next_obs = np.zeros_like(obs)
         timeout_or_dones = np.zeros_like(dones)
         rewards = np.zeros_like(dones)
-        next_state_discounts = np.ones_like(dones)
+        next_state_gammas = np.ones_like(dones)
 
         for n in range(nstep):
             idxs = (batch_inds + n) % self.buffer_size
@@ -511,14 +511,14 @@ class ReplayBuffer(BaseBuffer):
                     ),
                 )
 
-            next_state_discounts *= np.where(timeout_or_dones, 1, self.discount)
+            next_state_gammas *= np.where(timeout_or_dones, 1, self.gamma)
             dones = np.where(
[…]

### 2024-02-05 agents/iqrl.py
log the active percent of the codebook

@@ -1,5 +1,6 @@
 #!/usr/bin/env python3
 import logging
+import math
 from typing import List, Optional, Tuple
 
 import agents
@@ -101,6 +102,7 @@ class iQRL(Agent):
 
         ##### Encoder config #####
         self.use_fsq = use_fsq
+        self.num_codes = math.prod(fsq_levels)
         self.quantized = quantized
         if quantized:
             self.fsq_idx = 1

### 2024-02-05 train.py

@@ -220,8 +220,19 @@ def train(cfg):
                     z_batch = agent.enc(batch.observations[0])
                     rank_info = h.calc_rank(name="z", z=z_batch)
                     if cfg.agent.use_fsq:
+                        # Log  rank of NN output (before normalization)
                         mlp_z_batch = agent.enc.mlp(batch.observations[0])
                         rank_info.update(h.calc_rank(name="mlp-z", z=mlp_z_batch))
+
+                        # Log percent of codebook being used (i.e. the active percent)
+                        indices = agent.enc(batch.observations[0], quatized=True)
+                        eval_metrics.update(
+                            {
+                                "active_percent": indices.unique().numel()
+                                / agent.num_codes
+                                * 100
+                            }
+                        )
 
                     eval_metrics.update(rank_info)
 

### 2024-02-05 train.py
fix

@@ -225,7 +225,7 @@ def train(cfg):
                         rank_info.update(h.calc_rank(name="mlp-z", z=mlp_z_batch))
 
                         # Log percent of codebook being used (i.e. the active percent)
-                        indices = agent.enc(batch.observations[0], quatized=True)
+                        indices = agent.enc(batch.observations[0], quantized=True)
                         eval_metrics.update(
                             {
                                 "active_percent": indices.unique().numel()

### 2024-02-05 helper.py
use orthogoanl init for all MLPs

@@ -127,6 +127,7 @@ class FSQMLP(nn.Module):
             in_dim=in_dim, mlp_dims=mlp_dims, out_dim=self.out_dim, act_fn=None
         )
         self._fsq = FSQ(levels)
+        self.apply(orthogonal_init)
 
     def forward(self, x, quantized: bool = False, both: bool = False):
         flag = False
@@ -176,7 +177,9 @@ def mlp(in_dim, mlp_dims, out_dim, act_fn=None, dropout=0.0):
         if act_fn
         else nn.Linear(dims[-2], dims[-1])
     )
-    return nn.Sequential(*mlp)
+    mlp = nn.Sequential(*mlp)
+    mlp.apply(orthogonal_init)
+    return mlp
 
 
 def orthogonal_init(m):
@@ -232,6 +235,7 @@ class MLPResettable(nn.Module):
     def __init__(self, mlp):
         super().__init__()
         self.mlp = mlp
+        self.reset(reset_type="full")
 
     def forward(self, x):
         return self.mlp(x)

### 2024-02-05 paper/main.tex
fixed algorithm to update \eta target
adding clipping to policy smoothing

@@
           \STATE Update actor less frequently than critic:
-          \STATE $\quad \quad \eta \leftarrow \psi + \alpha \nabla \left( \mathcal{L}_{\pi}(\psi; \mathcal{D}) \right)$  \COMMENT{\cref{eq:policy-loss}}
+          \STATE $\quad \quad \eta \leftarrow \eta + \alpha \nabla \left( \mathcal{L}_{\pi}(\eta; \mathcal{D}) \right)$  \COMMENT{\cref{eq:policy-loss}}
           % \STATE $\bar{\eta} \leftarrow \tau \bar{\eta} + (1-\tau) \eta$
@@
         \STATE Update target networks:
-        \STATE $\quad \quad [\bar{\theta}, \bar{\phi}, \bar{\psi}] \leftarrow \tau [\bar{\theta}, \bar{\phi}, \bar{\psi}] + (1-\tau) [{\theta}, {\phi}, {\psi}]$
+        \STATE $\quad \quad [\bar{\theta}, \bar{\psi}, \bar{\eta}] \leftarrow \tau [\bar{\theta}, \bar{\psi}, \bar{\eta}] + (1-\tau) [{\theta}, {\psi}, \eta]$
     \ENDFOR
@@
 \end{align}
-where we use policy smoothing by adding clipped Gaussian noise $\epsilon_{t+n} \sim \mathcal{N} (0,\sigma^{2})$ to the
+where we use policy smoothing by adding clipped Gaussian noise $\epsilon_{t+n} \sim \text{clip} \left(\mathcal{N} (0,\sigma^{2}), -c, c \right)$ to the
 action $a_{t+n} = \pi_{\bar{\eta}}(z_{t+n}) + \epsilon_{t+n}$.

### 2024-02-20 paper/figs/plotting/reward-ablation/make_plot.py
Reward ablation plot

@@ -63,10 +63,10 @@ rename = {a: b for a, b in zip(keys, titles)}
 def plot(df, key="episode_reward"):
     # breakpoint()
     envs = np.sort(df.env.unique())
-    ncol = 6
+    ncol = 3
     # assert envs.shape[0] % ncol == 0
     # nrow = len(main_envs) // ncol
-    nrow = 1
+    nrow = 2
     # nrow = envs.shape[0] // ncol
 
     fig, axs = plt.subplots(nrow, ncol, figsize=(4 * ncol, 3.5 * nrow))
